---
title: "Choosing a pension plan: an investor's decision tool"
author: Martin Hoshi Vognsen
output:
  html_document:
    toc: true
    toc_depth: 3
    keep_md: yes
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
params:
  start_date: null         ## NULL = use the full comparable window (latest genuine common start); else a later "YYYY-MM-DD".
  monitor_months: 36       ## Length (months) of the "recent" window for the structure check.
  balance: 1000000         ## Current balance (kr) the decision applies to.
  flat_fee: 700            ## A provider's annual flat administration fee (kr); the marginal cost of a second provider.
  prop_rate_velliv: 0.007  ## Velliv annual proportional fee (fraction of balance).
  prop_rate_pfa: 0.007     ## PFA annual proportional fee (fraction of balance).
  risk_free: 0.0           ## Annual risk-free rate (e.g. 0.03 for a 3% deposit).
  gamma: 2                 ## Relative risk aversion.
  horizon: 20              ## Decision horizon in years.
  assumed_nu: 3.5          ## Assumed skew-t tail index for the stress simulation.
  assumed_xi: 0.7          ## Assumed skew-t skew for the stress simulation.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(scipen = 999)   # never print kroner amounts in scientific notation
library(here)
library(jsonlite)
library(ggplot2)
library(tidyr)
library(fGarch)
library(MASS)
source(here("src", "pension-returns_functions.R"))
```

```{r load-data}
## --- Velliv: low/medium/high price index -> monthly log-returns ---
v <- read.csv(here("data", "monthly", "velliv_monthly_prices.csv"))
v$date <- as.Date(v$date); v <- v[order(v$date), ]; v$ym <- format(v$date, "%Y-%m")
vlr <- data.frame(
  ym     = v$ym[-1],
  V_low  = diff(log(v$low_risk)),
  V_med  = diff(log(v$medium_risk)),
  V_high = diff(log(v$high_risk))
)

## --- PFA: profiles A/B/C/D at the 30y horizon (a30), % -> monthly log-returns ---
j <- fromJSON(here("data", "monthly", "pfa_monthly.json"), simplifyVector = FALSE)
rows <- list()
for (y in j$afkast_maaned$afkast) for (m in y$maaneder) if (m$maaned != "YTD")
  rows[[length(rows) + 1]] <- data.frame(
    ym  = sprintf("%d-%02d", y$aar, as.integer(m$maaned)),
    P_A = log(1 + m$a$a30 / 100), P_B = log(1 + m$b$a30 / 100),
    P_C = log(1 + m$c$a30 / 100), P_D = log(1 + m$d$a30 / 100))
P <- do.call(rbind, rows)

M <- merge(vlr, P, by = "ym"); M <- M[order(M$ym), ]
## Comparability: each provider's plans must be genuinely distinct over the window. Velliv's
## low/medium/high are a single backfilled series until the plans split (here 2012-07, the first
## month the three returns differ; they share a common 100 anchor at 2012-06). Earlier months are
## degenerate and excluded. PFA's profiles are distinct throughout. The comparable window starts at
## the latest genuine first return month across the providers; the user may start later, not earlier.
.v_start   <- min(vlr$ym[apply(vlr[, c("V_low", "V_med", "V_high")], 1,
                               function(r) length(unique(r)) > 1)])
.cmp_start <- max(.v_start, min(P$ym))
if (exists("params") && !is.null(params$start_date))
  .cmp_start <- max(.cmp_start, format(as.Date(params$start_date), "%Y-%m"))
M <- M[M$ym >= .cmp_start, ]
R <- M[, c("V_low", "V_med", "V_high", "P_A", "P_B", "P_C", "P_D")]
plan_labels <- c(V_low = "Velliv low", V_med = "Velliv med", V_high = "Velliv high",
                 P_A = "PFA low (A)", P_B = "PFA B", P_C = "PFA C", P_D = "PFA high (D)")
provider    <- c(rep("Velliv", 3), rep("PFA", 4))

## --- User inputs (set in the YAML header or via render params) ---
.p <- function(name, default) if (exists("params") && !is.null(params[[name]])) params[[name]] else default
balance    <- .p("balance", 1e6)
flat_fee   <- .p("flat_fee", 700)
rf         <- .p("risk_free", 0)
gamma      <- .p("gamma", 2)
horizon    <- .p("horizon", 20)
prop_rate  <- c(Velliv = .p("prop_rate_velliv", 0.007), PFA = .p("prop_rate_pfa", 0.007))
assumed_nu <- .p("assumed_nu", 3.5)
assumed_xi <- .p("assumed_xi", 0.7)
```

# Purpose

This is a companion to the two returns reports `pension-returns_monthly` and `pension-returns_annual`. Those reports characterise the *distribution* of each plan's returns -- fat tails, a persistent left skew, and a tail index the data cannot pin down. This tool turns those findings into the two decisions an investor actually faces:

1. **Which risk level** -- and is a higher-risk plan anything more than a more-leveraged
   version of a lower-risk one?
2. **One provider or both** -- is there diversification between PFA and Velliv, and could
   it plausibly beat the cost of splitting?

The returns reports' central caveat governs everything here: with about `r round(nrow(R) / 12)` years of
data and fat tails, point estimates (means, Sharpe ratios, correlations) are noisy and
regime-dependent. The tool is therefore built to expose *how robust* each conclusion is,
not to emit a single "best" plan. Returns are monthly, `r nrow(R)` months
(`r min(M$ym)` to `r max(M$ym)`), PFA at the 30-year horizon. Sharpe ratios use the risk-free rate set in the inputs below.

# Inputs

Every figure below is computed from your data and these settings. Change them in the YAML header at
the top of this document, or pass them as render parameters.

```{r inputs-table}
.sd <- .p("start_date", NULL)
knitr::kable(data.frame(
  setting = c("Balance (kr)", "Flat fee per provider (kr/yr)", "Proportional fee, Velliv",
              "Proportional fee, PFA", "Risk-free rate (annual)", "Risk aversion", "Horizon (years)",
              "Start date", "Monitor window (months)", "Assumed tail index", "Assumed skew"),
  value = c(format(round(balance), big.mark = ",", scientific = FALSE),
            format(round(flat_fee), big.mark = ","),
            sprintf("%.2f%%", 100 * prop_rate["Velliv"]),
            sprintf("%.2f%%", 100 * prop_rate["PFA"]),
            sprintf("%.1f%%", 100 * rf), gamma, horizon,
            if (is.null(.sd)) "full comparable window" else as.character(.sd),
            .p("monitor_months", 36), assumed_nu, assumed_xi),
  check.names = FALSE),
  caption = "Current inputs. Sharpe ratios, certainty equivalents, fee break-evens and the stress simulation all read from these.")
```

# 1. Within a provider: Is a higher-risk plan just more leverage?

A higher-risk plan is *pure leverage* of a lower-risk one when it has the same
reward-to-risk (Sharpe) and the same distributional shape, only scaled. If so, choosing a
risk level is a pure risk-appetite dial with no consequence for return-per-unit-risk. We
test this with the Sharpe ratio of each plan and, crucially, its stability across
sub-periods.

```{r sharpe-windows}
## Sub-period start dates spanning the supplied window: full, ~1/3 in, ~2/3 in.
.k <- length(M$ym); wins <- M$ym[unique(round(c(1, .k / 3, 2 * .k / 3)))]
tab <- t(sapply(wins, function(w) {
  S <- M[M$ym >= w, names(R)]
  round(sapply(S, ann_sharpe, rf = rf), 2)
}))
rownames(tab) <- paste0("from ", wins, "  (n=", sapply(wins, function(w) sum(M$ym >= w)), ")")
colnames(tab) <- plan_labels[colnames(tab)]
knitr::kable(tab, caption = "Annualised Sharpe ratio by plan and start date.")
## Derived figures used in the prose below.
.vshrp   <- mean(sapply(R[c("V_low", "V_med", "V_high")], ann_sharpe, rf = rf))   # mean Velliv Sharpe
.pfa     <- c("P_A", "P_B", "P_C", "P_D")
.bestpfa <- sapply(wins, function(w) { S <- M[M$ym >= w, .pfa]; .pfa[which.max(sapply(S, ann_sharpe, rf = rf))] })
## Leverage-ray verdict per provider: do the plans share a Sharpe within sampling error?
.ray <- function(plans) {
  S  <- sapply(R[plans], ann_sharpe, rf = rf)
  se <- sqrt(12 * (1 + mean(S)^2 / 24) / nrow(R))   # iid-normal SE of an annualised Sharpe
  list(spread = max(S) - min(S), se = se, ray = (max(S) - min(S)) < se)
}
.vr <- .ray(c("V_low", "V_med", "V_high")); .pr <- .ray(.pfa)
.pfa_stable <- length(unique(.bestpfa)) == 1
```

**Velliv.** The three plans' Sharpe ratios (around `r round(.vshrp, 1)`) span `r round(.vr$spread, 2)`
against a sampling error of about `r round(.vr$se, 2)`. That spread is `r if (.vr$ray) "within" else "beyond"`
the noise, so on this data Velliv `r if (.vr$ray) "behaves as a leverage ray" else "shows a genuine reward-per-risk difference across its plans"`.
`r if (.vr$ray) "Choosing a Velliv risk level is then a pure risk-appetite decision, with nothing to optimise." else "The risk level then carries a reward-per-risk consequence worth weighing, not just a risk-appetite choice."`

**PFA.** The four profiles' Sharpe ratios span `r round(.pr$spread, 2)` against a sampling error of
about `r round(.pr$se, 2)`, so the menu `r if (.pr$ray) "is, within the noise, also a single ray" else "curves rather than forming a single ray"`.
The highest-Sharpe profile by start window is `r paste0(plan_labels[.bestpfa], " from ", wins, collapse = "; ")`,
`r if (.pfa_stable) "the same across windows" else "and it is not stable across windows"`.
`r if (!.pr$ray && .pfa_stable) "A best-Sharpe end is identifiable here, though the reports caution that such gaps are fragile at this sample size." else "So no best-Sharpe blend is reliably identifiable, and you should not pay, in added risk or in fees, to chase one."`

The risk-return picture for all plans, with the leverage ray, is in the monthly report's Sharpe
section; this tool reports the verdict above rather than repeat the chart.

# 2. Across providers: Is there diversification?

```{r corr-matrix}
C <- cor(R); dimnames(C) <- list(plan_labels[rownames(C)], plan_labels[colnames(C)])
knitr::kable(round(C, 3), caption = "Correlation matrix of monthly returns across all plans.")
.offdiag <- C; diag(.offdiag) <- NA
.mincorr <- min(.offdiag, na.rm = TRUE)
.minpair <- which(.offdiag == .mincorr, arr.ind = TRUE)[1, ]
```

Every pair is highly correlated. The lowest pairwise correlation is `r round(.mincorr, 2)`,
because every plan is equity-dominated (see the "Path crossing" section of the returns reports:
even a "low-risk plan" is mostly equities). The least-correlated plans are
**`r rownames(C)[.minpair[1]]`** and **`r colnames(C)[.minpair[2]]`**.

```{r divers-table}
mixes <- list(
  c("P_A", "V_high"), c("V_low", "P_D"), c("V_high", "P_D"), c("V_med", "P_B"))
dt <- t(sapply(mixes, function(m) c(
  correlation = round(cor(R[[m[1]]], R[[m[2]]]), 3),
  `vol reduction` = sprintf("%.1f%%", 100 * (divr(R[[m[1]]], R[[m[2]]]) - 1)))))
rownames(dt) <- sapply(mixes, function(m) paste(plan_labels[m[1]], "+", plan_labels[m[2]]))
knitr::kable(dt, caption = "50/50 cross-provider mixes: Correlation and volatility reduction vs the weighted-average volatility.")
.mix_red  <- sapply(mixes, function(m) divr(R[[m[1]]], R[[m[2]]]) - 1)
.best_mix <- rownames(dt)[which.max(.mix_red)]; .max_red <- max(.mix_red)
```

The largest risk reduction among these cross-provider mixes is **`r .best_mix`**, at about
`r sprintf("%.1f%%", 100 * .max_red)` of volatility, consistent with the intuition that pairing the
most bond-heavy plan of one provider with the most equity-heavy of the other combines the
most-different compositions. `r if (.max_red < 0.05) "But the benefit is small, a few percent of volatility, not a step change. Diversifying across providers does something, just not much, in mean-variance terms." else "That is a material reduction, so cross-provider diversification is doing real work in mean-variance terms on this data."`

## The diversification that *does* work is within a provider

Splitting across providers pairs equity with equity. The diversification that actually
reduces risk is across *asset classes* -- bonds with equity -- and it lives inside a single
provider, harvested simply by choosing a risk level. Backing PFA's two building-block funds
out of its profiles (equity fund `H = D`, bond-like fund `L = 2B − D`, exact for simple
returns) and comparing the two kinds of 50/50 blend:

```{r within-divers}
.H <- simple_returns(R$P_D); .L <- 2 * simple_returns(R$P_B) - simple_returns(R$P_D)
knitr::kable(
  data.frame(
    blend = c("Within PFA: Mid-risk plan",
              "Across providers: Velliv-high + PFA-high (same risk level)"),
    correlation = round(c(cor(.L, .H), cor(R$V_high, R$P_D)), 2),
    `vol. reduction` = sprintf("%.0f%%", 100 * (c(divr(.L, .H), divr(R$V_high, R$P_D)) - 1)),
    check.names = FALSE),
  caption = "Volatility reduction from a 50/50 blend: within a provider (across asset classes) vs across providers (same risk level).")
.within_red <- divr(.L, .H) - 1; .cross_red <- divr(R$V_high, R$P_D) - 1
.ratio <- .within_red / .cross_red
```

The within-provider blend correlates about `r round(cor(.L, .H), 1)` and cuts volatility
`r if (.ratio >= 2) "several times more than" else "comparably to"` the cross-provider split.
`r if (.ratio >= 2) "A single provider's mid-risk plan is therefore already more diversified than two equity-heavy plans held across providers, at one set of fees, not two. The upshot, reinforcing Sections 1 and 4: diversify by lowering your risk level within one provider, not by adding a second." else "So on this data the within-provider and cross-provider routes are closer; weigh both against the cost of a second provider."`

# 3. Diversification vs. cost: Which fees actually matter

Fees split into two kinds, and only one bears on the *split* decision:

- **Proportional fees** (a percentage of assets under management). If both providers charge
  the same rate they are *split-neutral*: `p` on the whole pot costs the same whether it sits
  in one plan or two (`p·W = p·W₁ + p·W₂`). Equal proportional fees therefore drop out of the
  one-vs-two decision entirely.
- **Flat fees** (a fixed kr/year administration charge, independent of balance). These are
  paid *per provider*, so a second provider adds its flat fee `F₂` every year, however little
  you place there.

The split decision thus turns on flat fees -- weighed against the diversification benefit --
and, separately, on any *difference* in proportional rates between providers.

## Is the diversification worth a second flat fee?

Express the diversification benefit the way it actually accrues, as a boost to the **compound**
growth rate: the variance drain it removes, `Δg = (σ²_single − σ²_split)/2`. The monthly report
derives this variance drain, and the certainty equivalent used later in this section, from the
return moments.

```{r dg-chunk}
.av <- function(x) sd(x) * sqrt(12)
dg  <- (.av(R$V_high)^2 - .av(0.5 * R$V_high + 0.5 * R$P_D)^2) / 2
```

For a cross-provider same-risk split this is **`r round(1e4 * dg, 1)` basis points per
year** -- the entire "gain that compounds" from holding two providers. In kroner it is
`dg · W`, while the flat fee `F₂` is paid every year regardless of `W`. The split pays only
above a break-even balance `W* = F₂ / dg`:

```{r flatfee-table}
F2 <- round(flat_fee * c(0.5, 0.75, 1, 1.5, 2))
knitr::kable(
  data.frame(`flat fee per year (kr)` = F2,
             `break-even balance (kr)` = format(round(F2 / dg, -3), big.mark = ",", scientific = FALSE, trim = TRUE),
             check.names = FALSE),
  caption = "Balance above which the cross-provider diversification covers a second provider's flat fee.")
.be <- flat_fee / dg; .covers <- balance > .be   # break-even at the user's flat fee; does the balance clear it?
```

At your flat fee of kr `r format(round(flat_fee), big.mark = ",")`, the diversification covers it
only above a balance of about kr `r format(round(.be), big.mark = ",")`. Your balance of kr
`r format(round(balance), big.mark = ",")` is `r if (.covers) "above" else "below"` that, so on pure
diversification grounds `r if (.covers) "a second provider pays for itself here" else "one provider is enough"`.
The benefit is only about `r round(1e4 * dg, 1)` basis points a year, and the flat fee is paid in
every low-balance year on the way up besides. A risk-averse investor values the volatility cut
somewhat above the pure variance drain, but not enough to move the threshold much.

## How much cheaper must a second provider be?

The other reason to add a provider is that it is *cheaper proportionally*. Moving an amount
`W₂` to a provider whose rate is `Δp` lower saves `Δp · W₂` per year, which must cover its
flat fee `F₂`. So a second provider must undercut the first by at least

$$\Delta p \;\ge\; \frac{F_2}{W_2}.$$

```{r prop-table}
W2 <- round(balance * c(0.25, 0.5, 0.75, 1)); F2v <- round(flat_fee * c(0.5, 1, 1.5, 2))
tab <- outer(F2v, W2, function(f, w) sprintf("%.2f%%", 100 * f / w))
dimnames(tab) <- list(paste0("F2 = kr ", format(F2v, big.mark = ",", trim = TRUE)),
                      paste0("kr ", format(W2, big.mark = ",", scientific = FALSE, trim = TRUE)))
knitr::kable(tab, caption = "Proportional-rate discount a second provider must offer to justify its flat fee, by amount placed there (W2).")
```

At your flat fee of kr `r format(round(flat_fee), big.mark = ",")` and balance of kr
`r format(round(balance), big.mark = ",")`, a second provider must be at least
**`r sprintf("%.2f", 100 * flat_fee / balance)` percentage points/year** cheaper to break even on
fees. The diversification credit (`dg · W`) could be added to the left-hand side, but at about a
basis point it barely moves the bar. If a second provider is *strictly* cheaper on both fee types,
the question is not whether to *add* it but whether to *switch* entirely.

## The benefit the variance drain misses: Hedging an unreadable provider bet

The variance drain above compares the mix to the *average* single plan, treating the two
providers' expected returns as known and equal. But the real risk in the provider choice is
not knowing *which provider's drift will compound higher* -- and that difference is precisely
what the data cannot pin down.

```{r selrisk-setup}
.d <- R$P_D - R$V_high                 # monthly relative log-return, PFA high minus Velliv high
.n <- length(.d); .se <- sd(.d) / sqrt(.n)
.principal <- balance; .fee_yr <- flat_fee    # the saver's balance and second-provider flat fee
```

Over the sample `r if (mean(.d) >= 0) "PFA-high out-returned Velliv-high" else "Velliv-high out-returned PFA-high"`
by **`r round(1200 * abs(mean(.d)), 1)` percentage points a year**, with a standard error of
`r round(1200 * .se, 1)` pp/yr (t = `r round(mean(.d) / .se, 1)`), so it is
**`r if (abs(mean(.d) / .se) >= 2) "distinguishable from zero" else "not distinguishable from zero"`**.
`r if (abs(mean(.d) / .se) >= 2) "There is a measurable return edge on this sample, though it remains a single realised path." else "You could not have known in advance which would win, and the best-Sharpe profile in Section 1 was not stable across windows."` The monthly report's Sharpe section carries the reference-period version of this comparison and a plot of the two paths.

Committing to one provider is thus a bet on an unreadable coin, and the stakes grow with the
horizon -- increasingly from the *estimation* uncertainty in the drift (which accumulates
linearly) rather than path noise (which accumulates only as √horizon):

```{r selrisk-horizon}
.row <- function(H) { T <- 12 * H; par <- T * .se; tot <- sqrt(par^2 + T * sd(.d)^2)
  data.frame(horizon = paste0(H, " yr"),
    `provider gap (1-sd)` = sprintf("%.0f%%", 100 * (exp(tot) - 1)),
    `share from estimation` = sprintf("%.0f%%", 100 * par^2 / tot^2),
    `gap on kr 1,000,000` = format(round(.principal * (exp(tot) - 1)), big.mark = ",", trim = TRUE),
    `flat fee over horizon` = format(.fee_yr * H, big.mark = ",", trim = TRUE),
    check.names = FALSE) }
knitr::kable(do.call(rbind, lapply(c(5, 10, 20, 30), .row)),
  caption = "Dispersion between the two providers' cumulative outcomes by horizon, vs the second provider's flat fee.")
```

In kroner the gap dwarfs the flat fee. But the gap is symmetric, since you might land on either
side, so what justifies hedging is its **risk-adjusted** value, not its raw size. For a saver with
constant relative risk aversion `γ`, the certainty-equivalent value of holding both, agnostic
about which is better, is about `γ · σ²gap / 8`, the result derived in the monthly report. Here
`σ²gap` is the variance of the cumulative provider gap, combining path noise with the estimation
uncertainty in the relative drift:

```{r selrisk-ce}
.ce <- function(H, g) { T <- 12 * H; g * ((T * .se)^2 + T * sd(.d)^2) / 8 }
.gv <- sort(unique(c(1, gamma, 4))); .hv <- sort(unique(c(10, horizon)))
.tab <- sapply(.gv, function(g) sapply(.hv, function(H)
  format(round(.principal * .ce(H, g)), big.mark = ",", trim = TRUE)))
dimnames(.tab) <- list(paste0(.hv, " yr"), paste0("RRA ", .gv))
knitr::kable(.tab, caption = paste0(
  "Certainty-equivalent value (kr, on kr ", format(.principal, big.mark = ",", scientific = FALSE),
  ") of holding both providers instead of one, by relative risk aversion (RRA, columns). ",
  "Compare with the flat fee, kr ", .fee_yr, "/yr: kr ", format(.fee_yr * .hv[1], big.mark = ","),
  " over ", .hv[1], " yr, kr ", format(.fee_yr * .hv[length(.hv)], big.mark = ","),
  " over ", .hv[length(.hv)], " yr."))
.ce_user <- .principal * .ce(horizon, gamma); .fee_tot <- .fee_yr * horizon; .ce_clears <- .ce_user > .fee_tot
```

At your risk aversion of `r gamma` and a `r horizon`-year horizon, the certainty-equivalent value
of holding both providers is about kr `r format(round(.ce_user), big.mark = ",")`, against
kr `r format(round(.fee_tot), big.mark = ",")` of extra flat fees over that horizon. So the hedge
`r if (.ce_clears) "clears its cost: a second provider is worth its flat fee for you" else "falls short of its cost: for you it is close to a wash"`.
The large realised gap is hindsight, not an expected gain.
`r if (.ce_clears) "The case only strengthens with more risk aversion or a longer horizon." else "It would clear only for a more risk-averse saver or a longer horizon."`
The risk-level choice of Sections 1 and 4 remains the larger lever, and all of this assumes the
saver does not chase the apparent drift edge that Section 1 says cannot be trusted.

*To use this with real numbers, set `F2` (each provider's flat fee), the two proportional
rates, and your balance / contribution path; the break-evens above then read off directly.*

## A second lens: the split as insurance

The certainty equivalent prices the body of the distribution. A complementary view, free of any
distributional assumption, treats the split as insurance. A 50/50 buy-and-hold split ends at the
average of the two single outcomes, so its terminal wealth always lands between the worse and the
better provider. Splitting removes the risk of committing to the worse provider, and pays for it
by giving up the better provider's upside.

```{r split-insurance}
.iA <- exp(sum(R$P_D)); .iB <- exp(sum(R$V_high))   # growth of 1 unit over the sample, two high plans
.ihi <- max(.iA, .iB); .ilo <- min(.iA, .iB); .isplit <- 0.5 * (.iA + .iB)
.ihiname <- if (.iA >= .iB) "PFA-high" else "Velliv-high"
.iloname <- if (.iA >= .iB) "Velliv-high" else "PFA-high"
```

Over the sample, 1 kr grew to about `r sprintf("%.1f", .ihi)` kr in `r .ihiname` and about
`r sprintf("%.1f", .ilo)` kr in `r .iloname`. A 50/50 buy-and-hold split would have ended at about
`r sprintf("%.1f", .isplit)` kr, between the two. The premium is symmetric, and you cannot know in
advance which provider will win: the realised drift gap is not distinguishable from zero, as shown
above.

The honest limit of this lens is that the insurance cannot be priced reliably against the flat
fee. Its value depends on the dispersion of the provider gap, and under the heavy tails the
returns reports document that dispersion is not something the data pin down. The Monte Carlo in
Section 4, and the long-horizon simulations in the returns reports, show that compounded outcomes
over a long horizon can diverge to an extreme degree. So the mean-variance certainty equivalent
gives a point estimate that reads as a close call, while the insurance lens says the protection is
real but its price is genuinely uncertain. The decision rests on how much an unquantifiable hedge
against picking the worse provider is worth to the saver, set against a known annual flat fee.

# 4. Does diversifying across providers help in the tail? A joint simulation

Sections 1--3 are second-moment -- Sharpe, correlation, volatility. The returns reports show
the second moment is the wrong lens: the tail index `nu` is around 3--5, returns are
left-skewed, and correlation is itself unstable under fat tails. The decision-relevant
question is whether holding two providers protects the *tail* -- a crash -- and by how much.

The data fix each plan's *marginal* (mean, volatility, fat-tailed shape) and the *linear
correlation* between providers (about `r round(cor(R$V_high, R$P_D), 2)` for two high-risk plans), but they do **not** fix the
*tail dependence*: whether the two crash *together* (a systemic equity selloff) or whether
one can crash *alone*. So rather than a single number we simulate three models that **all
match the observed means, volatilities and correlation** and differ only in the tail -- a
**Gaussian** world (no fat tails); the fitted skewed-$t$ marginals with **independent**
crashes (a Gaussian copula); and the same marginals with **coincident** crashes (a
Student-$t$ copula). For each we compare holding one high-risk plan (100% Velliv high)
against a 50/50 provider split (Velliv high + PFA high) over **one year** -- the horizon at
which a crash actually bites (over twenty years the law of large numbers smooths the monthly
tails away and the split's effect all but vanishes).

```{r joint-sim}
.fit_sstd <- function(x) { nll <- function(p) sum(-dsstd(x, p[1], p[2], p[3], p[4], log = TRUE))
  setNames(optim(c(mean(x), sd(x), 4, 1), nll, method = "BFGS")$par, c("mean","sd","nu","xi")) }
.m1 <- .fit_sstd(R$V_high); .m2 <- .fit_sstd(R$P_D); .rho <- cor(R$V_high, R$P_D)
.clamp <- function(u) pmin(pmax(u, 1e-7), 1 - 1e-7)
.simyr <- function(model, n = 10000, H = 12, dfc = 4, seed = 99) {
  set.seed(seed); tot <- n * H; Z <- MASS::mvrnorm(tot, c(0, 0), matrix(c(1, .rho, .rho, 1), 2))
  if (model == "Gaussian") {
    r1 <- .m1["mean"] + .m1["sd"] * Z[, 1]; r2 <- .m2["mean"] + .m2["sd"] * Z[, 2]
  } else {
    if (model == "Fat, coincident crashes") { Z <- Z / sqrt(rchisq(tot, dfc) / dfc); U <- pt(Z, dfc) } else U <- pnorm(Z)
    r1 <- qsstd(.clamp(U[, 1]), .m1["mean"], .m1["sd"], .m1["nu"], .m1["xi"])
    r2 <- qsstd(.clamp(U[, 2]), .m2["mean"], .m2["sd"], .m2["nu"], .m2["xi"])
  }
  dim(r1) <- dim(r2) <- c(H, n)
  Ws  <- 100 * exp(colSums(r1))
  port <- 0.5 * (exp(r1) - 1) + 0.5 * (exp(r2) - 1)
  Wsp <- 100 * apply(pmax(1 + port, 0), 2, prod)
  c(`single 5th-pct` = as.numeric(quantile(Ws, .05)),  `single P(loss>10%)` = 100 * mean(Ws < 90),
    `split 5th-pct`  = as.numeric(quantile(Wsp, .05)),  `split P(loss>10%)`  = 100 * mean(Wsp < 90))
}
.mods <- c("Gaussian", "Fat, independent crashes", "Fat, coincident crashes")
.JT <- suppressWarnings(t(sapply(.mods, .simyr)))
knitr::kable(round(.JT, 2),
  caption = "One-year wealth (start = 100): one high-risk plan vs a 50/50 provider split, under three models matched on mean, volatility and correlation.")
```

Three things stand out. **Fat tails make the single plan riskier than a Gaussian view admits.**
The chance of a one-year loss worse than 10% rises from about `r sprintf('%.1f', .JT['Gaussian','single P(loss>10%)'])`%
under the Gaussian to about `r sprintf('%.1f', .JT['Fat, independent crashes','single P(loss>10%)'])`% with
fat tails, and for deeper losses the gap widens. **Splitting across two same-risk providers helps
only modestly.** It trims that probability to about `r sprintf('%.1f', .JT['Fat, independent crashes','split P(loss>10%)'])`%
and lifts the 5th-percentile floor by about `r sprintf('%.1f', .JT['Fat, independent crashes','split 5th-pct'] - .JT['Fat, independent crashes','single 5th-pct'])`
points. And crucially, **that modest benefit barely moves between the independent-crash and
coincident-crash models**: at a correlation of `r round(cor(R$V_high, R$P_D), 2)` the bulk
co-movement already caps the diversification, so the tail dependence we *cannot* estimate turns
out not to change the answer. Provider-splitting at the same risk level is a weak crash hedge,
since two equity-dominated providers mostly fall together.

The far larger lever for crash protection is the **risk level itself**: The bond sleeve of a
lower-risk plan cuts the crash probability much more than a second provider does (the flip
side of the "Path crossing" result in the returns reports -- the lower-risk plan genuinely
cushions a drawdown, at the cost of expected return). The one diversification a *returns*
analysis cannot see is **provider-specific operational risk** -- a fund or administrator
failing on its own -- which is the strongest remaining argument for splitting and lies
outside this data.

# 5. Using the tool

- **Risk level, Velliv:** A pure risk-appetite dial (one leverage ray). Pick by drawdown
  tolerance; there is no reward-per-risk to optimise.
- **Risk level, PFA:** The menu curves, but the best-Sharpe blend is not identifiable and
  flips across periods. Default to risk appetite; do not pay to chase a tangency.
- **One provider or both:** A small mean-variance diversification (best via PFA-low +
  Velliv-high), and the joint simulation shows it stays small in the tail too -- two
  equity-heavy providers mostly crash together, so a same-risk split is a weak crash hedge.
  The real crash lever is the *risk level*, not the provider. The strongest case for two
  providers is operational risk, which the return data cannot measure; weigh it against the
  cost of splitting.
- **All of the above rests on a benign, short, fat-tailed sample.** Treat the numbers as
  direction, not precision -- the same discipline the returns reports impose.

# 6. From a one-off decision to an ongoing monitor

These choices are not made once. The data that drive them -- the regime, the gap between
providers, what each plan holds, the fees -- change, and a wrong assumption ("the bull market
will continue") reveals itself only with time. This tool is therefore also meant to be
**re-run periodically as a monitor**, flagging when a decision needs revisiting:

```{r provider-spread}
.gf  <- function(x) exp(sum(x))            # growth factor of one unit over the sample
.wD  <- .gf(R$P_D); .wV <- .gf(R$V_high)
.hi  <- if (.wD >= .wV) "PFA-high" else "Velliv-high"
.lo  <- if (.wD >= .wV) "Velliv-high" else "PFA-high"
.whi <- max(.wD, .wV); .wlo <- min(.wD, .wV)
.gap_pp <- abs(1200 * (mean(R$P_D) - mean(R$V_high)))   # annualised log-return gap, pp/yr
```

- **Regime shifts** -- a change in the level of returns, in volatility, or in the tail that
  would alter the *risk-level* choice (e.g. when staying in a high-risk plan stops looking
  safe).
- **Provider divergence** -- a widening gap between the providers' cumulative outcomes. Over
  `r min(M$ym)` to `r max(M$ym)`, 1 m DKK in `r .hi` grew to about `r sprintf("%.1f", .whi)` m
  versus about `r sprintf("%.1f", .wlo)` m in `r .lo`, a spread of about
  `r sprintf("%.1f", .whi - .wlo)` m, with `r .hi` compounding about `r sprintf("%.1f", .gap_pp)` pp/year
  faster. No one could have called this in advance. A monitor shows such a gap opening and forces
  the question of whether it is signal or luck.
- **Composition or fee changes** -- e.g. PFA's 2024 move from four profiles (A--D) to three
  (Low/Medium/High), or any change in the flat fees that drive the Section 3 arithmetic.

## Distance to a path crossing

Section 1 showed the high plan is a leverage of the medium plan, and the monthly report shows the
flip side: the high plan trails the medium plan exactly while the medium plan sits below its entry
value. The distance to a crossing is therefore the cushion the medium plan has built above the
chosen entry, and a crossing begins once a drawdown gives that cushion back. The table reads this
off for each provider, from the saver's entry (the start of the supplied data by default) and from
the medium plan's most recent peak.

```{r crossing-distance}
## The high plan trails the medium plan while the medium plan is below its entry (monthly report).
## So the distance to a crossing is the medium plan's cushion above entry, given back by a drawdown.
.cross_row <- function(med, label) {
  Cm     <- sum(med)                              # cumulative medium log return since entry (data start by default)
  lvl    <- cumsum(med)                           # medium log index since entry
  peakdd <- 1 - exp(tail(lvl, 1) - max(lvl))      # current drawdown of the medium plan from its running peak
  k      <- min(6, length(med)); recent <- exp(sum(tail(med, k))) - 1
  data.frame(
    provider                 = label,
    "cushion since entry"    = sprintf("%+.0f%%", 100 * (exp(Cm) - 1)),
    "drawdown to a crossing" = sprintf("%.0f%%", 100 * (1 - exp(-Cm))),
    "drawdown from peak now" = sprintf("%.0f%%", 100 * peakdd),
    "medium return, 6m"      = sprintf("%+.1f%%", 100 * recent),
    check.names = FALSE)
}
knitr::kable(
  rbind(.cross_row(R$V_med, "Velliv (high vs med)"),
        .cross_row(R$P_B,   "PFA (D vs B)")),
  row.names = FALSE, align = "r",
  caption = "Distance to a within-provider path crossing. The high plan trails the medium plan once the medium plan falls back to its entry level. 'Drawdown to a crossing' is the fall from today, gradual or in a single month, that would trigger it; 'drawdown from peak now' is how far the medium plan already sits under its running high; 'medium return, 6m' shows whether the cushion is widening (positive) or narrowing.")
```

From an entry at the start of the supplied data the cushion is large, so a crossing needs a
near-total drawdown: a long-held high plan is far from falling behind its medium plan. The
fragility is all in the entry point. An investor who bought at the medium plan's most recent peak
is at the brink, since any down-month then puts the medium plan below entry. This is the COVID-eve
case from the monthly report. The six-month medium return shows the current direction: positive
means the cushion is widening and a crossing receding, negative means it is narrowing. This is the
within-provider crossing, the high plan against its own medium plan, and is separate from the
cross-provider spread above.

## Has the return structure changed?

The tool monitors *structure*, not only level. The clearest signal is the number of return
streams: when a provider splits, merges, or renames its profiles, the column count changes, as
PFA's 2024 move from four profiles to three does. Subtler shifts, a reweighting toward equities or
a change of underlying funds, leave the count intact but show up as drift in the per-provider
leverage, the volatility of the lower-risk plans, and the correlations. The check below compares a
baseline window with the most recent months.

```{r structure-check}
.mm   <- if (exists("params") && !is.null(params$monitor_months)) params$monitor_months else 36
.nrec <- min(.mm, nrow(R) - 24)                       # recent window; keep at least 24 months of baseline
.base <- head(R, nrow(R) - .nrec); .rec <- tail(R, .nrec)
.beta <- function(D, med, high) unname(coef(lm(D[[high]] ~ D[[med]]))[2])
.metric <- function(D) c(
  .beta(D, "V_med", "V_high"), .beta(D, "P_B", "P_D"),
  ann_vol(D$V_med),            ann_vol(D$P_B),
  cor(D$V_low, D$V_high),      cor(D$P_A, D$P_D))
.b <- .metric(.base); .r <- .metric(.rec)
.num <- function(x, i) sprintf("%.2f", x[i]); .pct <- function(x, i) sprintf("%.1f%%", 100 * x[i])
.tab <- data.frame(
  metric = c("Velliv leverage $\\beta$ (high on med)", "PFA leverage $\\beta$ (high on med)",
             "Velliv medium volatility (ann.)", "PFA medium volatility (ann.)",
             "Velliv low-high correlation",     "PFA low-high correlation"),
  baseline = c(.num(.b, 1:2), .pct(.b, 3:4), .num(.b, 5:6)),
  recent   = c(.num(.r, 1:2), .pct(.r, 3:4), .num(.r, 5:6)),
  change   = c(sprintf("%+.2f", .r[1:2] - .b[1:2]),
               sprintf("%+.1f pp", 100 * (.r[3:4] - .b[3:4])),
               sprintf("%+.2f", .r[5:6] - .b[5:6])),
  check.names = FALSE)
knitr::kable(.tab, row.names = FALSE, caption = paste0(
  "Structure check: baseline window vs the most recent ", .nrec, " months. The data carry ",
  ncol(R), " return streams; a change in that count is itself an obvious structural change."))
```

On the supplied data the structure is broadly stable, and any drift here reflects the market
regime rather than a policy change. Run on data that extends past a reweighting, the same check
would flag it: fewer streams, a higher volatility and a higher low-to-high correlation in the
plans moved toward equities, and a leverage slope pulled toward one. A flag is a prompt to ask
which of three causes is at work, market conditions, a change of policy, or a change in the
underlying funds and weights, and then to revisit the risk-level and provider decisions above.
