---
title: "Monthly pension returns analysis: \nVelliv June 2012 - April 2024"
author: Martin Hoshi Vognsen
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
   - \usepackage[fontsize=8pt]{scrextend}
mainfont: SourceSansPro
output: 
  html_document:
    toc: true
    toc_depth: 3
    keep_md: yes
  pdf_document:
    toc: true
    toc_depth: 3
#fontsize: 10pt # for pdf. Limited to 10pt, 11pt and 12pt. Else use scrextend.
params:
  run_sim: FALSE ## TRUE: Run simulations and write output. FALSE: Read saved
                ## simulations output from disk instead of running the simulations.
  run_fits: FALSE
  run_exploratory: TRUE ## Include exploratory report?
  run_individual: TRUE ## Include individual reports?
  run_comparison: TRUE ## Include comparison report? Depends on run_individual.
  run_mc_plot: TRUE ## Depends on run_individual.
  run_is_sim: TRUE
  run_is_plot: TRUE
  include_long: TRUE
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---

```{r chunk1, include=FALSE}
## ****************************** WARNING ******************************
## 
##             This document takes a long time to render!
##
## *********************************************************************
##
## To read previous simulation outputs from disk rather than run simulations,
## set YAML parameter run_sim: TRUE
##
## NOTE:
## When using the "Run All" command, YAML parameters will be ignored. An error
## message "Error in eval(ele) : object 'params' not found" will be printed,
## but the chunks will be evaluated.
```


```{r  chunk2, include=FALSE, eval=FALSE}
################################################################################
##
##                                      SETUP
##
################################################################################
```

```{r chunk3, setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

run_individual <- (params$run_individual + params$run_comparison + params$run_mc_plot > 0)
run_exploratory <- (params$run_exploratory + run_individual > 0)

## Note: In the second line (run_exploratory <- ...), run_individual is correct that - not params$run_individual, because if we choose params$run_exploratory=TRUE and params$run_comparison=TRUE, but params$run_individual=FALSE, we still have to do the individual calculations.
```

```{r chunk4}
library(here)
library(tidyr)
library(tidyjson)
library(lubridate)   # year() used below; attach explicitly for non-interactive render
```

```{r chunk5}
source(here("src", "pension-returns_functions.R"))
```

```{r chunk6}
################################################################################
##
##                          GENERATE HTML AND PDF
##                        (DON'T USE KNIT BUTTON!)
##
################################################################################
```

```{r chunk6a, eval=FALSE, include=FALSE}
## Render this document to html and pdf
rmarkdown::render(here::here("pension-returns_monthly.Rmd"), output_format ="all")
```


```{r chunk6b}
################################################################################
##
##              GENERATE PDF MANUALLY AFTER GENERATING HTML AND MD
##
################################################################################

## Note: Will not produce TOC and a few other things will be different.
```

```{bash chunk7, eval=FALSE, include=FALSE}
## Convert md to pdf
## Run manually:
pandoc pension-returns_monthly.md -o pension-returns_monthly.pdf
```


```{r chunk7b}
################################################################################
##
##                             COPY TO SITES FOLDER
##
################################################################################
```

```{bash eval=FALSE, include=FALSE}
## Run manually:
cp ~/R\ work/pension-returns/pension-returns_monthly.html ~/var/site/pension-returns_monthly.html
cp ~/R\ work/pension-returns/pension-returns_monthly.pdf ~/var/site/pension-returns_monthly.pdf
```


```{r chunk8}
################################################################################
##
##                                      INPUTS
##
################################################################################
```

```{r chunk9}
mc_num_paths <- 1e4
mc_num_periods <- 20 * 12
mc_dao <- TRUE
#mc_threshold <- 0.01

## Min saldo for Velliv: kr 12.300
## Assume initial capital of 400.000.
## 100 * 12300/400000 = 3
mc_threshold <- 3
```

# Summary

A comparison of returns from pension plans provided by PFA and Velliv respectively:  
- Velliv medium risk (vmr), June 2012 - April 2024  
- Velliv high risk (vhr), June 2012 - April 2024  
- PFA medium risk (pmr), June 2012 - April 2024  
- PFA high risk (phr), June 2012 - April 2024  

The combinations of low and high risk funds for each PFA plan can be found here:  
https://www.pfa.dk/privat/opsparing/aktivsammensaetning/

The Velliv website only provides a list of funds.  

The analysis also looks at these mixes of plans:  
- Mix medium risk (mmr), June 2012 - April 2024  
- Mix high risk (mhr), June 2012 - April 2024  
- Mix vmr+phr (vm_ph), June 2012 - April 2024  
- Mix vhr+pmr (mh_pm), June 2012 - April 2024





We observe that  
1. **The plans are weighted blends of the same two underlying funds.** Within
each provider the high-risk plan is almost exactly a scalar multiple of the medium-risk plan
($\beta \approx 1.2$ for Velliv, $\approx 1.5$ for PFA, intercept $\approx 0$), and a single
common factor dominates the variation across all four plans. PFA's own product description
confirms it: profiles B and D are fixed weighted blends of the *same two* underlying funds.

2. **"Risk" here means deeper drawdowns and start-date-dependent path ordering, not a
different kind of bet.** The high-risk plan amplifies the same factor in both directions: it
draws down more (worst peak-to-trough $-17\%$ vs $-15\%$ for Velliv, $-16\%$ vs $-12\%$ for
PFA), and its cumulative path drops below the medium plan's whenever the index is started
just before a drawdown -- as it does around the COVID crash.

3. **The sample still spans only about one market cycle.** Monthly data (142 observations) is
firmer than the annual report's 13, and -- unlike the annual figures -- it *does* capture the
COVID-2020 and 2022 drawdowns. But it contains no 2008-scale event; for that the long Velliv
series is needed. The qualitative structure -- fat tails, leverage, asymmetric drawdowns -- is
robust, while precise tail numbers should be read with Taleb's "better stay home" in mind.

The Discussion at the end develops these points.

# Returns data

The returns data for PFA can be seen here:  
https://www.pfa.dk/privat/opsparing/pfa-afkast/  

The data for Velliv can be seen here:  
https://www.velliv.dk/opsparing/vaekstpension/vaekstpension-aktiv/afkast  

Fit log returns to F-S skew standardized Student-t distribution.  
`m`  is the location parameter.  
`s` is the scale parameter.  
`nu` is the estimated shape parameter (degrees of freedom).  
`xi` is the estimated skewness parameter.  

The long version of Velliv medium risk data runs from January 2007 to April 2024 (incl).  
For January 2007 to May 2012 no low risk and high risk funds existed. For this period the medium risk data is copied into the other funds.  

The short version runs from June 2012 to April 2024.

Velliv returns are including bonus and "DinKapital.  
PFA returns are including "KundeKapital".  

```{r}
## The specified folder must exist in the `data` directory
data_dir <- "monthly"
```

```{r chunk10}
## Headlines for individual sections
headlines <- c(
  "Velliv medium risk (vmr), June 2012 - April 2024",
  "Velliv high risk (vhr), June 2012 - April 2024",
  "PFA medium risk (pmr), June 2012 - April 2024",
  "PFA high risk (phr), June 2012 - April 2024",
  "Mix medium risk (mmr), June 2012 - April 2024",
  "Mix high risk (mhr), June 2012 - April 2024",
  "Mix vmr+phr (vm_ph), June 2012 - April 2024",
  "Mix vhr+pmr (mh_pm), June 2012 - April 2024"
)
```

```{r}
## For Monte Carlo -- PURE BLACK BOX.
## Each fund (mixes included) is simulated from its OWN single fit to its observed
## total-value series: "simulation of the average", NOT "average of simulations".
## The former component-recombination (e.g. c(1, 3) for a mix) drew the two
## components independently, assuming zero correlation between near-collinear funds,
## and is abandoned. fits[[5..8]] are the single fits to the realised mix series
## (mmr, mhr, vmr_phr, vhr_pmr) built via mix_of_logreturns().
fit_id <- list(1, 2, 3, 4, 5, 6, 7, 8)

```


```{r}
pfa_json <- read_json(here("data", "monthly", "pfa_monthly.json"))
```

```{r eval=FALSE}
pfa_json %>% gather_object %>% json_types %>% count(name, type)
```

```{r eval=FALSE}
pfa_json %>% enter_object(afkast_maaned) %>% gather_object %>% json_types %>% count(name, type)
```
```{r}
pfa_monthly_json <- pfa_json %>% enter_object(afkast_maaned) %>% gather_object 
pfa_monthly_json <- pfa_monthly_json$..JSON[[2]]
#pfa_monthly_df <- pfa_monthly_df %>% spread_all
```

```{r}
## Input json: Percent returns
## Output df: Logreturns

pfa_monthly_df <- data.frame()
month_id <- 1
num_years <- length(pfa_monthly_json)
for(i in num_years - seq_along(pfa_monthly_json) + 1) {
  year <- pfa_monthly_json[[i]]$aar
  months <- pfa_monthly_json[[i]]$maaneder[-1]
  for(j in seq_along(months)) {
    pfa_monthly_df[month_id, "year"] <- year
    pfa_monthly_df[month_id, "month"] <- months[[j]]$maaned
    pfa_monthly_df[month_id, "pfa_a"] <- months[[j]]$a$a30
    pfa_monthly_df[month_id, "pfa_b"] <- months[[j]]$b$a30
    pfa_monthly_df[month_id, "pfa_c"] <- months[[j]]$c$a30
    pfa_monthly_df[month_id, "pfa_d"] <- months[[j]]$d$a30
    month_id <- month_id + 1
  }
}
pfa_monthly_df <- pfa_monthly_df %>% 
  ## Start at same date as short Velliv series: 2012-06-30
  ## Get last day of the month by taking first day of next month and subtracting 1
  mutate(date = as.Date(
    paste0(
      year + ifelse(as.numeric(month) == 12, 1, 0), 
      "-", 
      as.numeric(month) %% 12 + 1, 
      "-", 
      "01"
      )
    ) - 1 
  ) %>% 
  filter(date >= as.Date("2012-07-31", format = "%Y-%m-%d"))
pfa_monthly_df <- cbind(pfa_monthly_df$date, pfa_monthly_df[, c("pfa_a", "pfa_b", "pfa_c", "pfa_d")])
colnames(pfa_monthly_df)[1] <- "date"

## Convert percent to gross returns
pfa_monthly_logreturns_df <- pfa_monthly_df
pfa_monthly_logreturns_df[, 2:5] <- log((pfa_monthly_df[, 2:5] /100) + 1)
```

```{r}
## Calculate price index with 2012-06-30 being 100
pfa_monthly_prices_df <- pfa_monthly_logreturns_df
for(i in 2:5) {
  pfa_monthly_prices_df[, i] <- 100 * exp(cumsum(pfa_monthly_logreturns_df[, i]))
}
pfa_monthly_prices_df <- rbind(
    list(as.Date("2012-06-30", format = "%Y-%m-%d"), 100, 100, 100, 100),
    pfa_monthly_prices_df
  )
```

```{r}
## Load data
## Velliv long term log returns
velliv_monthly_prices_long_df <- read.csv(here("data", data_dir, "velliv_monthly_prices.csv"),
  colClasses = c("Date", "numeric", "numeric", "numeric", "numeric")
)

velliv_monthly_prices_df <- velliv_monthly_prices_long_df %>% 
  filter(date >= as.Date("2012-06-30", format = "%Y-%m-%d"))

velliv_monthly_logreturns_long_df <- velliv_monthly_prices_long_df[-1, ]
velliv_monthly_logreturns_long_df[, -1] <- as.data.frame(lapply(
  velliv_monthly_prices_long_df[, -1],
  function(x) {
    log(tail(x, -1) / head(x, -1))
  }
))

#velliv_monthly_logreturns_df <- velliv_monthly_logreturns_long_df[-(1:65), ]
velliv_monthly_logreturns_df <- velliv_monthly_logreturns_long_df %>% 
  filter(date >= as.Date("2012-07-31", format = "%Y-%m-%d"))
rownames(velliv_monthly_logreturns_df) <- as.character(1:nrow(velliv_monthly_logreturns_df))
```


```{r}
monthly_logreturns_df <- cbind(
  velliv_monthly_logreturns_df[, c("date", "medium_risk", "high_risk")],
  pfa_monthly_logreturns_df[, c("pfa_b", "pfa_d")]
)
colnames(monthly_logreturns_df)[2:5] <- c("vmr", "vhr", "pmr", "phr")

data_df_l <- data.frame(vmrl = velliv_monthly_logreturns_long_df$medium_risk) ## Only medium risk data

monthly_logreturns_df <- monthly_logreturns_df %>% 
  mutate(
    mmr = mix_of_logreturns(vmr, pmr),
    mhr = mix_of_logreturns(vhr, phr),
    vmr_phr = mix_of_logreturns(vmr, phr),
    vhr_pmr = mix_of_logreturns(vhr, pmr)
  )

monthly_prices_df <- data.frame(date = pfa_monthly_prices_df$date)
for(i in 2:ncol(monthly_logreturns_df)) {
  monthly_prices_df[, colnames(monthly_logreturns_df)[[i]]] <- prices_from_logreturns(monthly_logreturns_df[, colnames(monthly_logreturns_df)[[i]]])
}

data_df <- monthly_logreturns_df[, -1]

fund_names <- c("vmr", "vhr", "pmr", "phr", "mmr", "mhr", "vmr_phr", "vhr_pmr")
fund_names_l <- "vmrl"
```





```{r fig.width=10}
ggplot(
  pfa_monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("PFA monthly prices"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```

```{r fig.width=10}
ggplot(
  pfa_monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("PFA monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```


```{r chunk11, eval=FALSE}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
cols = gg_color_hue(4)

all_data <- velliv_monthly_prices_df[, -1]
plot(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$guarantee, 
  type = "l",  lty = 1, lwd = 1, 
  xlab = "year", 
  ylab = "Gross returns", 
  col = cols[1], 
  ylim = c(min(all_data), max(all_data)), 
  main = "Velliv gross returns 2007-2023"
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$low_risk, 
  col = cols[2],  
  lty = 1, 
  lwd = 0.8
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$medium_risk, 
  col = cols[3],  
  lty = 1, 
  lwd = 0.6
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$high_risk, 
  col = cols[4],  
  lty = 1, 
  lwd = 0.5
)
legend(
  "bottomright", 
  legend = c("guarantee", "low risk", "medium risk", "high risk"), 
  col = cols, 
  lty = 1, 
  lwd = 2, 
  cex = 0.75
)
```

```{r fig.width=10}
ggplot(
  velliv_monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly prices"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```


```{r eval=params$include_long, fig.width=10}
ggplot(
  velliv_monthly_prices_long_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly prices - long"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```


```{r chunk11b, eval=FALSE}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
cols = gg_color_hue(4)

all_data <- exp(velliv_monthly_logreturns_df[, -1])
plot(
  x = velliv_monthly_logreturns_df$date, 
  y = exp(velliv_monthly_logreturns_df$guarantee), 
  type = "l",  
  lty = 1, 
  lwd = 1, 
  xlab = "year", 
  ylab = "Gross returns", 
  col = cols[1], 
  ylim = c(min(all_data), max(all_data)), 
  main = "Velliv gross returns 2007-2023"
)
lines(
  x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$low_risk), col = cols[2],  lty = 1, lwd = 0.8)
lines(x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$medium_risk), col = cols[3],  lty = 1, lwd = 0.6)
lines(x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$high_risk), col = cols[4],  lty = 1, lwd = 0.5)
legend("bottomright", legend = c("guarantee", "low risk", "medium risk", "high risk"), col = cols, lty = 1, lwd = 2, cex = 0.75)
```

```{r fig.width=10}
ggplot(
  velliv_monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```

```{r eval=params$include_long, fig.width=10}
ggplot(
  velliv_monthly_logreturns_long_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly gross returns - long"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```

```{r fig.width=10}
ggplot(
  monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv and PFA monthly prices"),
    subtitle = "",
    x = "date",
    y = "price index"
  )
```

## Path crossing and the starting point

In the "Velliv and PFA monthly prices" plot above every plan starts at index 100 in
June 2012, and the riskier plans' paths never fall below the less-risky ones. It is
tempting to read this as *more risk always pays*. It could look like the different risk plans for each provider are simply different weighted shares of returns from a single portfolio. What is actually going on here?

The observation does not generalise: It is an artefact of **where the index starts**. Each high-risk plan is, in effect, a leveraged exposure to one common return factor. The high-risk profile is, to a good approximation, a *scaled* version of the medium-risk one -- a slope above 1, essentially zero intercept, and a high $R^2$:

```{r}
.cor_v <- cor(data_df$vmr, data_df$vhr); .cor_p <- cor(data_df$pmr, data_df$phr)
.fv <- lm(vhr ~ vmr, data = data_df);    .fp <- lm(phr ~ pmr, data = data_df)
cat(sprintf("Velliv  corr(medium, high) = %.3f ;  high = %.2f * medium + %.4f ,  R^2 = %.3f\n",
            .cor_v, coef(.fv)[2], coef(.fv)[1], summary(.fv)$r.squared))
cat(sprintf("PFA     corr(medium, high) = %.3f ;  high = %.2f * medium + %.4f ,  R^2 = %.3f\n",
            .cor_p, coef(.fp)[2], coef(.fp)[1], summary(.fp)$r.squared))
```

"Approximately" matters: The $R^2$ is below 1, and that residual is where the paths
separate. The profiles are not one stream scaled, but blends of two genuinely different funds -- a low-risk and a high-risk building block. Backing PFA's two funds out of the medium and high profiles (high fund $H = D$, low fund $L = 2B - D$, exact for simple returns):

```{r}
.simp <- function(lr) exp(lr) - 1
.H <- .simp(data_df$phr); .L <- 2 * .simp(data_df$pmr) - .simp(data_df$phr)
cat(sprintf("PFA building blocks (simple returns): corr(low fund, high fund) = %.2f ; volatility ratio high/low = %.1f\n",
            cor(.L, .H), sd(.H) / sd(.L)))
```

The two underlying funds are only moderately correlated and differ several-fold in
volatility -- far from interchangeable -- even though the *profiles* built from them are nearly collinear. That collinearity is not surprising: It is what two equity-dominated portfolios must produce. A "low-risk *plan*" is not a low-risk *fund*. By PFA's published strategic allocation, even the low-risk plan is roughly 76% equity-like (global, emerging, private equity, real estate) and only ~24% bonds. To see how little bond content it takes, treat a plan as a weight $w$ in equities and $(1-w)$ in bonds and ask which $w$ reproduces a given correlation with a 100%-equity plan (illustrative capital-market assumptions, not estimated from the data):

```{r}
.corr_mix <- function(w, sE, sB, rho)
  (w*sE^2 + (1-w)*rho*sE*sB) /
  (sE * sqrt(w^2*sE^2 + (1-w)^2*sB^2 + 2*w*(1-w)*rho*sE*sB))
.solve_w  <- function(target, sE, sB, rho)
  uniroot(function(w) .corr_mix(w, sE, sB, rho) - target, c(1e-3, 1))$root
.scen <- data.frame(
  `equity vol` = c("15%", "12%", "18%"),
  `bond vol`   = c("4%",  "5%",  "4%"),
  `equity weight for corr 0.986` =
    sprintf("%.0f%%", 100*c(.solve_w(.986,.15,.04,0), .solve_w(.986,.12,.05,0), .solve_w(.986,.18,.04,0))),
  `equity weight for corr 0.84`  =
    sprintf("%.0f%%", 100*c(.solve_w(.84,.15,.04,0),  .solve_w(.84,.12,.05,0),  .solve_w(.84,.18,.04,0))),
  check.names = FALSE)
knitr::kable(.scen)
```

An equity weight of only about two-thirds already yields a ~0.986 correlation with a pure-equity plan. The observed correlations are therefore unremarkable -- exactly what equity-heavy "low-risk" plans produce -- and not evidence of anything hidden.  
  
The implication for the choice is the point of this section. High correlation means the plans rise and fall together, so over a benign, rising sample the higher-risk plan simply ends up higher and the choice looks free. But correlation describes co-movement, not magnitude and not the tails. The lower-risk plan holds a real bond sleeve and a distinct, lower-volatility fund, and -- as the crossings above show -- that cushion is exactly what bites in a drawdown. The plans look alike because both are mostly equities; they are not alike when equities fall.  
  
In cumulative log terms $\log(W_\text{high}/100) \approx \beta\,\log(W_\text{med}/100)$ with slope $\beta > 1$ and intercept $\approx 0$. With $\beta > 1$ this gives $W_\text{high} > W_\text{med}$ **exactly when** $W_\text{med} > 100$. So the two paths can only cross while the *safer* plan is itself below its starting value -- equivalently, the high-risk plan sits below the medium-risk plan precisely over any window in which the cumulative return *since the start date* is negative.  
  
June 2012 was a low point just before a long bull market, so the safer plans never
revisited 100 and the paths never crossed. Start the index instead just before a drawdown and the high-risk path falls below. The COVID crash is a clean realised example: rebasing every plan to 100 at the end of February 2020, the high-risk plans drop **below** the medium-risk plans through the March-2020 trough -- and stay there for about nine months -- before recovering.  

```{r fig.width=10}
## Rebase the cumulative index to 100 at a chosen start month, then plot medium vs high.
crossing_start <- as.Date("2020-02-01")
.cs <- monthly_logreturns_df %>% filter(date >= crossing_start)
.idx <- function(lr) 100 * exp(cumsum(c(0, lr)))     # 100 at the start month
crossing_df <- data.frame(
  date = c(crossing_start, .cs$date),
  `Velliv medium` = .idx(.cs$vmr),
  `Velliv high`   = .idx(.cs$vhr),
  `PFA medium`    = .idx(.cs$pmr),
  `PFA high`      = .idx(.cs$phr),
  check.names = FALSE
) %>% gather(key = "plan", value = "price", -date)

ggplot(crossing_df, aes(x = date, y = price, color = plan)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 100, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("6 months"), labels = date_format("%b %Y")) +
  labs(
    title = "Same plans, index started just before the COVID crash (end-Feb 2020 = 100)",
    subtitle = "High-risk paths fall below medium-risk through the drawdown, then recover",
    x = "date", y = "price index"
  )
```

The high-risk plans regain and overtake the medium-risk plans only once the cumulative return since the start date turns positive again. Whether the realised paths cross is therefore not a property of the plans but of the chosen starting point relative to the market cycle. (A deeper or more prolonged regime -- e.g. one in which the index spends years below its start, as in 2008-2009 -- would keep the high-risk plan below for correspondingly longer.)

```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv and PFA monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```
```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    group_by(date = lubridate::year(date)) %>%
    summarize(
      vmr = sum(vmr),
      vhr = sum(vhr),
      pmr = sum(pmr),
      phr = sum(phr),
      mmr = sum(mmr),
      mhr = sum(mhr),
      vmr_phr = sum(vmr_phr),
      vhr_pmr = sum(vhr_pmr)
    ) %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)),
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  #scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  scale_x_continuous(breaks = lubridate::year(monthly_logreturns_df$date)) +
  labs(
    title = paste("Velliv+PFA annual gross returns"), 
    subtitle = "Aggregated from monthly gross returns", 
    x = "date", 
    y = "gross return"
  )
```








```{r chunk12}
################################################################################
##
##                                   CALCULATIONS
##
################################################################################
```

```{r chunk13}
##                                       FITS
```

```{r chunk14, eval=(run_individual && params$run_fits)}
fits <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), distribution = "sstd")
  })}
)

fits_std <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "std")
  })}
)

fits_norm <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "normal")
  })}
)
```


```{r eval=(run_individual && params$include_long && params$run_fits)}
fits_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), distribution = "sstd")
  })}
)

fits_std_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), method = "Nelder-Mead", distribution = "std")
  })}
)

fits_norm_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), method = "Nelder-Mead", distribution = "normal")
  })}
)
```

```{r chunk15, eval=run_individual && params$run_fits}
saveRDS(fits, file = here("data", data_dir,  "fits.RData"))
saveRDS(fits_std, file = here("data", data_dir, "fits_std.RData"))
saveRDS(fits_norm, file = here("data", data_dir, "fits_norm.RData"))
```

```{r chunk15b, eval=(run_individual && params$include_long && params$run_fits)}
saveRDS(fits_l, file = here("data", data_dir,  "fits_l.RData"))
saveRDS(fits_std_l, file = here("data", data_dir, "fits_std_l.RData"))
saveRDS(fits_norm_l, file = here("data", data_dir, "fits_norm_l.RData"))
```

```{r chunk15c, eval=(run_individual && !params$run_fits)}
fits <- readRDS(file = here("data", data_dir,  "fits.RData"))
fits_std <- readRDS(file = here("data", data_dir, "fits_std.RData"))
fits_norm <- readRDS(file = here("data", data_dir, "fits_norm.RData"))
```

```{r chunk15d, eval=(run_individual && !params$run_fits && params$include_long)}
fits_l <- readRDS(file = here("data", data_dir,  "fits_l.RData"))
fits_std_l <- readRDS(file = here("data", data_dir, "fits_std_l.RData"))
fits_norm_l <- readRDS(file = here("data", data_dir, "fits_norm_l.RData"))
```


```{r eval=run_individual}
fit_summary <- as.data.frame(lapply(
  fits,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary) <- fund_names
rownames(fit_summary) <- c("m", "s", "nu", "xi", "R^2")
```

```{r eval=run_individual}
fit_summary_std <- as.data.frame(lapply(
  fits_std,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_std) <- fund_names
rownames(fit_summary_std) <- c("m", "s", "nu", "R^2")
```

```{r eval=run_individual}
fit_summary_norm <- as.data.frame(lapply(
  fits_norm,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_norm) <- fund_names
rownames(fit_summary_norm) <- c("m", "s", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_l <- as.data.frame(lapply(
  fits_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_l) <- fund_names_l
rownames(fit_summary_l) <- c("m", "s", "nu", "xi", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_std_l <- as.data.frame(lapply(
  fits_std_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_std_l) <- fund_names_l
rownames(fit_summary_std_l) <- c("m", "s", "nu", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_norm_l <- as.data.frame(lapply(
  fits_norm_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_norm_l) <- fund_names_l
rownames(fit_summary_norm_l) <- c("m", "s", "R^2")
```


```{r eval=(params$run_comparison && params$run_sim)}
kappa_df <- data.frame(matrix(numeric(length(fund_names)), ncol = length(fund_names)))

for(i in seq_along(fits)) {
  kappa_df[1, i] <- f_kappa(
    n0 = 1, 
    n = 30, 
    mean = fits[[i]]$dist_params[1], 
    sd = fits[[i]]$dist_params[2], 
    nu = fits[[i]]$dist_params[3], 
    xi = fits[[i]]$dist_params[4], 
    num_sim = 1e5
  )
}

colnames(kappa_df) <- fund_names
```

```{r eval=(params$run_comparison && params$run_sim)}
saveRDS(kappa_df, file = here("data", data_dir,  "kappa_df.RData"))
```

```{r eval=(params$run_comparison && !params$run_sim)}
kappa_df <- readRDS(file = here("data", data_dir,  "kappa_df.RData"))
```

```{r eval=(params$run_comparison && params$run_sim)}
n_min_df <- data.frame(matrix(numeric(length(fund_names)), ncol = length(fund_names)))

for(i in seq_along(fits)) {
  n_min_df[1, i] <- f_n_min(
    n_g = 30, 
    mean = fits[[i]]$dist_params[1], 
    sd = fits[[i]]$dist_params[2], 
    nu = fits[[i]]$dist_params[3], 
    xi = fits[[i]]$dist_params[4], 
    num_sim = 1e5, 
    approx = FALSE
  )
}

colnames(n_min_df) <- fund_names
```

```{r eval=(params$run_comparison && params$run_sim)}
saveRDS(n_min_df, file = here("data", data_dir,  "n_min_df.RData"))
```

```{r eval=(params$run_comparison && !params$run_sim)}
n_min_df <- readRDS(file = here("data", data_dir,  "n_min_df.RData"))
```


```{r chunk16}
##                                     MONTE CARLO
```

```{r chunk17, eval=(params$run_sim && run_individual)}
num_fits <- length(fits)

mc_output <- list()
for(i in seq_along(fits)) {
  message(paste0("\nsstd: ", i, " of ", num_fits, "\n"))
  mc_output[[i]] <- mc_simulation(
    fits[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "sstd"
  )
}

mc_output_std <- list()
for(i in seq_along(fits_std)) {
  message(paste0("\nstd: ", i, " of, ", num_fits, "\n"))
  mc_output_std[[i]] <- mc_simulation(
    fits_std[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "std"
  )
}

mc_output_norm <- list()
for(i in seq_along(fits_norm)) {
  message(paste0("\nnormal: ", i, " of, ", num_fits, "\n"))
  mc_output_norm[[i]] <- mc_simulation(
    fits_norm[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "normal"
  )
}
```

```{r chunk17a, eval=(params$run_sim && run_individual && params$include_long)}
fit_id_l <- list(
  1
)

num_fits_l <- length(fits_l)

mc_output_l <- list()
for(i in seq_along(fits_l)) {
  message(paste0("\nsstd: ", i, " of ", num_fits_l, "\n"))
  mc_output_l[[i]] <- mc_simulation(
    fits_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "sstd"
  )
}

mc_output_std_l <- list()
for(i in seq_along(fits_std_l)) {
  message(paste0("\nstd: ", i, " of, ", num_fits, "\n"))
  mc_output_std_l[[i]] <- mc_simulation(
    fits_std_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "std"
  )
}

mc_output_norm_l <- list()
for(i in seq_along(fits_norm_l)) {
  message(paste0("\nnormal: ", i, " of, ", num_fits, "\n"))
  mc_output_norm_l[[i]] <- mc_simulation(
    fits_norm_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "normal"
  )
}
```


```{r chunk17b, eval=(params$run_sim && run_individual)}
saveRDS(mc_output, file = here("data", data_dir, "mc_output.RData"))
saveRDS(mc_output_std, file = here("data", data_dir, "mc_output_std.RData"))
saveRDS(mc_output_norm, file = here("data", data_dir, "mc_output_norm.RData"))
```

```{r chunk17c, eval=(params$run_sim && run_individual && params$include_long)}
saveRDS(mc_output_l, file = here("data", data_dir, "mc_output_l.RData"))
saveRDS(mc_output_std_l, file = here("data", data_dir, "mc_output_std_l.RData"))
saveRDS(mc_output_norm_l, file = here("data", data_dir, "mc_output_norm_l.RData"))
```

```{r chunk18, eval=(run_individual && !params$run_sim)}
mc_output <- readRDS(file = here("data", data_dir, "mc_output.RData"))
mc_output_std <- readRDS(file = here("data", data_dir, "mc_output_std.RData"))
mc_output_norm <- readRDS(file = here("data", data_dir, "mc_output_norm.RData"))
```

```{r chunk18b, eval=(run_individual && !params$run_sim && params$include_long)}
mc_output_l <- readRDS(file = here("data", data_dir, "mc_output_l.RData"))
mc_output_std_l <- readRDS(file = here("data", data_dir, "mc_output_std_l.RData"))
mc_output_norm_l<- readRDS(file = here("data", data_dir, "mc_output_norm_l.RData"))
```


```{r chunk19}
##                                     MAX SUM PLOTS
```

```{r chunk20, eval=(params$run_sim && run_individual)}
max_sum_plots <- list()
for(i in seq_along(fits)) {
  max_sum_plots[[i]] <- plot_max_sum(
    fits[fit_id[[i]]], 
    mc_num_paths,
    distribution = "sstd"
  )
}

max_sum_plots_std <- list()
for(i in seq_along(fits_std)) {
  max_sum_plots_std[[i]] <- plot_max_sum(
    fits_std[fit_id[[i]]], 
    mc_num_paths, 
    distribution = "std"
  )
}

max_sum_plots_norm <- list()
for(i in seq_along(fits_norm)) {
  max_sum_plots_norm[[i]] <- plot_max_sum(
    fits_norm[fit_id[[i]]], 
    mc_num_paths, 
    distribution = "norm"
  )
}
```

```{r chunk20a, eval=(params$run_sim && run_individual && params$include_long)}
max_sum_plots_l <- list()
for(i in seq_along(fits_l)) {
  max_sum_plots_l[[i]] <- plot_max_sum(
    fits_l[fit_id_l[[i]]], 
    mc_num_paths,
    distribution = "sstd"
  )
}

max_sum_plots_std_l <- list()
for(i in seq_along(fits_std_l)) {
  max_sum_plots_std_l[[i]] <- plot_max_sum(
    fits_std_l[fit_id_l[[i]]], 
    mc_num_paths, 
    distribution = "std"
  )
}

max_sum_plots_norm_l <- list()
for(i in seq_along(fits_norm_l)) {
  max_sum_plots_norm_l[[i]] <- plot_max_sum(
    fits_norm_l[fit_id_l[[i]]], 
    mc_num_paths, 
    distribution = "norm"
  )
}
```

```{r chunk20b, eval=(params$run_sim && run_individual)}
saveRDS(max_sum_plots, here("data", data_dir, "max_sum_plots.RData"))
saveRDS(max_sum_plots_std, here("data", data_dir, "max_sum_plots_std.RData"))
saveRDS(max_sum_plots_norm, here("data", data_dir, "max_sum_plots_norm.RData"))
```

```{r chunk20c, eval=(params$run_sim && run_individual && params$include_long)}
saveRDS(max_sum_plots_l, here("data", data_dir, "max_sum_plots_l.RData"))
saveRDS(max_sum_plots_std_l, here("data", data_dir, "max_sum_plots_std_l.RData"))
saveRDS(max_sum_plots_norm_l, here("data", data_dir, "max_sum_plots_norm_l.RData"))
```



```{r chunk21, eval=(run_individual && !params$run_sim)}
max_sum_plots <- readRDS(file = here("data", data_dir, "max_sum_plots.RData"))
max_sum_plots_std <- readRDS(file = here("data", data_dir, "max_sum_plots_std.RData"))
max_sum_plots_norm <- readRDS(file = here("data", data_dir, "max_sum_plots_norm.RData"))
```

```{r chunk21b, eval=(run_individual && !params$run_sim)}
max_sum_plots_l <- readRDS(file = here("data", data_dir, "max_sum_plots_l.RData"))
max_sum_plots_std_l <- readRDS(file = here("data", data_dir, "max_sum_plots_std_l.RData"))
max_sum_plots_norm_l <- readRDS(file = here("data", data_dir, "max_sum_plots_norm_l.RData"))
```


```{r chunk22}
##                                  IMPORTANCE SAMPLING
```

```{r chunk23, eval=(params$run_sim && run_individual && params$run_is_sim)}
is_g_fits <-  list()
is_output <- list()
for(i in seq_along(fits)) {
  is_g_fits[[i]] <- is_proposal(
    x_i_fit = fits[fit_id[[i]]],
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    obj_func_plot = TRUE,
    init_par = c(2, 0.5),
    method = "L-BFGS-B",
    lower = c(1, 0.1),
    upper = c(3, 0.9)
  )
  
  is_output[[i]] <- importance_sampling(
    x_i_fit = fits[fit_id[[i]]], 
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    g_n_params = is_g_fits[[i]]$par, 
    mode = 1
  )
}
```

```{r chunk23b, eval=(params$run_sim && run_individual && params$run_is_sim)}
saveRDS(is_g_fits, here("data", data_dir, "is_g_fits.RData"))
saveRDS(is_output, here("data", data_dir, "is_output.RData"))
```

```{r chunk24, eval=(run_individual && !params$run_sim && !params$run_is_sim && params$run_is_plot)}
is_g_fits <- readRDS(file = here("data", data_dir, "is_g_fits.RData"))
is_output <- readRDS(file = here("data", data_dir, "is_output.RData"))
```


```{r chunk25}
###########################################################################
##
##                            GENERATE REPORTS
##
###########################################################################
```

```{r chunk26, eval=params$run_exploratory, results='asis'}
exploratory_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-exploratory.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(exploratory_report), sep = '\n')
```

```{r chunk27, eval=params$run_comparison, results='asis'}
comparison_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-comparison.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(comparison_report), sep = '\n')
```

```{r chunk28, eval=params$run_individual, results='asis'}
long <- FALSE
individual_reports <- lapply(
  seq_along(data_df),
  function(i) {
    knitr::knit_child(
      here("child_docs", "pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```

```{r chunk28b, eval=params$run_individual, results='asis', eval=params$include_long}
long <- TRUE
individual_reports <- lapply(
  seq_along(data_df_l),
  function(i) {
    knitr::knit_child(
      here("child_docs", "pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```







# Discussion

## The plans are one portfolio at different leverage

The most consequential finding is structural. The four plans are not four different
stock/bond mixes; they are weighted blends of the same two underlying funds (a low-risk and a
high-risk building block), so the high-risk plan behaves like a *leveraged* version of the
medium-risk plan -- a near-scalar multiple with intercept $\approx 0$, and a single common
factor dominates the joint variation across all four plans. Practically, the choice between
risk profiles within a provider is a choice of **how much** of one portfolio to hold, not
**which** portfolio: higher profiles buy a higher expected growth rate at the price of
proportionally larger swings, with essentially no diversification difference between them.

## What "risk" looks like in the data

Because $\beta>1$ with near-zero intercept, the high-risk plan amplifies the common factor
symmetrically: higher average growth, but deeper drawdowns (worst peak-to-trough $-17.3\%$ vs
$-15.2\%$ for Velliv, $-15.9\%$ vs $-11.9\%$ for PFA) and underperformance over any window in
which the cumulative return since the start is negative. The much-noted property that the
cumulative high-risk path never dips below the medium-risk path is, as the "Path crossing"
section shows, an artefact of starting the index in June 2012 at a low point before a long
bull market -- rebasing to just before the COVID crash, the high-risk paths fall below the
medium-risk paths for about nine months before recovering.

## What the data can and cannot tell us

Monthly resolution helps: 142 observations identify the fits better than the annual report's
13, and the series captures real drawdowns (COVID-2020, 2022) rather than averaging them away
as the annual figures do. But 142 months is still roughly one market cycle, monthly returns
carry serial dependence, and there is no 2008-scale event in the short series. Taleb's
$\kappa$ / $n_{min}$ machinery (appendix) still implies that a Student-$t$ with $\nu\approx 3$
needs far more data than a Gaussian to pin down, so quantitative tail estimates remain
fragile -- even if less so than annually. The long Velliv series (`vmrl`, back to 2007) is the
one place a 2008-type left-tail event enters the estimate, which is the strongest reason to
keep it.

## The mix and the fat-tail reading

The Monte-Carlo results are now produced as a single "black box" fit to each realised series
rather than by combining independent simulations, which removes a spurious diversification
effect that previously made the Velliv+PFA mixes look artificially safe. What remains -- and
what correlation alone misses -- is a genuine *tail* benefit from mixing two fat-tailed
providers: one provider's idiosyncratic crash is softened because the mix always lies between
the two. Quantifying that benefit against its cost is the subject of a separate analysis. On
the distributional side, the fitted $\nu\approx 3$ implies (near-)infinite variance; the
symmetric-$t$ Monte Carlo for the high-risk PFA profile diverges to absurd values, which is
not a numerical error but the fit telling us its variance does not exist.

## Bottom line

Even with monthly data spanning about one cycle, we can say a great deal *qualitatively* --
the plans are leverage on one portfolio, the tails are fat, drawdowns are asymmetric, and
apparent path-dominance is a starting-point artefact -- while being honest that *quantitative*
tail estimates are still fragile at this sample size. The next step is to turn the qualitative
structure into the two decisions an investor actually faces: which risk level, and whether to
split across providers.

# Appendix

## Infinite variance

Taleb, Statistical Consequences Of Fat Tails, p. 97:  
"the variance of a finite variance random variable with tail exponent $< 4$ will be infinite".

And p. 363:  
"The hedging errors for an option portfolio (under a daily revision regime) over 3000 days, under a constant volatility Student T with tail exponent $\alpha = 3$. Technically the errors should not converge in finite time as their distribution has infinite variance."


## QQ lines  

Note: QQ lines by design pass through 1st and 3rd quantiles. They are not trendlines in the sense of linear regression.  

## Arithmetic vs geometric mean
Let $m$ be the number of steps in each path and $n$ be the number of paths.
$a$ is the initial capital.
Use arithmetic mean for mean of all paths at time $t$:
$$\dfrac{a (e^{z_1} + e^{z_2} + \dots + e^{z_n})}{n}$$
where
$$z_i := x_{i, 1} + x_{i, 2} + \dots + x_{i, m}$$
Use geometric mean for mean of all steps in a single path $i$:
$$a e^{\frac{x_{i, 1} + x_{i, 2} + \dots + x_{i, m}}{m}} = a \sqrt[m]{e^{x_{i, 1} + x_{i, 2} + \dots + x_{i, m}}}$$

So for **Monte Carlo** of returns after $m$ periods, we 

+ fit a skewed t-distribution to log-returns and use that distribution to simulate $\{x_{i, j}\}_j^m$,
+ for each path $i$, calculate $100\cdot e^{z_i}$,
+ calculate the mean of $\{z_i\}_i^n$:
    + $$\bar{z} = 100\dfrac{e^{z_1} + e^{z_2} + \dots + e^{z_n}}{n}$$

For **Importance Sampling**, we

+ model log-returns on a skewed t-distribution,
+ for each path $i$, calculate $100\cdot e^{z_i}$,
+ fit a skewed t-distribution to $\{z_i\}_i^n$ and use it as our $f$ density function from which we simulate $\{h_i\}_i^n$,
    + In our case $h$ and $z$ are identical, because we have an idea for a distribution to simulate $z$, but in general for IS $h$ could be a function of $z$.
+ calculate $w* = \frac{f}{g^*}$, where $g*$ is our proposal distribution, which minimizes the variance of $h\cdot w$.
+ calculate the arithmetic mean of $\{h_i w_i^{*}\}_i^n$: 
    + $$100 \dfrac{e^{h_1 w_1^{*}} +  e^{h_2 w_2^{*}} + \dots +  e^{h_n w_n^{*}}}{n}$$


## Average of returns vs returns of average

### Math

$$\text{Avg. of returns} := \dfrac{ \left(\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} \right) }{2}$$
$$\text{Returns of avg.} := \left(\dfrac{ x_t + y_t }{2}\right) \Big/ \left(\dfrac{ x_{t-1} + y_{t-1} }{2}\right) \equiv \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

For which $x_1$ and $y_1$ are
$\text{Avg. of returns} = \text{Returns of avg.}$?

$$\dfrac{ \left(\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} \right) }{2} = \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

$$\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} = 2 \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

$$(x_{t-1} + y_{t-1}) x_t y_{t-1} + (x_{t-1} + y_{t-1}) x_{t-1} y_t = 2 (x_{t-1}y_{t-1}x_t + x_{t-1}y_{t-1}y_t)$$

$$(x_{t-1}x_ty_{t-1} + y_{t-1}x_ty_{t-1}) + (x_{t-1}x_{t-1}y_t + x_{t-1}y_{t-1}y_t) = 2(x_{t-1}y_{t-1}x_t + x_{t-1}y_{t-1}y_t)$$
This is not generally true, but true if for instance
$x_{t-1} = y_{t-1}$.

### Example

```{r}
x0 <- 100
y0 <- 200
Rx <- 0.5
Ry <- 1.5
```

Definition: `R = 1+r`

```{r}
cat(paste0("Let x_0 be ", x0, ".\n"))
cat(paste0("Let y_0 be ", y0, ".\n"))
cat("So the initial value of the pf is", x0 + y0, ".\n")
cat("\n")
cat(paste0("Let R_x be ", Rx, ".\n"))
cat(paste0("Let R_y be ", Ry, ".\n"))
```

Then,

```{r}
cat(paste0("x_1 is R_x * x_0 = ", Rx * x0, ".\n"))
cat(paste0("y_1 is R_y * y_0 = ", Ry * y0, ".\n"))
```

Average of returns:

```{r}
cat("0.5 * (R_x + R_y) =", 0.5 * (Rx + Ry), "\n")
```

So here the value of the pf at t=1 should be unchanged from t=0:

```{r}
cat("(x_0 + y_0) * 0.5 * (R_x + R_y) =", (x0 + y0) * 0.5 * (Rx + Ry), "\n")
```

But this is clearly not the case:

```{r}
cat("0.5 * (x_1 + y_1) = 0.5 * (R_x * x_0 + R_y * y_0) =", 0.5 * (Rx * x0 + Ry * y0), "\n")
```

Therefore we should take returns of average, not average of returns!

Let's take the average of log returns instead:

```{r}
cat("0.5 * (log(R_x) + log(R_y)) =", 0.5 * (log(Rx) + log(Ry)), "\n")
```

We now get:

```{r}
cat("(x_0 + y_0) * exp(0.5 * (log(Rx) + log(Ry))) =", (x0 + y0) * exp(0.5 * (log(Rx) + log(Ry))), "\n")
```

So taking the average of log returns doesn't work either.

## Simulation of mix vs mix of simulations

Test if a simulation of a mix (average) of two returns series has the
same distribution as a mix of two simulated returns series.

```{r}
mc_sim <- function(
    num_runs = 1, 
    num_paths = 1000, 
    num_periods = 20,
    m_a = 0,
    s_a = 0.4,
    m_b = 10,
    s_b = 3
    ) {
  
  data_x <- rnorm(num_periods, m_a, s_a)
  data_y <- rnorm(num_periods, m_b, s_b)
  
  m_data_x <- mean(data_x)
  s_data_x <- sd(data_x)
  m_data_y <- mean(data_y)
  s_data_y <- sd(data_y)
  
  cat("m(data_x):", m_data_x, "\n")
  cat("s(data_x):", s_data_x, "\n")
  cat("m(data_y):", m_data_y, "\n")
  cat("s(data_y):", s_data_y, "\n")
  cat("\n")
  
  m_data_xy <- mean(0.5 * data_x + 0.5 * data_y)
  s_data_xy <- sd(0.5 * data_x + 0.5 * data_y)
  
  cat("m(data_x + data_y):", m_data_xy, "\n")
  cat("s(data_x + data_y):", s_data_xy, "\n")
  cat("\n")
  
  run_sim <- function(num_runs) {
    df <- data.frame(
      m_a = rep(0, num_runs), m_b = rep(0, num_runs), 
      s_a = rep(0, num_runs), s_b = rep(0, num_runs)
    )
    for(j in 1:num_runs) {
      sim_x <- rep(0, num_paths)
      sim_y <- rep(0, num_paths)
      sim_xy <- rep(0, num_paths)
      for(i in 1:num_paths) {
        sim_x[i] <- sum(rnorm(num_periods, m_data_x, s_data_x))
        sim_y[i] <- sum(rnorm(num_periods, m_data_y, s_data_y))
        sim_xy[i] <- sum(rnorm(num_periods, m_data_xy, s_data_xy))
      }

      df$m_a[j] <-  mean(0.5 * sim_x + 0.5 * sim_y)
      df$m_b[j] <-  mean(sim_xy)
      df$s_a[j] <-  sd(0.5 * sim_x + 0.5 * sim_y)
      df$s_b[j] <-  sd(sim_xy)
    }
    df
  }
  
  run_sim(num_runs)
}
```

```{r}
mc_sim_df <- mc_sim(
  num_runs = 10, 
  num_paths = 1000, 
  num_periods = 20,
  m_a = 0,
  s_a = 0.4,
  m_b = 10,
  s_b = 3
)
```

m and s of final state of all paths.\
`_a` is mix of simulated returns.\
`_b` is simulated mixed returns.

```{r}
knitr::kable(mc_sim_df, digits = 3)
```

```{r}
summary(mc_sim_df)
```

`_a` and `_b` are very close to equal.\
We attribute the differences to differences in estimating the
distributions in version a and b.

The final state is independent of the order of the preceding steps:

```{r}
vect1 <- c(rnorm(100))
vect2 <- c(sample(vect1, 100))
vect3 <- c(sample(vect1, 100))
path1 <- c(0, cumsum(vect1))
path2 <- c(0, cumsum(vect2))
path3 <- c(0, cumsum(vect3))
plot(path1, type = "l", col = "blue", 
     ylim = c(
       min(c(path1, path2, path3)), 
       max(c(path1, path2, path3))
     )
)
lines(path2, col = "red")
lines(path3, col = "green")
```

So does the order of the steps in the two processes matter, when mixing
simulated returns?

```{r}
vect1a <- c(rnorm(100, 0.05, 0.06))
vect1b <- c(sample(vect1a, 100))
vect2a <- c(rnorm(100, 0.05, 0.06))
vect2b <- c(sample(vect2a, 100))

path1a <- 100 * c(1, cumprod(1 + vect1a))
path1b <- 100 * c(1, cumprod(1 + vect1b))
path2a <- 100 * c(1, cumprod(1 + vect2a))
path2b <- 100 * c(1, cumprod(1 + vect2b))

mix_path_a <- 0.5 * path1a + 0.5 * path2a
mix_path_b <- 0.5 * path1b + 0.5 * path2b

plot(path1a, type = "l", lty = 1, col = "blue", 
     ylim = c(
       min(c(path1a, path1b, path2a, path2b)), 
       max(c(path1a, path1b, path2a, path2b))
     )
)
lines(path1b, lty = 2, col = "blue")
lines(path2a, lty = 1, col = "red")
lines(path2b, lty = 2, col = "red")
```

```{r}
plot(mix_path_a, type = "l", lty = 1, col = "blue", 
     ylim = c(
       min(c(mix_path_a, mix_path_b)), 
       max(c(mix_path_a, mix_path_b))
     )
)
lines(mix_path_b, lty = 2, col = "blue")
```

The order of steps in the individual paths do not matter, because the
mix of simulated paths is a sum of a sum, so the order of terms doesn't
affect the sum. If there is variation it is because the sets preceding
steps are not the same. For instance, the steps between step 1 and 60 in
the plot above are not the same for the two lines.

Recall,
$$\text{Var}(aX+bY) = a^2 \text{Var}(X) + b^2 \text{Var}(Y) + 2ab \text{Cov}(a, b)$$

```{r include=TRUE, echo=TRUE}
var(0.5 * data_df$vhr + 0.5 * data_df$phr)
0.5^2 * var(data_df$vhr) + 0.5^2 * var(data_df$phr) + 2 * 0.5 * 0.5 * cov(data_df$vhr, data_df$phr)
```

Our distribution estimate is based on 142 monthly observations. Is that
enough for a robust estimate? What if we suddenly hit a month like the
2008 crash? How would that affect our estimate?\
Let's try to include the Velliv data going back to 2007, which spans the GFC.\
We do this by sampling 142 observations from the long series `vmrl`
(`data_df_l$vmrl`, 207 monthly observations incl. 2007-2012).

```{r}
n <- 50
test_df <- data.frame(m = rep(0, n), s = rep(0, n))
for(i in 1:n) {
  vmrl_smp <- sample(data_df_l$vmrl, 142)
  test_df$m[i] <- mean(0.5 * vmrl_smp + 0.5 * data_df$phr)
  test_df$s[i] <- sd(0.5 * vmrl_smp + 0.5 * data_df$phr)
}
summary(test_df)
```

## The meaning of `xi`

The fit for `mhr` has the highest `xi` value of all. This suggests
right-skew:

```{r}
plot((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 1), xlim = c(-20, 20), ylim = c(0, 0.1), type = "l", xlab = "value", ylab = "likelihood", main = "Skew t-distribution density")
lines((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 0.5), col="red")
lines((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 2), col="blue")
legend("topright", c("xi=1", "xi=0.5", "xi=2"), col = c("black", "red", "blue"), lty = 1)
```

## Max vs sum plot

If the Law Of Large Numbers holds true,
$$\dfrac{\max (X_1^p, ..., X^p)}{\sum_{i=1}^n X_i^p} \rightarrow 0$$ for
$n \rightarrow \infty$.

If not, $X$ doesn't have a $p$'th moment.

See Taleb: The Statistical Consequences Of Fat Tails, p. 192


