---
title: "Annual pension returns analysis"
author: Martin Hoshi Vognsen
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
   - \usepackage[fontsize=8pt]{scrextend}
mainfont: SourceSansPro
output: 
  html_document:
    toc: true
    toc_depth: 3
    keep_md: yes
  pdf_document:
    toc: true
    toc_depth: 3
#fontsize: 10pt # for pdf. Limited to 10pt, 11pt and 12pt. Else use scrextend.
params:
  run_sim: FALSE ## TRUE: Run simulations and write output. FALSE: Read saved
                ## simulations output from disk instead of running the simulations.
  run_fits: FALSE
  run_exploratory: TRUE ## Include exploratory report?
  run_individual: TRUE ## Include individual reports? Depends on run_individual.
  run_comparison: TRUE ## Include comparison report? Depends on run_individual.
  run_comments: FALSE ## !!! OBS! Comments in separarte child doc deprecated
  run_appendix: FALSE ## !!! OBS! Appendix in separarte child doc deprecated
  run_mc_plot: TRUE
  run_is_sim: TRUE
  run_is_plot: TRUE
  include_long: TRUE
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---

```{r chunk1, include=FALSE}
## ****************************** WARNING ******************************
## 
##             This document takes a long time to render!
##
## *********************************************************************
##
## To read previous simulation outputs from disk rather than run simulations,
## set YAML parameter run_sim: TRUE
##
## NOTE:
## When using the "Run All" command, YAML parameters will be ignored. An error
## message "Error in eval(ele) : object 'params' not found" will be printed,
## but the chunks will be evaluated.
```



```{r  chunk2, include=FALSE, eval=FALSE}
################################################################################
##
##                                      SETUP
##
################################################################################
```

```{r chunk3, setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

run_individual <- (params$run_individual + params$run_comparison + params$run_appendix + params$run_mc_plot > 0)
run_exploratory <- (params$run_exploratory + run_individual > 0)

## Note: In the second line (run_exploratory <- ...), run_individual is correct that - not params$run_individual, because if we choose params$run_exploratory=TRUE and params$run_comparison=TRUE, but params$run_individual=FALSE, we still have to do the individual calculations.
```

```{r chunk4}
library(here)
library(tidyr)   # gather() used below; attach explicitly for non-interactive render
```

```{r chunk5}
source(here("src", "pension-returns_functions.R"))
```

```{r chunk6}
################################################################################
##
##                          GENERATE HTML AND PDF
##                        (DON'T USE KNIT BUTTON!)
##
################################################################################
```

```{r chunk6a, eval=FALSE, include=FALSE}
## Render this document to html and pdf
rmarkdown::render(here::here("pension-returns_annual.Rmd"), output_format ="all")
```


```{r chunk6b}
################################################################################
##
##              GENERATE PDF MANUALLY AFTER GENERATING HTML AND MD
##
################################################################################

## Note: Will not produce TOC and a few other things will be different.
```

```{bash chunk7, eval=FALSE, include=FALSE}
## Convert md to pdf
## Run manually:
pandoc pension-returns_annual.md -o pension-returns_annual.pdf
```


```{r chunk7b}
################################################################################
##
##                             COPY TO SITES FOLDER
##
################################################################################
```

```{bash eval=FALSE, include=FALSE}
## Run manually:
cp ~/R\ work/pension-returns/pension-returns_annual.html ~/var/site/pension-returns_annual.html
cp ~/R\ work/pension-returns/pension-returns_annual.pdf ~/var/site/pension-returns_annual.pdf
```


```{r chunk8}
################################################################################
##
##                                      INPUTS
##
################################################################################
```

```{r chunk9}
mc_num_paths <- 1e4
mc_num_periods <- 20
mc_dao <- TRUE
#mc_threshold <- 0.01

## Min saldo for Velliv: kr 12.300
## Assume initial capital of 400.000.
## 100 * 12300/400000 = 3
mc_threshold <- 3
```

# Summary

This is a **case study on a deliberately frozen data set** (annual returns 2011-2023),
not an up-to-date monitor. Fixing the data is what lets us interpret a specific set of
numbers and graphs. Three findings stand out.

1. **The "plans" are leverage levels on one portfolio, not distinct asset mixes.** Within
each provider the high-risk plan is almost exactly a scalar multiple of the medium-risk
plan ($\beta \approx 1.2$ for Velliv, $\approx 1.6$ for PFA, intercept $\approx 0$), and a
single common factor explains about 96% of the variation across all four plans. PFA's own
product description confirms it: profiles B and D are fixed weighted blends of the *same
two* underlying funds.

2. **"Risk" here means deeper drawdowns and start-date-dependent path ordering, not a
different kind of bet.** The high-risk plan amplifies the same factor in both directions:
it loses more in bad years (it underperformed the medium plan in 3-4 of 13 years), and its
cumulative path drops below the medium plan's whenever the index is started just before a
drawdown.

3. **The sample is far too short to estimate the tail, and -- at annual resolution -- it
contains no crash**, so every risk number is both barely identified and optimistically
biased. For *precise* tail estimates the honest verdict is close to Taleb's "better stay
home"; what survives is the *qualitative* structure -- fat tails, leverage, and asymmetric
drawdowns.

The Discussion at the end develops these points.

# Returns data 2011-2023.

Fit log returns to F-S skew standardized Student-t distribution.
`m`  is the location parameter.
`s` is the scale parameter.
`nu` is the estimated shape parameter (degrees of freedom).
`xi` is the estimated skewness parameter.

For 2011, medium risk data is used in the high risk data set, as no high risk fund data is available prior to 2012.  
`vmrl` is a long version of Velliv medium risk data, from 2007 to 2023. For 2007 to 2011 (both included) no high risk data is available.

PFA medium risk is risk profile B.  
PFA high risk is risk profile D.  

```{r chunk10}
## Headlines for individual sections
headlines <- c(
  "Velliv medium risk (vmr), 2011 - 2023",
  "Velliv high risk (vhr), 2011 - 2023",
  "PFA medium risk (pmr), 2011 - 2023",
  "PFA high risk (phr), 2011 - 2023",
  "Mix medium risk (mmr), 2011 - 2023",
  "Mix high risk (mhr), 2011 - 2023",
  "Mix vmr+phr (vm_ph), 2011 - 2023",
  "Mix vhr+pmr (mh_pm), 2011 - 2023"
)

## Log returns data 2011-2023.
## Velliv medium risk returns
vmr <- log(c(0.986, 1.098, 1.157, 1.136, 1.058, 1.044, 1.100, 0.963, 1.168, 1.097, 1.141, 0.868, 1.094))

## Velliv medium risk return, long period: 2007-2023
vmrl <- log(c(1.058, 0.801, 1.193, 1.129, 1.035, 0.996, 1.133, 1.092, 1.085, 1.013, 1.095, 1.011, 1.129, 1.101, 1.128, 0.988, 1.048))

## Velliv high risk returns
vhr <- log(c(0.986, 1.093, 1.214, 1.160, 1.075, 1.039, 1.118, 0.952, 1.205, 1.099, 1.194, 0.849, 1.122))

## PFA medium risk returns
pmr <- log(c(1.004, 1.141, 1.111, 1.107, 1.084, 1.073, 1.089, 0.976, 1.119, 1.042, 1.107, 0.904, 1.084))

## PFA high risk return
phr <- log(c(0.934, 1.182, 1.208, 1.128, 1.123, 1.082, 1.135, 0.943, 1.190, 1.068, 1.200, 0.878, 1.159))


## Mix Velliv medium risk + PFA high risk
# vmr_phr <- (c(100, 100 * cumprod(exp(vmr))) + c(100, 100 * cumprod(exp(phr))))/2
# vmr_phr <- log(tail(vmr_phr, -1)/head(vmr_phr, -1))

## Mix Velliv high risk + PFA medium risk
# vhr_pmr <- (c(100, 100 * cumprod(exp(vhr))) + c(100, 100 * cumprod(exp(pmr))))/2
# vhr_pmr <- log(tail(vhr_pmr, -1)/head(vhr_pmr, -1))

data_df <- data.frame(
  vmr = vmr,
  vhr = vhr,
  pmr = pmr,
  phr = phr
)

data_df <- data_df %>% 
  mutate(
    mmr = mix_of_logreturns(vmr, pmr),
    mhr = mix_of_logreturns(vhr, phr),
    vmr_phr = mix_of_logreturns(vmr, phr),
    vhr_pmr = mix_of_logreturns(vhr, pmr)
  )


data_df_l <- data.frame(
  vmrl = vmrl
)

fund_names <- colnames(data_df)
fund_names_l <- colnames(data_df_l)   # long (vmrl) series; was relied on from global env
```

```{r}
## The specified folder must exist in the `data` directory
data_dir <- "annual"
```


```{r chunk11, eval=FALSE}
all_data <- exp(c(data_df$vmr, data_df$vhr, data_df$pmr, data_df$phr, data_df$mmr, data_df$mhr))
plot(x = 2011:2023, y = exp(data_df$vmr), type = "l",  lty = 1, lwd = 2, xlab = "year", ylab = "Gross returns", col = "#76EEC6", ylim = c(min(all_data), max(all_data)), main = "Gross returns 2011-2023")
lines(x = 2011:2023, y = exp(data_df$vhr), col = "#76EEC6",  lty = 2, lwd = 1.5)
lines(x = 2011:2023, y = exp(data_df$pmr), col = "#CD6600",  lty = 1, lwd = 2)
lines(x = 2011:2023, y = exp(data_df$phr), col = "#CD6600",  lty = 2, lwd = 1.5)
lines(x = 2011:2023, y = exp(data_df$mmr), col = "#FF00FF",  lty = 1, lwd = 2)
lines(x = 2011:2023, y = exp(data_df$mhr), col = "#FF00FF",  lty = 2, lwd = 1.5)
legend("bottom", legend = c("vmr", "vhr", "pmr", "phr", "mmr", "mhr"), col = c("#76EEC6", "#76EEC6", "#CD6600", "#CD6600", "#FF00FF", "#FF00FF"), lty = c(1, 2, 1, 2, 1, 2), lwd = 2, ncol = 3 )
```

```{r}
ggplot(
  data_df[, 1:6] %>% 
    mutate(date = 2011:2023) %>% 
    gather(key = "fund", value = "gross_return", -date),
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund)) +
  scale_x_continuous(breaks = 2011:2023) +
  labs(
    title = paste("Gross returns 2011-2023"),
    subtitle = "",
    x = "date",
    y = "gross return"
  )
```

## Path crossing and the starting point

Plotting the *cumulative* value of each plan from a common starting index makes the
riskier plans appear to dominate -- their paths sit above the less-risky ones and rarely
fall below. It is tempting to read this as *more risk always pays*. It does not
generalise: it depends on **where the index starts**.

Each high-risk plan is, in effect, a leveraged exposure to one common return factor: in
cumulative log terms $\log(W_\text{high}/100) \approx \beta\,\log(W_\text{med}/100)$ with
slope $\beta > 1$ and intercept $\approx 0$. With $\beta > 1$ this gives
$W_\text{high} > W_\text{med}$ **exactly when** $W_\text{med} > 100$, so the high-risk plan
can only fall below the medium-risk plan over a window in which the cumulative return
*since the start date* is negative.

Two realised illustrations. First, no rebasing is even needed for PFA: started in 2011,
the high-risk plan (`phr`, $-6.6\%$ in 2011) begins **below** the medium-risk plan (`pmr`,
$+0.4\%$) and only overtakes it in 2013. Second, rebasing every plan to 100 at the end of
2021 -- just before the 2022 drawdown -- the high-risk plans fall below the medium-risk
plans in 2022 and recross above only in 2023:

```{r fig.width=10}
## Annual cumulative index, rebased to 100 just before the 2022 drawdown.
.yrs <- 2011:2023
.rebase_year <- 2021
.sel <- .yrs > .rebase_year                            # 2022, 2023
.idx <- function(lr) 100 * cumprod(c(1, exp(lr[.sel]))) # 100 at 2021, then 2022, 2023
crossing_df <- data.frame(
  year = c(.rebase_year, .yrs[.sel]),
  `Velliv medium` = .idx(data_df$vmr),
  `Velliv high`   = .idx(data_df$vhr),
  `PFA medium`    = .idx(data_df$pmr),
  `PFA high`      = .idx(data_df$phr),
  check.names = FALSE
) %>% gather(key = "plan", value = "price", -year)

ggplot(crossing_df, aes(x = year, y = price, color = plan)) +
  geom_line(linewidth = 0.6) +
  geom_point() +
  geom_hline(yintercept = 100, linetype = 2, colour = "grey50") +
  scale_x_continuous(breaks = c(.rebase_year, .yrs[.sel])) +
  labs(
    title = "Annual index started just before the 2022 drawdown (end-2021 = 100)",
    subtitle = "High-risk plans fall below medium-risk in 2022, then recross above by 2023",
    x = "year", y = "price index"
  )
```

Whether the realised paths cross is therefore a property of the starting point relative
to the market cycle, not of the plans themselves. A deeper or longer regime (e.g. a
2008-2009-type drawdown, present in the long Velliv series `vmrl`) would hold the
high-risk plan below for correspondingly longer.




```{r chunk12}
################################################################################
##
##                                   CALCULATIONS
##
################################################################################
```

```{r chunk13}
##                                       FITS
```

```{r chunk14, eval=(params$run_fits && run_individual)}
fits <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "sstd")
  })}
)

fits_std <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "std")
  })}
)

fits_norm <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "normal")
  })}
)
```

```{r chunk15, eval=(params$run_fits && run_individual)}
saveRDS(fits, file = here("data", data_dir, "fits.RData"))
saveRDS(fits_std, file = here("data", data_dir, "fits_std.RData"))
saveRDS(fits_norm, file = here("data", data_dir, "fits_norm.RData"))
```

```{r chunk15c, eval=(run_individual && !params$run_fits)}
fits <- readRDS(file = here("data", data_dir,  "fits.RData"))
fits_std <- readRDS(file = here("data", data_dir, "fits_std.RData"))
fits_norm <- readRDS(file = here("data", data_dir, "fits_norm.RData"))
```

```{r eval=(params$run_fits && run_individual && params$include_long)}
fits_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), distribution = "sstd")
  })}
)

fits_std_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), method = "Nelder-Mead", distribution = "std")
  })}
)

fits_norm_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), method = "Nelder-Mead", distribution = "normal")
  })}
)
```

```{r chunk15b, eval=(params$run_fits && run_individual && params$include_long)}
saveRDS(fits_l, file = here("data", data_dir,  "fits_l.RData"))
saveRDS(fits_std_l, file = here("data", data_dir, "fits_std_l.RData"))
saveRDS(fits_norm_l, file = here("data", data_dir, "fits_norm_l.RData"))
```

```{r chunk15d, eval=(run_individual && !params$run_fits && params$include_long)}
fits_l <- readRDS(file = here("data", data_dir,  "fits_l.RData"))
fits_std_l <- readRDS(file = here("data", data_dir, "fits_std_l.RData"))
fits_norm_l <- readRDS(file = here("data", data_dir, "fits_norm_l.RData"))
```


```{r eval=run_individual}
fit_summary <- as.data.frame(lapply(
  fits,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary) <- fund_names
rownames(fit_summary) <- c("m", "s", "nu", "xi", "R^2")
```

```{r eval=run_individual}
fit_summary_std <- as.data.frame(lapply(
  fits_std,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_std) <- fund_names
rownames(fit_summary_std) <- c("m", "s", "nu", "R^2")
```

```{r eval=run_individual}
fit_summary_norm <- as.data.frame(lapply(
  fits_norm,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_norm) <- fund_names
rownames(fit_summary_norm) <- c("m", "s", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_l <- as.data.frame(lapply(
  fits_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_l) <- fund_names_l
rownames(fit_summary_l) <- c("m", "s", "nu", "xi", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_std_l <- as.data.frame(lapply(
  fits_std_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_std_l) <- fund_names_l
rownames(fit_summary_std_l) <- c("m", "s", "nu", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_norm_l <- as.data.frame(lapply(
  fits_norm_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_norm_l) <- fund_names_l
rownames(fit_summary_norm_l) <- c("m", "s", "R^2")
```

```{r eval=(params$run_comparison && params$run_sim)}
kappa_df <- data.frame(matrix(numeric(length(fund_names)), ncol = length(fund_names)))

for(i in seq_along(fits)) {
  kappa_df[1, i] <- f_kappa(
    n0 = 1, 
    n = 30, 
    mean = fits[[i]]$dist_params[1], 
    sd = fits[[i]]$dist_params[2], 
    nu = fits[[i]]$dist_params[3], 
    xi = fits[[i]]$dist_params[4], 
    num_sim = 1e5
  )
}

colnames(kappa_df) <- fund_names
```

```{r eval=(params$run_comparison && params$run_sim)}
saveRDS(kappa_df, file = here("data", data_dir,  "kappa_df.RData"))
```

```{r eval=(params$run_comparison && !params$run_sim)}
kappa_df <- readRDS(file = here("data", data_dir,  "kappa_df.RData"))
```

```{r eval=(params$run_comparison && params$run_sim)}
n_min_df <- data.frame(matrix(numeric(length(fund_names)), ncol = length(fund_names)))

for(i in seq_along(fits)) {
  n_min_df[1, i] <- f_n_min(
    n_g = 30, 
    mean = fits[[i]]$dist_params[1], 
    sd = fits[[i]]$dist_params[2], 
    nu = fits[[i]]$dist_params[3], 
    xi = fits[[i]]$dist_params[4], 
    num_sim = 1e5, 
    approx = FALSE
  )
}

colnames(n_min_df) <- fund_names
```

```{r eval=(params$run_comparison && params$run_sim)}
saveRDS(n_min_df, file = here("data", data_dir,  "n_min_df.RData"))
```

```{r eval=(params$run_comparison && !params$run_sim)}
n_min_df <- readRDS(file = here("data", data_dir,  "n_min_df.RData"))
```



```{r chunk16}
##                                     MONTE CARLO
```

```{r chunk17, eval=(params$run_sim && run_individual)}
## PURE BLACK BOX: each fund (mixes included) is simulated from its OWN single fit
## to its observed total-value series -- i.e. "simulation of the average", NOT
## "average of simulations". The former component-recombination (e.g. c(1, 3) for a
## mix) drew the two components independently, assuming zero correlation between
## near-collinear funds, and is abandoned. fits[[5..8]] are the single fits to the
## realised mix series (mmr, mhr, vmr_phr, vhr_pmr) built via mix_of_logreturns().
fit_id <- list(1, 2, 3, 4, 5, 6, 7, 8)

num_fits <- length(fits)

mc_output <- list()
for(i in seq_along(fits)) {
  message(paste0("\nsstd: ", i, " of ", num_fits, "\n"))
  mc_output[[i]] <- mc_simulation(
    fits[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "sstd"
  )
}

mc_output_std <- list()
for(i in seq_along(fits_std)) {
  message(paste0("\nstd: ", i, " of, ", num_fits, "\n"))
  mc_output_std[[i]] <- mc_simulation(
    fits_std[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "std"
  )
}

mc_output_norm <- list()
for(i in seq_along(fits_norm)) {
  message(paste0("\nnormal: ", i, " of, ", num_fits, "\n"))
  mc_output_norm[[i]] <- mc_simulation(
    fits_norm[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "normal"
  )
}
```


```{r chunk17a, eval=(params$run_sim && run_individual && params$include_long)}
## Long-series MC (single fund vmrl). Ported from the monthly report so the annual
## report is self-contained for the long series (it previously relied on these
## objects lingering in a shared session).
fit_id_l <- list(
  1
)

num_fits_l <- length(fits_l)

mc_output_l <- list()
for(i in seq_along(fits_l)) {
  mc_output_l[[i]] <- mc_simulation(
    fits_l[fit_id_l[[i]]],
    num_paths = mc_num_paths,
    num_periods = mc_num_periods,
    dao = mc_dao,
    threshold = mc_threshold,
    distribution = "sstd"
  )
}

mc_output_std_l <- list()
for(i in seq_along(fits_std_l)) {
  mc_output_std_l[[i]] <- mc_simulation(
    fits_std_l[fit_id_l[[i]]],
    num_paths = mc_num_paths,
    num_periods = mc_num_periods,
    dao = mc_dao,
    threshold = mc_threshold,
    distribution = "std"
  )
}

mc_output_norm_l <- list()
for(i in seq_along(fits_norm_l)) {
  mc_output_norm_l[[i]] <- mc_simulation(
    fits_norm_l[fit_id_l[[i]]],
    num_paths = mc_num_paths,
    num_periods = mc_num_periods,
    dao = mc_dao,
    threshold = mc_threshold,
    distribution = "normal"
  )
}
```

```{r chunk17b, eval=(params$run_sim && run_individual)}
saveRDS(mc_output, file = here("data", data_dir, "mc_output.RData"))
saveRDS(mc_output_std, file = here("data", data_dir, "mc_output_std.RData"))
saveRDS(mc_output_norm, file = here("data", data_dir, "mc_output_norm.RData"))
```

```{r chunk17c, eval=(params$run_sim && run_individual && params$include_long)}
saveRDS(mc_output_l, file = here("data", data_dir, "mc_output_l.RData"))
saveRDS(mc_output_std_l, file = here("data", data_dir, "mc_output_std_l.RData"))
saveRDS(mc_output_norm_l, file = here("data", data_dir, "mc_output_norm_l.RData"))
```

```{r chunk18, eval=(run_individual && !params$run_sim)}
mc_output <- readRDS(file = here("data", data_dir, "mc_output.RData"))
mc_output_std <- readRDS(file = here("data", data_dir, "mc_output_std.RData"))
mc_output_norm <- readRDS(file = here("data", data_dir, "mc_output_norm.RData"))
```

```{r chunk18b, eval=(run_individual && !params$run_sim && params$include_long)}
mc_output_l <- readRDS(file = here("data", data_dir, "mc_output_l.RData"))
mc_output_std_l <- readRDS(file = here("data", data_dir, "mc_output_std_l.RData"))
mc_output_norm_l <- readRDS(file = here("data", data_dir, "mc_output_norm_l.RData"))
```

```{r chunk19}
##                                     MAX SUM PLOTS
```

```{r chunk20, eval=(params$run_sim && run_individual)}
max_sum_plots <- list()
for(i in seq_along(fits)) {
  max_sum_plots[[i]] <- plot_max_sum(
    fits[fit_id[[i]]], 
    mc_num_paths,
    distribution = "sstd"
  )
}

max_sum_plots_std <- list()
for(i in seq_along(fits_std)) {
  max_sum_plots_std[[i]] <- plot_max_sum(
    fits_std[fit_id[[i]]], 
    mc_num_paths, 
    distribution = "std"
  )
}

max_sum_plots_norm <- list()
for(i in seq_along(fits_norm)) {
  max_sum_plots_norm[[i]] <- plot_max_sum(
    fits_norm[fit_id[[i]]], 
    mc_num_paths, 
    distribution = "norm"
  )
}
```

```{r chunk20a, eval=(params$run_sim && run_individual && params$include_long)}
max_sum_plots_l <- list()
for(i in seq_along(fits_l)) {
  max_sum_plots_l[[i]] <- plot_max_sum(
    fits_l[fit_id_l[[i]]],
    mc_num_paths,
    distribution = "sstd"
  )
}

max_sum_plots_std_l <- list()
for(i in seq_along(fits_std_l)) {
  max_sum_plots_std_l[[i]] <- plot_max_sum(
    fits_std_l[fit_id_l[[i]]],
    mc_num_paths,
    distribution = "std"
  )
}

max_sum_plots_norm_l <- list()
for(i in seq_along(fits_norm_l)) {
  max_sum_plots_norm_l[[i]] <- plot_max_sum(
    fits_norm_l[fit_id_l[[i]]],
    mc_num_paths,
    distribution = "norm"
  )
}
```

```{r chunk20b, eval=(params$run_sim && run_individual)}
saveRDS(max_sum_plots, here("data", data_dir, "max_sum_plots.RData"))
saveRDS(max_sum_plots_std, here("data", data_dir, "max_sum_plots_std.RData"))
saveRDS(max_sum_plots_norm, here("data", data_dir, "max_sum_plots_norm.RData"))
```

```{r chunk20c, eval=(params$run_sim && run_individual && params$include_long)}
saveRDS(max_sum_plots_l, here("data", data_dir, "max_sum_plots_l.RData"))
saveRDS(max_sum_plots_std_l, here("data", data_dir, "max_sum_plots_std_l.RData"))
saveRDS(max_sum_plots_norm_l, here("data", data_dir, "max_sum_plots_norm_l.RData"))
```



```{r chunk21, eval=(run_individual && !params$run_sim)}
max_sum_plots <- readRDS(file = here("data", data_dir, "max_sum_plots.RData"))
max_sum_plots_std <- readRDS(file = here("data", data_dir, "max_sum_plots_std.RData"))
max_sum_plots_norm <- readRDS(file = here("data", data_dir, "max_sum_plots_norm.RData"))
```

```{r chunk21b, eval=(run_individual && !params$run_sim && params$include_long)}
max_sum_plots_l <- readRDS(file = here("data", data_dir, "max_sum_plots_l.RData"))
max_sum_plots_std_l <- readRDS(file = here("data", data_dir, "max_sum_plots_std_l.RData"))
max_sum_plots_norm_l <- readRDS(file = here("data", data_dir, "max_sum_plots_norm_l.RData"))
```

```{r chunk22}
##                                  IMPORTANCE SAMPLING
```

```{r chunk23, eval=(params$run_sim && run_individual && params$run_is_sim)}
is_g_fits <-  list()
is_output <- list()
for(i in seq_along(fits)) {
  is_g_fits[[i]] <- is_proposal(
    x_i_fit = fits[fit_id[[i]]],
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    obj_func_plot = TRUE,
    init_par = c(2, 0.5),
    method = "L-BFGS-B",
    lower = c(1, 0.1),
    upper = c(3, 0.9)
  )
  
  is_output[[i]] <- importance_sampling(
    x_i_fit = fits[fit_id[[i]]], 
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    g_n_params = is_g_fits[[i]]$par, 
    mode = 1
  )
}
```

```{r chunk23b, eval=(params$run_sim && run_individual && params$run_is_sim)}
saveRDS(is_g_fits, here("data", data_dir, "is_g_fits.RData"))
saveRDS(is_output, here("data", data_dir, "is_output.RData"))
```

```{r chunk24, eval=(run_individual && !params$run_sim && !params$run_is_sim && params$run_is_plot)}
is_g_fits <- readRDS(file = here("data", data_dir, "is_g_fits.RData"))
is_output <- readRDS(file = here("data", data_dir, "is_output.RData"))
```


```{r chunk25}
###########################################################################
##
##                            GENERATE REPORTS
##
###########################################################################
```

```{r chunk26, eval=params$run_exploratory, results='asis'}
exploratory_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-exploratory.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(exploratory_report), sep = '\n')
```

```{r chunk27, eval=params$run_comparison, results='asis'}
comparison_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-comparison.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(comparison_report), sep = '\n')
```

```{r chunk28, eval=params$run_individual, results='asis'}
long <- FALSE
individual_reports <- lapply(
  seq_along(data_df),
  function(i) {
    knitr::knit_child(
      here("child_docs", "pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```

```{r chunk28b, eval=params$run_individual, results='asis', eval=params$include_long}
long <- TRUE
individual_reports <- lapply(
  seq_along(data_df_l),
  function(i) {
    knitr::knit_child(
      here("child_docs", "pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```


```{r chunk29, eval=params$run_comments, results='asis'}
comments_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-comments.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(comments_report), sep = '\n')
```





```{r chunk30, eval=params$run_appendix, results='asis'}
appendix_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-appendix.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(appendix_report), sep = '\n')
```



# Discussion

With the data frozen we can step back and ask what these numbers and graphs actually tell
us.

## The plans are one portfolio at different leverage

The most consequential finding is structural. The four plans are not four different
stock/bond mixes; they are weighted blends of the same two underlying funds (a low-risk and
a high-risk building block), so the high-risk plan behaves like a *leveraged* version of the
medium-risk plan. Empirically the high-risk series is a near-scalar multiple of the
medium-risk series ($R^2$ of 0.91-0.98 with intercept $\approx 0$), and the first principal
component explains about 96% of the joint variation across all four plans. Practically, the
choice between risk profiles within a provider is a choice of **how much** of one portfolio
to hold, not **which** portfolio -- higher profiles buy a higher expected growth rate at the
price of proportionally larger swings, with essentially no diversification difference between
them.

## What "risk" looks like in the data

Because $\beta>1$ with near-zero intercept, the high-risk plan amplifies the common factor
symmetrically. Over 2011-2023 that meant higher average returns (Velliv 8.5% vs 7.0%, PFA
9.5% vs 6.5%) but deeper drawdowns (worst year $-15.1\%$ vs $-13.2\%$ for Velliv, $-12.2\%$
vs $-9.6\%$ for PFA) and outright underperformance in flat or down years. The much-noted
property that the cumulative high-risk path never dips below the medium-risk path is, as the
"Path crossing" section shows, an artefact of starting the index in 2012 at a low point
before a long bull market -- not evidence that more risk is free.

## What the data can and cannot tell us

This is the heart of the case study. The annual fits rest on **13 observations** for up to
four parameters. The skewness parameter `xi` is driven to its boundary in every fit -- a
direct symptom that the shape simply is not identifiable from so few points. Taleb's
$\kappa$ / $n_{min}$ machinery (appendix) quantifies the gap: matching the convergence of 30
Gaussian observations would take far more than 30 observations from a Student-$t$ with
$\nu\approx 3$. So for *point estimates of tail risk*, the honest answer is close to "stay
home."

Two things sharpen this. First, at annual resolution the 2011-2023 window contains **no
crash** -- even 2020 nets to $+9.7\%$ because the COVID drawdown recovered within the year.
The realised maximum gains actually exceed the maximum losses in this sample, although the
fitted distributions are left-skewed: the benign window has not yet delivered the large
drawdown the fit predicts. Second, the long Velliv series (`vmrl`, back to 2007) *does*
contain the 2008 crash, and is the one place in the data where a real left-tail event enters
the estimate. That is the strongest reason to keep it: it is our only empirical anchor for
the very tail the whole study is about.

## The mix and the fat-tail reading

The Monte-Carlo results are now produced as a single "black box" fit to each realised series
rather than by combining independent simulations, which removes a spurious diversification
effect that previously made the Velliv+PFA mixes look artificially safe. What remains -- and
what correlation alone misses -- is a genuine *tail* benefit from mixing two fat-tailed
providers: one provider's idiosyncratic crash is softened because the mix always lies between
the two. Quantifying that benefit against its cost is the subject of a separate analysis. On
the distributional side, the fitted $\nu\approx 3$ implies (near-)infinite variance; the
symmetric-$t$ Monte Carlo for the high-risk PFA profile diverges to absurd values, which is
not a numerical error but the fit telling us its variance does not exist.

## Bottom line

For a fixed, crash-free, 13-point sample we can say a great deal *qualitatively* -- the plans
are leverage on one portfolio, the tails are fat, drawdowns are asymmetric, and apparent
path-dominance is a starting-point artefact -- while being honest that *quantitative* tail
estimates are not trustworthy at this sample size. The next step is to turn the qualitative
structure into the two decisions an investor actually faces: which risk level, and whether to
split across providers.

# Appendix

## Infinite variance

Taleb, Statistical Consequences Of Fat Tails, p. 97:  
"the variance of a finite variance random variable with tail exponent $< 4$ will be infinite".

And p. 363:  
"The hedging errors for an option portfolio (under a daily revision regime) over 3000 days, under a constant volatility Student T with tail exponent $\alpha = 3$. Technically the errors should not converge in finite time as their distribution has infinite variance."


## QQ lines  

Note: QQ lines by design pass through 1st and 3rd quantiles. They are not trendlines in the sense of linear regression.  

## Arithmetic vs geometric mean
Let $m$ be the number of steps in each path and $n$ be the number of paths.
$a$ is the initial capital.
Use arithmetic mean for mean of all paths at time $t$:
$$\dfrac{a (e^{z_1} + e^{z_2} + \dots + e^{z_n})}{n}$$
where
$$z_i := x_{i, 1} + x_{i, 2} + \dots + x_{i, m}$$
Use geometric mean for mean of all steps in a single path $i$:
$$a e^{\frac{x_{i, 1} + x_{i, 2} + \dots + x_{i, m}}{m}} = a \sqrt[m]{e^{x_{i, 1} + x_{i, 2} + \dots + x_{i, m}}}$$

So for **Monte Carlo** of returns after $m$ periods, we 

+ fit a skewed t-distribution to log-returns and use that distribution to simulate $\{x_{i, j}\}_j^m$,
+ for each path $i$, calculate $100\cdot e^{z_i}$,
+ calculate the mean of $\{z_i\}_i^n$:
    + $$\bar{z} = 100\dfrac{e^{z_1} + e^{z_2} + \dots + e^{z_n}}{n}$$

For **Importance Sampling**, we

+ model log-returns on a skewed t-distribution,
+ for each path $i$, calculate $100\cdot e^{z_i}$,
+ fit a skewed t-distribution to $\{z_i\}_i^n$ and use it as our $f$ density function from which we simulate $\{h_i\}_i^n$,
    + In our case $h$ and $z$ are identical, because we have an idea for a distribution to simulate $z$, but in general for IS $h$ could be a function of $z$.
+ calculate $w* = \frac{f}{g^*}$, where $g*$ is our proposal distribution, which minimizes the variance of $h\cdot w$.
+ calculate the arithmetic mean of $\{h_i w_i^{*}\}_i^n$: 
    + $$100 \dfrac{e^{h_1 w_1^{*}} +  e^{h_2 w_2^{*}} + \dots +  e^{h_n w_n^{*}}}{n}$$


## Average of returns vs returns of average

### Math

$$\text{Avg. of returns} := \dfrac{ \left(\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} \right) }{2}$$
$$\text{Returns of avg.} := \left(\dfrac{ x_t + y_t }{2}\right) \Big/ \left(\dfrac{ x_{t-1} + y_{t-1} }{2}\right) \equiv \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

For which $x_1$ and $y_1$ are
$\text{Avg. of returns} = \text{Returns of avg.}$?

$$\dfrac{ \left(\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} \right) }{2} = \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

$$\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} = 2 \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

$$(x_{t-1} + y_{t-1}) x_t y_{t-1} + (x_{t-1} + y_{t-1}) x_{t-1} y_t = 2 (x_{t-1}y_{t-1}x_t + x_{t-1}y_{t-1}y_t)$$

$$(x_{t-1}x_ty_{t-1} + y_{t-1}x_ty_{t-1}) + (x_{t-1}x_{t-1}y_t + x_{t-1}y_{t-1}y_t) = 2(x_{t-1}y_{t-1}x_t + x_{t-1}y_{t-1}y_t)$$
This is not generally true, but true if for instance
$x_{t-1} = y_{t-1}$.

### Example

```{r}
x0 <- 100
y0 <- 200
Rx <- 0.5
Ry <- 1.5
```

Definition: `R = 1+r`

```{r}
cat(paste0("Let x_0 be ", x0, ".\n"))
cat(paste0("Let y_0 be ", y0, ".\n"))
cat("So the initial value of the pf is", x0 + y0, ".\n")
cat("\n")
cat(paste0("Let R_x be ", Rx, ".\n"))
cat(paste0("Let R_y be ", Ry, ".\n"))
```

Then,

```{r}
cat(paste0("x_1 is R_x * x_0 = ", Rx * x0, ".\n"))
cat(paste0("y_1 is R_y * y_0 = ", Ry * y0, ".\n"))
```

Average of returns:

```{r}
cat("0.5 * (R_x + R_y) =", 0.5 * (Rx + Ry), "\n")
```

So here the value of the pf at t=1 should be unchanged from t=0:

```{r}
cat("(x_0 + y_0) * 0.5 * (R_x + R_y) =", (x0 + y0) * 0.5 * (Rx + Ry), "\n")
```

But this is clearly not the case:

```{r}
cat("0.5 * (x_1 + y_1) = 0.5 * (R_x * x_0 + R_y * y_0) =", 0.5 * (Rx * x0 + Ry * y0), "\n")
```

Therefore we should take returns of average, not average of returns!

Let's take the average of log returns instead:

```{r}
cat("0.5 * (log(R_x) + log(R_y)) =", 0.5 * (log(Rx) + log(Ry)), "\n")
```

We now get:

```{r}
cat("(x_0 + y_0) * exp(0.5 * (log(Rx) + log(Ry))) =", (x0 + y0) * exp(0.5 * (log(Rx) + log(Ry))), "\n")
```

So taking the average of log returns doesn't work either.

## Simulation of mix vs mix of simulations

Test if a simulation of a mix (average) of two returns series has the
same distribution as a mix of two simulated returns series.

```{r}
mc_sim <- function(
    num_runs = 1, 
    num_paths = 1000, 
    num_periods = 20,
    m_a = 0,
    s_a = 0.4,
    m_b = 10,
    s_b = 3
    ) {
  
  data_x <- rnorm(num_periods, m_a, s_a)
  data_y <- rnorm(num_periods, m_b, s_b)
  
  m_data_x <- mean(data_x)
  s_data_x <- sd(data_x)
  m_data_y <- mean(data_y)
  s_data_y <- sd(data_y)
  
  cat("m(data_x):", m_data_x, "\n")
  cat("s(data_x):", s_data_x, "\n")
  cat("m(data_y):", m_data_y, "\n")
  cat("s(data_y):", s_data_y, "\n")
  cat("\n")
  
  m_data_xy <- mean(0.5 * data_x + 0.5 * data_y)
  s_data_xy <- sd(0.5 * data_x + 0.5 * data_y)
  
  cat("m(data_x + data_y):", m_data_xy, "\n")
  cat("s(data_x + data_y):", s_data_xy, "\n")
  cat("\n")
  
  run_sim <- function(num_runs) {
    df <- data.frame(
      m_a = rep(0, num_runs), m_b = rep(0, num_runs), 
      s_a = rep(0, num_runs), s_b = rep(0, num_runs)
    )
    for(j in 1:num_runs) {
      sim_x <- rep(0, num_paths)
      sim_y <- rep(0, num_paths)
      sim_xy <- rep(0, num_paths)
      for(i in 1:num_paths) {
        sim_x[i] <- sum(rnorm(num_periods, m_data_x, s_data_x))
        sim_y[i] <- sum(rnorm(num_periods, m_data_y, s_data_y))
        sim_xy[i] <- sum(rnorm(num_periods, m_data_xy, s_data_xy))
      }

      df$m_a[j] <-  mean(0.5 * sim_x + 0.5 * sim_y)
      df$m_b[j] <-  mean(sim_xy)
      df$s_a[j] <-  sd(0.5 * sim_x + 0.5 * sim_y)
      df$s_b[j] <-  sd(sim_xy)
    }
    df
  }
  
  run_sim(num_runs)
}
```

```{r}
mc_sim_df <- mc_sim(
  num_runs = 10, 
  num_paths = 1000, 
  num_periods = 20,
  m_a = 0,
  s_a = 0.4,
  m_b = 10,
  s_b = 3
)
```

m and s of final state of all paths.\
`_a` is mix of simulated returns.\
`_b` is simulated mixed returns.

```{r}
knitr::kable(mc_sim_df, digits = 3)
```

```{r}
summary(mc_sim_df)
```

`_a` and `_b` are very close to equal.\
We attribute the differences to differences in estimating the
distributions in version a and b.

The final state is independent of the order of the preceding steps:

```{r}
vect1 <- c(rnorm(100))
vect2 <- c(sample(vect1, 100))
vect3 <- c(sample(vect1, 100))
path1 <- c(0, cumsum(vect1))
path2 <- c(0, cumsum(vect2))
path3 <- c(0, cumsum(vect3))
plot(path1, type = "l", col = "blue", 
     ylim = c(
       min(c(path1, path2, path3)), 
       max(c(path1, path2, path3))
     )
)
lines(path2, col = "red")
lines(path3, col = "green")
```

So does the order of the steps in the two processes matter, when mixing
simulated returns?

```{r}
vect1a <- c(rnorm(100, 0.05, 0.06))
vect1b <- c(sample(vect1a, 100))
vect2a <- c(rnorm(100, 0.05, 0.06))
vect2b <- c(sample(vect2a, 100))

path1a <- 100 * c(1, cumprod(1 + vect1a))
path1b <- 100 * c(1, cumprod(1 + vect1b))
path2a <- 100 * c(1, cumprod(1 + vect2a))
path2b <- 100 * c(1, cumprod(1 + vect2b))

mix_path_a <- 0.5 * path1a + 0.5 * path2a
mix_path_b <- 0.5 * path1b + 0.5 * path2b

plot(path1a, type = "l", lty = 1, col = "blue", 
     ylim = c(
       min(c(path1a, path1b, path2a, path2b)), 
       max(c(path1a, path1b, path2a, path2b))
     )
)
lines(path1b, lty = 2, col = "blue")
lines(path2a, lty = 1, col = "red")
lines(path2b, lty = 2, col = "red")
```

```{r}
plot(mix_path_a, type = "l", lty = 1, col = "blue", 
     ylim = c(
       min(c(mix_path_a, mix_path_b)), 
       max(c(mix_path_a, mix_path_b))
     )
)
lines(mix_path_b, lty = 2, col = "blue")
```

The order of steps in the individual paths do not matter, because the
mix of simulated paths is a sum of a sum, so the order of terms doesn't
affect the sum. If there is variation it is because the sets preceding
steps are not the same. For instance, the steps between step 1 and 60 in
the plot above are not the same for the two lines.

Recall,
$$\text{Var}(aX+bY) = a^2 \text{Var}(X) + b^2 \text{Var}(Y) + 2ab \text{Cov}(a, b)$$

```{r include=TRUE, echo=TRUE}
var(0.5 * vhr + 0.5 * phr)
0.5^2 * var(vhr) + 0.5^2 * var(phr) + 2 * 0.5 * 0.5 * cov(vhr, phr)
```

Our distribution estimate is based on 13 observations. Is that enough
for a robust estimate? What if we suddenly hit a year like 2008? How
would that affect our estimate?\
Let's try to include the Velliv data from 2007-2010.\
We do this by sampling 13 observations from `vmrl`.

```{r}
n <- 50
test_df <- data.frame(m = rep(0, n), s = rep(0, n))
for(i in 1:n) {
  vmrl_smp <- sample(vmrl, 13)
  test_df$m[i] <- mean(0.5 * vmrl_smp + 0.5 * phr)
  test_df$s[i] <- sd(0.5 * vmrl_smp + 0.5 * phr)
}
summary(test_df)
```

## The meaning of `xi`

The fit for `mhr` has the highest `xi` value of all. This suggests
right-skew:

```{r}
plot((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 1), xlim = c(-20, 20), ylim = c(0, 0.1), type = "l", xlab = "value", ylab = "likelihood", main = "Skew t-distribution density")
lines((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 0.5), col="red")
lines((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 2), col="blue")
legend("topright", c("xi=1", "xi=0.5", "xi=2"), col = c("black", "red", "blue"), lty = 1)
```

## Max vs sum plot

If the Law Of Large Numbers holds true,
$$\dfrac{\max (X_1^p, ..., X^p)}{\sum_{i=1}^n X_i^p} \rightarrow 0$$ for
$n \rightarrow \infty$.

If not, $X$ doesn't have a $p$'th moment.

See Taleb: The Statistical Consequences Of Fat Tails, p. 192


