---
title: "Monthly pension returns analysis: \nVelliv June 2012 - April 2024"
author: Martin Hoshi Vognsen
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
   - \usepackage[fontsize=8pt]{scrextend}
mainfont: SourceSansPro
output: 
  html_document:
    toc: true
    toc_depth: 3
    keep_md: yes
  pdf_document:
    toc: true
    toc_depth: 3
#fontsize: 10pt # for pdf. Limited to 10pt, 11pt and 12pt. Else use scrextend.
params:
  run_sim: FALSE ## TRUE: Run simulations and write output. FALSE: Read saved
                ## simulations output from disk instead of running the simulations.
  run_fits: FALSE
  run_exploratory: TRUE ## Include exploratory report?
  run_individual: TRUE ## Include individual reports?
  run_comparison: TRUE ## Include comparison report? Depends on run_individual.
  run_mc_plot: TRUE ## Depends on run_individual.
  run_is_sim: TRUE
  run_is_plot: TRUE
  include_long: TRUE
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---

```{r chunk1, include=FALSE}
## ****************************** WARNING ******************************
## 
##             This document takes a long time to render!
##
## *********************************************************************
##
## To read previous simulation outputs from disk rather than run simulations,
## set YAML parameter run_sim: TRUE
##
## NOTE:
## When using the "Run All" command, YAML parameters will be ignored. An error
## message "Error in eval(ele) : object 'params' not found" will be printed,
## but the chunks will be evaluated.
```


```{r  chunk2, include=FALSE, eval=FALSE}
################################################################################
##
##                                      SETUP
##
################################################################################
```

```{r chunk3, setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

run_individual <- (params$run_individual + params$run_comparison + params$run_mc_plot > 0)
run_exploratory <- (params$run_exploratory + run_individual > 0)

## Note: In the second line (run_exploratory <- ...), run_individual is correct that - not params$run_individual, because if we choose params$run_exploratory=TRUE and params$run_comparison=TRUE, but params$run_individual=FALSE, we still have to do the individual calculations.
```

```{r chunk4}
library(here)
library(tidyr)
library(tidyjson)
```

```{r chunk5}
source(here("src", "pension-returns_functions.R"))
```

```{r chunk6}
################################################################################
##
##                          GENERATE HTML AND PDF
##                        (DON'T USE KNIT BUTTON!)
##
################################################################################
```

```{r chunk6a, eval=FALSE, include=FALSE}
## Render this document to html and pdf
rmarkdown::render(here::here("pension-returns_monthly.Rmd"), output_format ="all")
```


```{r chunk6b}
################################################################################
##
##              GENERATE PDF MANUALLY AFTER GENERATING HTML AND MD
##
################################################################################

## Note: Will not produce TOC and a few other things will be different.
```

```{bash chunk7, eval=FALSE, include=FALSE}
## Convert md to pdf
## Run manually:
pandoc pension-returns_monthly.md -o pension-returns_monthly.pdf
```


```{r chunk7b}
################################################################################
##
##                             COPY TO SITES FOLDER
##
################################################################################
```

```{bash eval=FALSE, include=FALSE}
## Run manually:
cp ~/R\ work/pension-returns/pension-returns_monthly.html ~/var/site/pension-returns_monthly.html
cp ~/R\ work/pension-returns/pension-returns_monthly.pdf ~/var/site/pension-returns_monthly.pdf
```


```{r chunk8}
################################################################################
##
##                                      INPUTS
##
################################################################################
```

```{r chunk9}
mc_num_paths <- 1e4
mc_num_periods <- 20 * 12
mc_dao <- TRUE
#mc_threshold <- 0.01

## Min saldo for Velliv: kr 12.300
## Assume initial capital of 400.000.
## 100 * 12300/400000 = 3
mc_threshold <- 3
```

Fit log returns to F-S skew standardized Student-t distribution.  
`m`  is the location parameter.  
`s` is the scale parameter.  
`nu` is the estimated shape parameter (degrees of freedom).  
`xi` is the estimated skewness parameter.  

# Returns data  

The long version of Velliv medium risk data runs from January 2007 to April 2024 (incl).  
For January 2007 to May 2012 no low risk and high risk funds existed. For this period the medium risk data is copied into the other funds.  

The short version runs from June 2012 to April 2024.

Velliv returns are including bonus and "DinKapital.  
PFA returns are including "KundeKapital".  

```{r}
## The specified folder must exist in the `data` directory
data_dir <- "monthly"
```

```{r chunk10}
## Headlines for individual sections
headlines <- c(
  "Velliv medium risk (vmr), June 2012 - April 2024",
  "Velliv high risk (vhr), June 2012 - April 2024",
  "PFA medium risk (pmr), June 2012 - April 2024",
  "PFA high risk (phr), June 2012 - April 2024",
  "Mix medium risk (mmr), June 2012 - April 2024",
  "Mix high risk (mhr), June 2012 - April 2024",
  "Mix vmr+phr (vm_ph), June 2012 - April 2024",
  "Mix vhr+pmr (mh_pm), June 2012 - April 2024"
)
```

```{r}
## For Monte Carlo
## When mixing two funds, we are doing average of simulations, not simulation of average!
fit_id <- list(
  1,
  2,
  3,
  4,
  c(1, 3),
  c(2, 4),
  c(1, 4),
  c(2, 3)
)

```


```{r}
pfa_json <- read_json(here("data", "monthly", "pfa_monthly.json"))
```

```{r eval=FALSE}
pfa_json %>% gather_object %>% json_types %>% count(name, type)
```

```{r eval=FALSE}
pfa_json %>% enter_object(afkast_maaned) %>% gather_object %>% json_types %>% count(name, type)
```
```{r}
pfa_monthly_json <- pfa_json %>% enter_object(afkast_maaned) %>% gather_object 
pfa_monthly_json <- pfa_monthly_json$..JSON[[2]]
#pfa_monthly_df <- pfa_monthly_df %>% spread_all
```

```{r}
## Input json: Percent returns
## Output df: Logreturns

pfa_monthly_df <- data.frame()
month_id <- 1
num_years <- length(pfa_monthly_json)
for(i in num_years - seq_along(pfa_monthly_json) + 1) {
  year <- pfa_monthly_json[[i]]$aar
  months <- pfa_monthly_json[[i]]$maaneder[-1]
  for(j in seq_along(months)) {
    pfa_monthly_df[month_id, "year"] <- year
    pfa_monthly_df[month_id, "month"] <- months[[j]]$maaned
    pfa_monthly_df[month_id, "pfa_a"] <- months[[j]]$a$a30
    pfa_monthly_df[month_id, "pfa_b"] <- months[[j]]$b$a30
    pfa_monthly_df[month_id, "pfa_c"] <- months[[j]]$c$a30
    pfa_monthly_df[month_id, "pfa_d"] <- months[[j]]$d$a30
    month_id <- month_id + 1
  }
}
pfa_monthly_df <- pfa_monthly_df %>% 
  ## Start at same date as short Velliv series: 2012-06-30
  ## Get last day of the month by taking first day of next month and subtracting 1
  mutate(date = as.Date(
    paste0(
      year + ifelse(as.numeric(month) == 12, 1, 0), 
      "-", 
      as.numeric(month) %% 12 + 1, 
      "-", 
      "01"
      )
    ) - 1 
  ) %>% 
  filter(date >= as.Date("2012-07-31", format = "%Y-%m-%d"))
pfa_monthly_df <- cbind(pfa_monthly_df$date, pfa_monthly_df[, c("pfa_a", "pfa_b", "pfa_c", "pfa_d")])
colnames(pfa_monthly_df)[1] <- "date"

## Convert percent to gross returns
pfa_monthly_logreturns_df <- pfa_monthly_df
pfa_monthly_logreturns_df[, 2:5] <- log((pfa_monthly_df[, 2:5] /100) + 1)
```

```{r}
## Calculate price index with 2012-06-30 being 100
pfa_monthly_prices_df <- pfa_monthly_logreturns_df
for(i in 2:5) {
  pfa_monthly_prices_df[, i] <- 100 * exp(cumsum(pfa_monthly_logreturns_df[, i]))
}
pfa_monthly_prices_df <- rbind(
    list(as.Date("2012-06-30", format = "%Y-%m-%d"), 100, 100, 100, 100),
    pfa_monthly_prices_df
  )
```

```{r}
## Load data
## Velliv long term log returns
velliv_monthly_prices_long_df <- read.csv(here("data", data_dir, "velliv_monthly_prices.csv"),
  colClasses = c("Date", "numeric", "numeric", "numeric", "numeric")
)

velliv_monthly_prices_df <- velliv_monthly_prices_long_df %>% 
  filter(date >= as.Date("2012-06-30", format = "%Y-%m-%d"))

velliv_monthly_logreturns_long_df <- velliv_monthly_prices_long_df[-1, ]
velliv_monthly_logreturns_long_df[, -1] <- as.data.frame(lapply(
  velliv_monthly_prices_long_df[, -1],
  function(x) {
    log(tail(x, -1) / head(x, -1))
  }
))

#velliv_monthly_logreturns_df <- velliv_monthly_logreturns_long_df[-(1:65), ]
velliv_monthly_logreturns_df <- velliv_monthly_logreturns_long_df %>% 
  filter(date >= as.Date("2012-07-31", format = "%Y-%m-%d"))
rownames(velliv_monthly_logreturns_df) <- as.character(1:nrow(velliv_monthly_logreturns_df))
```


```{r}
monthly_logreturns_df <- cbind(
  velliv_monthly_logreturns_df[, c("date", "medium_risk", "high_risk")],
  pfa_monthly_logreturns_df[, c("pfa_b", "pfa_d")]
)
colnames(monthly_logreturns_df)[2:5] <- c("vmr", "vhr", "pmr", "phr")

data_df_l <- data.frame(vmrl = velliv_monthly_logreturns_long_df$medium_risk) ## Only medium risk data

monthly_logreturns_df <- monthly_logreturns_df %>% 
  mutate(
    mmr = mix_of_logreturns(vmr, pmr),
    mhr = mix_of_logreturns(vhr, phr),
    vmr_phr = mix_of_logreturns(vmr, phr),
    vhr_pmr = mix_of_logreturns(vhr, pmr)
  )

monthly_prices_df <- data.frame(date = pfa_monthly_prices_df$date)
for(i in 2:ncol(monthly_logreturns_df)) {
  monthly_prices_df[, colnames(monthly_logreturns_df)[[i]]] <- prices_from_logreturns(monthly_logreturns_df[, colnames(monthly_logreturns_df)[[i]]])
}

data_df <- monthly_logreturns_df[, -1]

fund_names <- c("vmr", "vhr", "pmr", "phr", "mmr", "mhr", "vmr_phr", "vhr_pmr")
fund_names_l <- "vmrl"
```





```{r fig.width=10}
ggplot(
  pfa_monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("PFA monthly prices"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```

```{r fig.width=10}
ggplot(
  pfa_monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("PFA monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```


```{r chunk11, eval=FALSE}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
cols = gg_color_hue(4)

all_data <- velliv_monthly_prices_df[, -1]
plot(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$guarantee, 
  type = "l",  lty = 1, lwd = 1, 
  xlab = "year", 
  ylab = "Gross returns", 
  col = cols[1], 
  ylim = c(min(all_data), max(all_data)), 
  main = "Velliv gross returns 2007-2023"
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$low_risk, 
  col = cols[2],  
  lty = 1, 
  lwd = 0.8
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$medium_risk, 
  col = cols[3],  
  lty = 1, 
  lwd = 0.6
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$high_risk, 
  col = cols[4],  
  lty = 1, 
  lwd = 0.5
)
legend(
  "bottomright", 
  legend = c("guarantee", "low risk", "medium risk", "high risk"), 
  col = cols, 
  lty = 1, 
  lwd = 2, 
  cex = 0.75
)
```

```{r fig.width=10}
ggplot(
  velliv_monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly prices"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```


```{r eval=params$include_long, fig.width=10}
ggplot(
  velliv_monthly_prices_long_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly prices - long"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```


```{r chunk11b, eval=FALSE}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
cols = gg_color_hue(4)

all_data <- exp(velliv_monthly_logreturns_df[, -1])
plot(
  x = velliv_monthly_logreturns_df$date, 
  y = exp(velliv_monthly_logreturns_df$guarantee), 
  type = "l",  
  lty = 1, 
  lwd = 1, 
  xlab = "year", 
  ylab = "Gross returns", 
  col = cols[1], 
  ylim = c(min(all_data), max(all_data)), 
  main = "Velliv gross returns 2007-2023"
)
lines(
  x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$low_risk), col = cols[2],  lty = 1, lwd = 0.8)
lines(x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$medium_risk), col = cols[3],  lty = 1, lwd = 0.6)
lines(x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$high_risk), col = cols[4],  lty = 1, lwd = 0.5)
legend("bottomright", legend = c("guarantee", "low risk", "medium risk", "high risk"), col = cols, lty = 1, lwd = 2, cex = 0.75)
```

```{r fig.width=10}
ggplot(
  velliv_monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```

```{r eval=params$include_long, fig.width=10}
ggplot(
  velliv_monthly_logreturns_long_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly gross returns - long"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```

```{r fig.width=10}
ggplot(
  monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv and PFA monthly prices"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```

```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv and PFA monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```
```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    group_by(date = lubridate::year(date)) %>%
    summarize(
      vmr = sum(vmr),
      vhr = sum(vhr),
      pmr = sum(pmr),
      phr = sum(phr),
      mmr = sum(mmr),
      mhr = sum(mhr),
      vmr_phr = sum(vmr_phr),
      vhr_pmr = sum(vhr_pmr)
    ) %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)),
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  #scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  scale_x_continuous(breaks = lubridate::year(monthly_logreturns_df$date)) +
  labs(
    title = paste("Velliv+PFA annual gross returns"), 
    subtitle = "Aggregated from monthly gross returns", 
    x = "date", 
    y = "gross return"
  )
```





```{r chunk12}
################################################################################
##
##                                   CALCULATIONS
##
################################################################################
```

```{r chunk13}
##                                       FITS
```

```{r chunk14, eval=(run_individual && params$run_fits)}
fits <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), distribution = "sstd")
  })}
)

fits_std <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "std")
  })}
)

fits_norm <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "normal")
  })}
)
```


```{r eval=(run_individual && params$include_long && params$run_fits)}
fits_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), distribution = "sstd")
  })}
)

fits_std_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), method = "Nelder-Mead", distribution = "std")
  })}
)

fits_norm_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), method = "Nelder-Mead", distribution = "normal")
  })}
)
```

```{r chunk15, eval=run_individual && params$run_fits}
saveRDS(fits, file = here("data", data_dir,  "fits.RData"))
saveRDS(fits_std, file = here("data", data_dir, "fits_std.RData"))
saveRDS(fits_norm, file = here("data", data_dir, "fits_norm.RData"))
```

```{r chunk15b, eval=(run_individual && params$include_long && params$run_fits)}
saveRDS(fits_l, file = here("data", data_dir,  "fits_l.RData"))
saveRDS(fits_std_l, file = here("data", data_dir, "fits_std_l.RData"))
saveRDS(fits_norm_l, file = here("data", data_dir, "fits_norm_l.RData"))
```

```{r chunk15c, eval=(run_individual && !params$run_fits)}
fits <- readRDS(file = here("data", data_dir,  "fits.RData"))
fits_std <- readRDS(file = here("data", data_dir, "fits_std.RData"))
fits_norm <- readRDS(file = here("data", data_dir, "fits_norm.RData"))
```

```{r chunk15d, eval=(run_individual && !params$run_fits && params$include_long)}
fits_l <- readRDS(file = here("data", data_dir,  "fits_l.RData"))
fits_std_l <- readRDS(file = here("data", data_dir, "fits_std_l.RData"))
fits_norm_l <- readRDS(file = here("data", data_dir, "fits_norm_l.RData"))
```


```{r eval=run_individual}
fit_summary <- as.data.frame(lapply(
  fits,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary) <- fund_names
rownames(fit_summary) <- c("m", "s", "nu", "xi", "R^2")
```

```{r eval=run_individual}
fit_summary_std <- as.data.frame(lapply(
  fits_std,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_std) <- fund_names
rownames(fit_summary_std) <- c("m", "s", "nu", "R^2")
```

```{r eval=run_individual}
fit_summary_norm <- as.data.frame(lapply(
  fits_norm,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_norm) <- fund_names
rownames(fit_summary_norm) <- c("m", "s", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_l <- as.data.frame(lapply(
  fits_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_l) <- fund_names_l
rownames(fit_summary_l) <- c("m", "s", "nu", "xi", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_std_l <- as.data.frame(lapply(
  fits_std_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_std_l) <- fund_names_l
rownames(fit_summary_std_l) <- c("m", "s", "nu", "R^2")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_norm_l <- as.data.frame(lapply(
  fits_norm_l,
  function(fit) {
     c(fit$dist_params, fit$r_squared)
  }
))
colnames(fit_summary_norm_l) <- fund_names_l
rownames(fit_summary_norm_l) <- c("m", "s", "R^2")
```


```{r eval=(params$run_comparison && params$run_sim)}
kappa_df <- data.frame(matrix(numeric(length(fund_names)), ncol = length(fund_names)))

for(i in seq_along(fits)) {
  kappa_df[1, i] <- f_kappa(
    n0 = 1, 
    n = 30, 
    mean = fits[[i]]$dist_params[1], 
    sd = fits[[i]]$dist_params[2], 
    nu = fits[[i]]$dist_params[3], 
    xi = fits[[i]]$dist_params[4], 
    num_sim = 1e5
  )
}

colnames(kappa_df) <- fund_names
```

```{r eval=(params$run_comparison && params$run_sim)}
saveRDS(kappa_df, file = here("data", data_dir,  "kappa_df.RData"))
```

```{r eval=(params$run_comparison && !params$run_sim)}
kappa_df <- readRDS(file = here("data", data_dir,  "kappa_df.RData"))
```

```{r eval=(params$run_comparison && params$run_sim)}
n_min_df <- data.frame(matrix(numeric(length(fund_names)), ncol = length(fund_names)))

for(i in seq_along(fits)) {
  n_min_df[1, i] <- f_n_min(
    n_g = 30, 
    mean = fits[[i]]$dist_params[1], 
    sd = fits[[i]]$dist_params[2], 
    nu = fits[[i]]$dist_params[3], 
    xi = fits[[i]]$dist_params[4], 
    num_sim = 1e5, 
    approx = FALSE
  )
}

colnames(n_min_df) <- fund_names
```

```{r eval=(params$run_comparison && params$run_sim)}
saveRDS(n_min_df, file = here("data", data_dir,  "n_min_df.RData"))
```

```{r eval=(params$run_comparison && !params$run_sim)}
n_min_df <- readRDS(file = here("data", data_dir,  "n_min_df.RData"))
```


```{r chunk16}
##                                     MONTE CARLO
```

```{r chunk17, eval=(params$run_sim && run_individual)}
num_fits <- length(fits)

mc_output <- list()
for(i in seq_along(fits)) {
  message(paste0("\nsstd: ", i, " of ", num_fits, "\n"))
  mc_output[[i]] <- mc_simulation(
    fits[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "sstd"
  )
}

mc_output_std <- list()
for(i in seq_along(fits_std)) {
  message(paste0("\nstd: ", i, " of, ", num_fits, "\n"))
  mc_output_std[[i]] <- mc_simulation(
    fits_std[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "std"
  )
}

mc_output_norm <- list()
for(i in seq_along(fits_norm)) {
  message(paste0("\nnormal: ", i, " of, ", num_fits, "\n"))
  mc_output_norm[[i]] <- mc_simulation(
    fits_norm[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "normal"
  )
}
```

```{r chunk17a, eval=(params$run_sim && run_individual && params$include_long)}
fit_id_l <- list(
  1
)

num_fits_l <- length(fits_l)

mc_output_l <- list()
for(i in seq_along(fits_l)) {
  message(paste0("\nsstd: ", i, " of ", num_fits_l, "\n"))
  mc_output_l[[i]] <- mc_simulation(
    fits_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "sstd"
  )
}

mc_output_std_l <- list()
for(i in seq_along(fits_std_l)) {
  message(paste0("\nstd: ", i, " of, ", num_fits, "\n"))
  mc_output_std_l[[i]] <- mc_simulation(
    fits_std_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "std"
  )
}

mc_output_norm_l <- list()
for(i in seq_along(fits_norm_l)) {
  message(paste0("\nnormal: ", i, " of, ", num_fits, "\n"))
  mc_output_norm_l[[i]] <- mc_simulation(
    fits_norm_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "normal"
  )
}
```


```{r chunk17b, eval=(params$run_sim && run_individual)}
saveRDS(mc_output, file = here("data", data_dir, "mc_output.RData"))
saveRDS(mc_output_std, file = here("data", data_dir, "mc_output_std.RData"))
saveRDS(mc_output_norm, file = here("data", data_dir, "mc_output_norm.RData"))
```

```{r chunk17c, eval=(params$run_sim && run_individual && params$include_long)}
saveRDS(mc_output_l, file = here("data", data_dir, "mc_output_l.RData"))
saveRDS(mc_output_std_l, file = here("data", data_dir, "mc_output_std_l.RData"))
saveRDS(mc_output_norm_l, file = here("data", data_dir, "mc_output_norm_l.RData"))
```

```{r chunk18, eval=(run_individual && !params$run_sim)}
mc_output <- readRDS(file = here("data", data_dir, "mc_output.RData"))
mc_output_std <- readRDS(file = here("data", data_dir, "mc_output_std.RData"))
mc_output_norm <- readRDS(file = here("data", data_dir, "mc_output_norm.RData"))
```

```{r chunk18b, eval=(run_individual && !params$run_sim && params$include_long)}
mc_output_l <- readRDS(file = here("data", data_dir, "mc_output_l.RData"))
mc_output_std_l <- readRDS(file = here("data", data_dir, "mc_output_std_l.RData"))
mc_output_norm_l<- readRDS(file = here("data", data_dir, "mc_output_norm_l.RData"))
```


```{r chunk19}
##                                     MAX SUM PLOTS
```

```{r chunk20, eval=(params$run_sim && run_individual)}
max_sum_plots <- list()
for(i in seq_along(fits)) {
  max_sum_plots[[i]] <- plot_max_sum(
    fits[fit_id[[i]]], 
    mc_num_paths,
    distribution = "sstd"
  )
}

max_sum_plots_std <- list()
for(i in seq_along(fits_std)) {
  max_sum_plots_std[[i]] <- plot_max_sum(
    fits_std[fit_id[[i]]], 
    mc_num_paths, 
    distribution = "std"
  )
}

max_sum_plots_norm <- list()
for(i in seq_along(fits_norm)) {
  max_sum_plots_norm[[i]] <- plot_max_sum(
    fits_norm[fit_id[[i]]], 
    mc_num_paths, 
    distribution = "norm"
  )
}
```

```{r chunk20a, eval=(params$run_sim && run_individual && params$include_long)}
max_sum_plots_l <- list()
for(i in seq_along(fits_l)) {
  max_sum_plots_l[[i]] <- plot_max_sum(
    fits_l[fit_id_l[[i]]], 
    mc_num_paths,
    distribution = "sstd"
  )
}

max_sum_plots_std_l <- list()
for(i in seq_along(fits_std_l)) {
  max_sum_plots_std_l[[i]] <- plot_max_sum(
    fits_std_l[fit_id_l[[i]]], 
    mc_num_paths, 
    distribution = "std"
  )
}

max_sum_plots_norm_l <- list()
for(i in seq_along(fits_norm_l)) {
  max_sum_plots_norm_l[[i]] <- plot_max_sum(
    fits_norm_l[fit_id_l[[i]]], 
    mc_num_paths, 
    distribution = "norm"
  )
}
```

```{r chunk20b, eval=(params$run_sim && run_individual)}
saveRDS(max_sum_plots, here("data", data_dir, "max_sum_plots.RData"))
saveRDS(max_sum_plots_std, here("data", data_dir, "max_sum_plots_std.RData"))
saveRDS(max_sum_plots_norm, here("data", data_dir, "max_sum_plots_norm.RData"))
```

```{r chunk20c, eval=(params$run_sim && run_individual && params$include_long)}
saveRDS(max_sum_plots_l, here("data", data_dir, "max_sum_plots_l.RData"))
saveRDS(max_sum_plots_std_l, here("data", data_dir, "max_sum_plots_std_l.RData"))
saveRDS(max_sum_plots_norm_l, here("data", data_dir, "max_sum_plots_norm_l.RData"))
```



```{r chunk21, eval=(run_individual && !params$run_sim)}
max_sum_plots <- readRDS(file = here("data", data_dir, "max_sum_plots.RData"))
max_sum_plots_std <- readRDS(file = here("data", data_dir, "max_sum_plots_std.RData"))
max_sum_plots_norm <- readRDS(file = here("data", data_dir, "max_sum_plots_norm.RData"))
```

```{r chunk21b, eval=(run_individual && !params$run_sim)}
max_sum_plots_l <- readRDS(file = here("data", data_dir, "max_sum_plots_l.RData"))
max_sum_plots_std_l <- readRDS(file = here("data", data_dir, "max_sum_plots_std_l.RData"))
max_sum_plots_norm_l <- readRDS(file = here("data", data_dir, "max_sum_plots_norm_l.RData"))
```


```{r chunk22}
##                                  IMPORTANCE SAMPLING
```

```{r chunk23, eval=(params$run_sim && run_individual && params$run_is_sim)}
is_g_fits <-  list()
is_output <- list()
for(i in seq_along(fits)) {
  is_g_fits[[i]] <- is_proposal(
    x_i_fit = fits[fit_id[[i]]],
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    obj_func_plot = TRUE,
    init_par = c(2, 0.5),
    method = "L-BFGS-B",
    lower = c(1, 0.1),
    upper = c(3, 0.9)
  )
  
  is_output[[i]] <- importance_sampling(
    x_i_fit = fits[fit_id[[i]]], 
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    g_n_params = is_g_fits[[i]]$par, 
    mode = 1
  )
}
```

```{r chunk23b, eval=(params$run_sim && run_individual && params$run_is_sim)}
saveRDS(is_g_fits, here("data", data_dir, "is_g_fits.RData"))
saveRDS(is_output, here("data", data_dir, "is_output.RData"))
```

```{r chunk24, eval=(run_individual && !params$run_sim && !params$run_is_sim && params$run_is_plot)}
is_g_fits <- readRDS(file = here("data", data_dir, "is_g_fits.RData"))
is_output <- readRDS(file = here("data", data_dir, "is_output.RData"))
```


```{r chunk25}
###########################################################################
##
##                            GENERATE REPORTS
##
###########################################################################
```

```{r chunk26, eval=params$run_exploratory, results='asis'}
exploratory_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-exploratory.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(exploratory_report), sep = '\n')
```

```{r chunk27, eval=params$run_comparison, results='asis'}
comparison_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-comparison.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(comparison_report), sep = '\n')
```

```{r chunk28, eval=params$run_individual, results='asis'}
long <- FALSE
individual_reports <- lapply(
  seq_along(data_df),
  function(i) {
    knitr::knit_child(
      here("child_docs", "pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```

```{r chunk28b, eval=params$run_individual, results='asis', eval=params$include_long}
long <- TRUE
individual_reports <- lapply(
  seq_along(data_df_l),
  function(i) {
    knitr::knit_child(
      here("child_docs", "pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```





