---
title: "Monthly pension returns analysis: \nVelliv June 2012 - April 2024"
author: Martin Hoshi Vognsen
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[fontsize=8pt]{scrextend}
output: 
  html_document:
    toc: true
    toc_depth: 3
    keep_md: yes
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
#fontsize: 10pt # for pdf. Limited to 10pt, 11pt and 12pt. Else use scrextend.
params:
  run_sim: FALSE ## TRUE: Run simulations and write output. FALSE: Read saved
                ## simulations output from disk instead of running the simulations.
  run_fits: FALSE
  run_exploratory: TRUE ## Include exploratory report?
  run_individual: TRUE ## Include individual reports?
  run_comparison: TRUE ## Include comparison report? Depends on run_individual.
  run_mc_plot: TRUE ## Depends on run_individual.
  run_is_sim: TRUE
  run_is_plot: TRUE
  include_long: TRUE
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---

```{r chunk1, include=FALSE}
## ****************************** WARNING ******************************
## 
##             This document takes a long time to render!
##
## *********************************************************************
##
## To read previous simulation outputs from disk rather than run simulations,
## set YAML parameter run_sim: TRUE
##
## NOTE:
## When using the "Run All" command, YAML parameters will be ignored. An error
## message "Error in eval(ele) : object 'params' not found" will be printed,
## but the chunks will be evaluated.
```


```{r  chunk2, include=FALSE, eval=FALSE}
################################################################################
##
##                                      SETUP
##
################################################################################
```

```{r chunk3, setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

run_individual <- (params$run_individual + params$run_comparison + params$run_mc_plot > 0)
run_exploratory <- (params$run_exploratory + run_individual > 0)

## Note: In the second line (run_exploratory <- ...), run_individual is correct that - not params$run_individual, because if we choose params$run_exploratory=TRUE and params$run_comparison=TRUE, but params$run_individual=FALSE, we still have to do the individual calculations.
```

```{r chunk4}
library(here)
library(tidyr)
library(tidyjson)
library(lubridate)   # year() used below; attach explicitly for non-interactive render
```

```{r chunk5}
source(here("src", "pension-returns_functions.R"))
```

```{r chunk6}
################################################################################
##
##                          GENERATE HTML AND PDF
##                        (DON'T USE KNIT BUTTON!)
##
################################################################################
```

```{r chunk6a, eval=FALSE, include=FALSE}
## Render this document to html and pdf
rmarkdown::render(here::here("pension-returns_monthly.Rmd"), output_format ="all")
```


```{r chunk6b}
################################################################################
##
##              GENERATE PDF MANUALLY AFTER GENERATING HTML AND MD
##
################################################################################

## Note: Will not produce TOC and a few other things will be different.
```

```{bash chunk7, eval=FALSE, include=FALSE}
## Convert md to pdf
## Run manually:
pandoc pension-returns_monthly.md -o pension-returns_monthly.pdf
```


```{r chunk7b}
################################################################################
##
##                             COPY TO SITES FOLDER
##
################################################################################
```

```{bash eval=FALSE, include=FALSE}
## Run manually:
cp ~/R\ work/pension-returns/pension-returns_monthly.html ~/var/site/pension-returns_monthly.html
cp ~/R\ work/pension-returns/pension-returns_monthly.pdf ~/var/site/pension-returns_monthly.pdf
```


```{r chunk8}
################################################################################
##
##                                      INPUTS
##
################################################################################
```

```{r chunk9}
mc_num_paths <- 1e4
mc_num_periods <- 20 * 12
mc_dao <- TRUE
#mc_threshold <- 0.01

## Min saldo for Velliv: kr 12.300
## Assume initial capital of 400.000.
## 100 * 12300/400000 = 3
mc_threshold <- 3
```

# Summary

A comparison of returns from pension plans provided by PFA and Velliv respectively:  
- Velliv medium risk (vmr), June 2012 - April 2024  
- Velliv high risk (vhr), June 2012 - April 2024  
- PFA medium risk (pmr), June 2012 - April 2024  
- PFA high risk (phr), June 2012 - April 2024  

The combinations of low and high risk funds for each PFA plan can be found here:  
https://www.pfa.dk/privat/opsparing/aktivsammensaetning/

The Velliv website only provides a list of funds.  

The analysis also looks at these mixes of plans:  
- Mix medium risk (mmr), June 2012 - April 2024  
- Mix high risk (mhr), June 2012 - April 2024  
- Mix vmr+phr (vm_ph), June 2012 - April 2024  
- Mix vhr+pmr (mh_pm), June 2012 - April 2024





We observe that  
1. **The plans are weighted blends of the same two underlying funds.** Within
each provider the high-risk plan is almost exactly a scalar multiple of the medium-risk plan
($\beta \approx 1.2$ for Velliv, $\approx 1.5$ for PFA, intercept $\approx 0$), and a single
common factor dominates the variation across all four plans. PFA's own product description
confirms it: profiles B and D are fixed weighted blends of the *same two* underlying funds.

2. **"Risk" here means deeper drawdowns and start-date-dependent path ordering, not a
different kind of bet.** The high-risk plan amplifies the same factor in both directions: it
draws down more (worst peak-to-trough $-17\%$ vs $-15\%$ for Velliv, $-16\%$ vs $-12\%$ for
PFA), and its cumulative path drops below the medium plan's whenever the index is started
just before a drawdown -- as it does around the COVID crash.

3. **The sample still spans only about one market cycle.** Monthly data (142 observations) is
firmer than the annual report's 13, and -- unlike the annual figures -- it *does* capture the
COVID-2020 and 2022 drawdowns. But it contains no 2008-scale event; for that the long Velliv
series is needed. The qualitative structure -- fat tails, leverage, asymmetric drawdowns -- is
robust, while precise tail numbers should be read with Taleb's "better stay home" in mind.

The Discussion at the end develops these points.

# Returns data

The returns data for PFA can be seen here:  
https://www.pfa.dk/privat/opsparing/pfa-afkast/  

The data for Velliv can be seen here:  
https://www.velliv.dk/opsparing/vaekstpension/vaekstpension-aktiv/afkast  

Fit log returns to F-S skew standardized Student-t distribution.  
`m`  is the location parameter.  
`s` is the scale parameter.  
`nu` is the estimated shape parameter (degrees of freedom).  
`xi` is the estimated skewness parameter.  

The long version of Velliv medium risk data runs from January 2007 to April 2024 (incl).  
For January 2007 to May 2012 no low risk and high risk funds existed. For this period the medium risk data is copied into the other funds.  

The short version runs from June 2012 to April 2024.

Velliv returns are including bonus and "DinKapital.  
PFA returns are including "KundeKapital".  

```{r}
## The specified folder must exist in the `data` directory
data_dir <- "monthly"
```

```{r chunk10}
## Headlines for individual sections
headlines <- c(
  "Velliv medium risk (vmr), June 2012 - April 2024",
  "Velliv high risk (vhr), June 2012 - April 2024",
  "PFA medium risk (pmr), June 2012 - April 2024",
  "PFA high risk (phr), June 2012 - April 2024",
  "Mix medium risk (mmr), June 2012 - April 2024",
  "Mix high risk (mhr), June 2012 - April 2024",
  "Mix vmr+phr (vm_ph), June 2012 - April 2024",
  "Mix vhr+pmr (mh_pm), June 2012 - April 2024"
)
```

```{r}
## For Monte Carlo -- PURE BLACK BOX.
## Each fund (mixes included) is simulated from its OWN single fit to its observed
## total-value series: "simulation of the average", NOT "average of simulations".
## The former component-recombination (e.g. c(1, 3) for a mix) drew the two
## components independently, assuming zero correlation between near-collinear funds,
## and is abandoned. fits[[5..8]] are the single fits to the realised mix series
## (mmr, mhr, vmr_phr, vhr_pmr) built via mix_of_logreturns().
fit_id <- list(1, 2, 3, 4, 5, 6, 7, 8)

```


```{r}
pfa_json <- read_json(here("data", "monthly", "pfa_monthly.json"))
```

```{r eval=FALSE}
pfa_json %>% gather_object %>% json_types %>% count(name, type)
```

```{r eval=FALSE}
pfa_json %>% enter_object(afkast_maaned) %>% gather_object %>% json_types %>% count(name, type)
```
```{r}
pfa_monthly_json <- pfa_json %>% enter_object(afkast_maaned) %>% gather_object 
pfa_monthly_json <- pfa_monthly_json$..JSON[[2]]
#pfa_monthly_df <- pfa_monthly_df %>% spread_all
```

```{r}
## Input json: Percent returns
## Output df: Logreturns

pfa_monthly_df <- data.frame()
month_id <- 1
num_years <- length(pfa_monthly_json)
for(i in num_years - seq_along(pfa_monthly_json) + 1) {
  year <- pfa_monthly_json[[i]]$aar
  months <- pfa_monthly_json[[i]]$maaneder[-1]
  for(j in seq_along(months)) {
    pfa_monthly_df[month_id, "year"] <- year
    pfa_monthly_df[month_id, "month"] <- months[[j]]$maaned
    pfa_monthly_df[month_id, "pfa_a"] <- months[[j]]$a$a30
    pfa_monthly_df[month_id, "pfa_b"] <- months[[j]]$b$a30
    pfa_monthly_df[month_id, "pfa_c"] <- months[[j]]$c$a30
    pfa_monthly_df[month_id, "pfa_d"] <- months[[j]]$d$a30
    month_id <- month_id + 1
  }
}
pfa_monthly_df <- pfa_monthly_df %>% 
  ## Start at same date as short Velliv series: 2012-06-30
  ## Get last day of the month by taking first day of next month and subtracting 1
  mutate(date = as.Date(
    paste0(
      year + ifelse(as.numeric(month) == 12, 1, 0), 
      "-", 
      as.numeric(month) %% 12 + 1, 
      "-", 
      "01"
      )
    ) - 1 
  ) %>% 
  filter(date >= as.Date("2012-07-31", format = "%Y-%m-%d"))
pfa_monthly_df <- cbind(pfa_monthly_df$date, pfa_monthly_df[, c("pfa_a", "pfa_b", "pfa_c", "pfa_d")])
colnames(pfa_monthly_df)[1] <- "date"

## Convert percent to gross returns
pfa_monthly_logreturns_df <- pfa_monthly_df
pfa_monthly_logreturns_df[, 2:5] <- log((pfa_monthly_df[, 2:5] /100) + 1)
```

```{r}
## Calculate price index with 2012-06-30 being 100
pfa_monthly_prices_df <- pfa_monthly_logreturns_df
for(i in 2:5) {
  pfa_monthly_prices_df[, i] <- 100 * exp(cumsum(pfa_monthly_logreturns_df[, i]))
}
pfa_monthly_prices_df <- rbind(
    list(as.Date("2012-06-30", format = "%Y-%m-%d"), 100, 100, 100, 100),
    pfa_monthly_prices_df
  )
```

```{r}
## Load data
## Velliv long term log returns
velliv_monthly_prices_long_df <- read.csv(here("data", data_dir, "velliv_monthly_prices.csv"),
  colClasses = c("Date", "numeric", "numeric", "numeric", "numeric")
)

velliv_monthly_prices_df <- velliv_monthly_prices_long_df %>% 
  filter(date >= as.Date("2012-06-30", format = "%Y-%m-%d"))

velliv_monthly_logreturns_long_df <- velliv_monthly_prices_long_df[-1, ]
velliv_monthly_logreturns_long_df[, -1] <- as.data.frame(lapply(
  velliv_monthly_prices_long_df[, -1],
  function(x) {
    log(tail(x, -1) / head(x, -1))
  }
))

#velliv_monthly_logreturns_df <- velliv_monthly_logreturns_long_df[-(1:65), ]
velliv_monthly_logreturns_df <- velliv_monthly_logreturns_long_df %>% 
  filter(date >= as.Date("2012-07-31", format = "%Y-%m-%d"))
rownames(velliv_monthly_logreturns_df) <- as.character(1:nrow(velliv_monthly_logreturns_df))
```


```{r}
monthly_logreturns_df <- cbind(
  velliv_monthly_logreturns_df[, c("date", "medium_risk", "high_risk")],
  pfa_monthly_logreturns_df[, c("pfa_b", "pfa_d")]
)
colnames(monthly_logreturns_df)[2:5] <- c("vmr", "vhr", "pmr", "phr")

data_df_l <- data.frame(vmrl = velliv_monthly_logreturns_long_df$medium_risk) ## Only medium risk data

monthly_logreturns_df <- monthly_logreturns_df %>% 
  mutate(
    mmr = mix_of_logreturns(vmr, pmr),
    mhr = mix_of_logreturns(vhr, phr),
    vmr_phr = mix_of_logreturns(vmr, phr),
    vhr_pmr = mix_of_logreturns(vhr, pmr)
  )

monthly_prices_df <- data.frame(date = pfa_monthly_prices_df$date)
for(i in 2:ncol(monthly_logreturns_df)) {
  monthly_prices_df[, colnames(monthly_logreturns_df)[[i]]] <- prices_from_logreturns(monthly_logreturns_df[, colnames(monthly_logreturns_df)[[i]]])
}

data_df <- monthly_logreturns_df[, -1]

fund_names <- c("vmr", "vhr", "pmr", "phr", "mmr", "mhr", "vmr_phr", "vhr_pmr")
fund_names_l <- "vmrl"
```





```{r fig.width=10}
ggplot(
  pfa_monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("PFA monthly prices"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```

```{r fig.width=10}
ggplot(
  pfa_monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("PFA monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```


```{r chunk11, eval=FALSE}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
cols = gg_color_hue(4)

all_data <- velliv_monthly_prices_df[, -1]
plot(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$guarantee, 
  type = "l",  lty = 1, lwd = 1, 
  xlab = "year", 
  ylab = "Gross returns", 
  col = cols[1], 
  ylim = c(min(all_data), max(all_data)), 
  main = "Velliv gross returns 2007-2023"
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$low_risk, 
  col = cols[2],  
  lty = 1, 
  lwd = 0.8
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$medium_risk, 
  col = cols[3],  
  lty = 1, 
  lwd = 0.6
)
lines(
  x = velliv_monthly_prices_df$date, 
  y = velliv_monthly_prices_df$high_risk, 
  col = cols[4],  
  lty = 1, 
  lwd = 0.5
)
legend(
  "bottomright", 
  legend = c("guarantee", "low risk", "medium risk", "high risk"), 
  col = cols, 
  lty = 1, 
  lwd = 2, 
  cex = 0.75
)
```

```{r fig.width=10}
ggplot(
  velliv_monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly prices"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```


```{r eval=params$include_long, fig.width=10}
ggplot(
  velliv_monthly_prices_long_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly prices - long"), 
    subtitle = "", 
    x = "date", 
    y = "price index"
  )
```


```{r chunk11b, eval=FALSE}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
cols = gg_color_hue(4)

all_data <- exp(velliv_monthly_logreturns_df[, -1])
plot(
  x = velliv_monthly_logreturns_df$date, 
  y = exp(velliv_monthly_logreturns_df$guarantee), 
  type = "l",  
  lty = 1, 
  lwd = 1, 
  xlab = "year", 
  ylab = "Gross returns", 
  col = cols[1], 
  ylim = c(min(all_data), max(all_data)), 
  main = "Velliv gross returns 2007-2023"
)
lines(
  x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$low_risk), col = cols[2],  lty = 1, lwd = 0.8)
lines(x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$medium_risk), col = cols[3],  lty = 1, lwd = 0.6)
lines(x = velliv_monthly_logreturns_df$date, y = exp(velliv_monthly_logreturns_df$high_risk), col = cols[4],  lty = 1, lwd = 0.5)
legend("bottomright", legend = c("guarantee", "low risk", "medium risk", "high risk"), col = cols, lty = 1, lwd = 2, cex = 0.75)
```

```{r fig.width=10}
ggplot(
  velliv_monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```

```{r eval=params$include_long, fig.width=10}
ggplot(
  velliv_monthly_logreturns_long_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv monthly gross returns - long"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```

```{r fig.width=10}
ggplot(
  monthly_prices_df %>% gather(key = "fund", value = "price", -date), 
  aes(x = date, y = price)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv and PFA monthly prices"),
    subtitle = "",
    x = "date",
    y = "price index"
  )
```




```{r chunk12}
################################################################################
##
##                                   CALCULATIONS
##
################################################################################
```

```{r chunk13}
##                                       FITS
```

```{r chunk14, eval=(run_individual && params$run_fits)}
fits <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), distribution = "sstd")
  })}
)

fits_std <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "std")
  })}
)

fits_norm <- lapply(
  seq_along(data_df),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df[i]), method = "Nelder-Mead", distribution = "normal")
  })}
)
```


```{r eval=(run_individual && params$include_long && params$run_fits)}
fits_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), distribution = "sstd")
  })}
)

fits_std_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), method = "Nelder-Mead", distribution = "std")
  })}
)

fits_norm_l <- lapply(
  seq_along(data_df_l),
  function(i) {suppressWarnings({
    fit_distribution(unlist(data_df_l[i]), method = "Nelder-Mead", distribution = "normal")
  })}
)
```

```{r chunk15, eval=run_individual && params$run_fits}
saveRDS(fits, file = here("data", data_dir,  "fits.RData"))
saveRDS(fits_std, file = here("data", data_dir, "fits_std.RData"))
saveRDS(fits_norm, file = here("data", data_dir, "fits_norm.RData"))
```

```{r chunk15b, eval=(run_individual && params$include_long && params$run_fits)}
saveRDS(fits_l, file = here("data", data_dir,  "fits_l.RData"))
saveRDS(fits_std_l, file = here("data", data_dir, "fits_std_l.RData"))
saveRDS(fits_norm_l, file = here("data", data_dir, "fits_norm_l.RData"))
```

```{r chunk15c, eval=(run_individual && !params$run_fits)}
fits <- readRDS(file = here("data", data_dir,  "fits.RData"))
fits_std <- readRDS(file = here("data", data_dir, "fits_std.RData"))
fits_norm <- readRDS(file = here("data", data_dir, "fits_norm.RData"))
```

```{r chunk15d, eval=(run_individual && !params$run_fits && params$include_long)}
fits_l <- readRDS(file = here("data", data_dir,  "fits_l.RData"))
fits_std_l <- readRDS(file = here("data", data_dir, "fits_std_l.RData"))
fits_norm_l <- readRDS(file = here("data", data_dir, "fits_norm_l.RData"))
```


```{r chunk15e-ad, eval=(run_individual && params$run_fits)}
## Anderson-Darling A^2 and parametric-bootstrap p-value per fund x distribution.
## Heavy (refits inside the bootstrap), so gated on run_fits and cached like the fits.
ad_B    <- 499
ad_df   <- data.frame(matrix(numeric(3 * length(fund_names)), ncol = length(fund_names)))
ad_p_df <- data.frame(matrix(numeric(3 * length(fund_names)), ncol = length(fund_names)))
for(i in seq_along(fits)) {
  rk <- ad_gof(fits[[i]],      "sstd",   B = ad_B)
  rs <- ad_gof(fits_std[[i]],  "std",    B = ad_B)
  rn <- ad_gof(fits_norm[[i]], "normal", B = ad_B)
  ad_df[, i]   <- c(rk$A2, rs$A2, rn$A2)
  ad_p_df[, i] <- c(rk$p,  rs$p,  rn$p)
}
colnames(ad_df) <- colnames(ad_p_df) <- fund_names
rownames(ad_df) <- rownames(ad_p_df) <- c("sstd", "std", "normal")
saveRDS(ad_df,   file = here("data", data_dir, "ad_df.RData"))
saveRDS(ad_p_df, file = here("data", data_dir, "ad_p_df.RData"))
```

```{r chunk15f-ad, eval=(run_individual && !params$run_fits)}
ad_df   <- readRDS(file = here("data", data_dir, "ad_df.RData"))
ad_p_df <- readRDS(file = here("data", data_dir, "ad_p_df.RData"))
```


```{r eval=run_individual}
fit_summary <- as.data.frame(lapply(
  fits,
  function(fit) {
     c(fit$dist_params, fit$ppcc)
  }
))
colnames(fit_summary) <- fund_names
rownames(fit_summary) <- c("m", "s", "nu", "xi", "PPCC")
```

```{r eval=run_individual}
fit_summary_std <- as.data.frame(lapply(
  fits_std,
  function(fit) {
     c(fit$dist_params, fit$ppcc)
  }
))
colnames(fit_summary_std) <- fund_names
rownames(fit_summary_std) <- c("m", "s", "nu", "PPCC")
```

```{r eval=run_individual}
fit_summary_norm <- as.data.frame(lapply(
  fits_norm,
  function(fit) {
     c(fit$dist_params, fit$ppcc)
  }
))
colnames(fit_summary_norm) <- fund_names
rownames(fit_summary_norm) <- c("m", "s", "PPCC")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_l <- as.data.frame(lapply(
  fits_l,
  function(fit) {
     c(fit$dist_params, fit$ppcc)
  }
))
colnames(fit_summary_l) <- fund_names_l
rownames(fit_summary_l) <- c("m", "s", "nu", "xi", "PPCC")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_std_l <- as.data.frame(lapply(
  fits_std_l,
  function(fit) {
     c(fit$dist_params, fit$ppcc)
  }
))
colnames(fit_summary_std_l) <- fund_names_l
rownames(fit_summary_std_l) <- c("m", "s", "nu", "PPCC")
```

```{r eval=(run_individual && params$include_long)}
fit_summary_norm_l <- as.data.frame(lapply(
  fits_norm_l,
  function(fit) {
     c(fit$dist_params, fit$ppcc)
  }
))
colnames(fit_summary_norm_l) <- fund_names_l
rownames(fit_summary_norm_l) <- c("m", "s", "PPCC")
```


```{r eval=(params$run_comparison && params$run_sim)}
kappa_df <- data.frame(matrix(numeric(length(fund_names)), ncol = length(fund_names)))

for(i in seq_along(fits)) {
  kappa_df[1, i] <- f_kappa(
    n0 = 1, 
    n = 30, 
    mean = fits[[i]]$dist_params[1], 
    sd = fits[[i]]$dist_params[2], 
    nu = fits[[i]]$dist_params[3], 
    xi = fits[[i]]$dist_params[4], 
    num_sim = 1e5
  )
}

colnames(kappa_df) <- fund_names
```

```{r eval=(params$run_comparison && params$run_sim)}
saveRDS(kappa_df, file = here("data", data_dir,  "kappa_df.RData"))
```

```{r eval=(params$run_comparison && !params$run_sim)}
kappa_df <- readRDS(file = here("data", data_dir,  "kappa_df.RData"))
```

```{r eval=(params$run_comparison && params$run_sim)}
n_min_df <- data.frame(matrix(numeric(length(fund_names)), ncol = length(fund_names)))

for(i in seq_along(fits)) {
  n_min_df[1, i] <- f_n_min(
    n_g = 30, 
    mean = fits[[i]]$dist_params[1], 
    sd = fits[[i]]$dist_params[2], 
    nu = fits[[i]]$dist_params[3], 
    xi = fits[[i]]$dist_params[4], 
    num_sim = 1e5, 
    approx = FALSE
  )
}

colnames(n_min_df) <- fund_names
```

```{r eval=(params$run_comparison && params$run_sim)}
saveRDS(n_min_df, file = here("data", data_dir,  "n_min_df.RData"))
```

```{r eval=(params$run_comparison && !params$run_sim)}
n_min_df <- readRDS(file = here("data", data_dir,  "n_min_df.RData"))
```


```{r chunk16}
##                                     MONTE CARLO
```

```{r chunk17, eval=(params$run_sim && run_individual)}
num_fits <- length(fits)

mc_output <- list()
for(i in seq_along(fits)) {
  message(paste0("\nsstd: ", i, " of ", num_fits, "\n"))
  mc_output[[i]] <- mc_simulation(
    fits[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "sstd"
  )
}

mc_output_std <- list()
for(i in seq_along(fits_std)) {
  message(paste0("\nstd: ", i, " of, ", num_fits, "\n"))
  mc_output_std[[i]] <- mc_simulation(
    fits_std[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "std"
  )
}

mc_output_norm <- list()
for(i in seq_along(fits_norm)) {
  message(paste0("\nnormal: ", i, " of, ", num_fits, "\n"))
  mc_output_norm[[i]] <- mc_simulation(
    fits_norm[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "normal"
  )
}
```

```{r chunk17a, eval=(params$run_sim && run_individual && params$include_long)}
fit_id_l <- list(
  1
)

num_fits_l <- length(fits_l)

mc_output_l <- list()
for(i in seq_along(fits_l)) {
  message(paste0("\nsstd: ", i, " of ", num_fits_l, "\n"))
  mc_output_l[[i]] <- mc_simulation(
    fits_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "sstd"
  )
}

mc_output_std_l <- list()
for(i in seq_along(fits_std_l)) {
  message(paste0("\nstd: ", i, " of, ", num_fits, "\n"))
  mc_output_std_l[[i]] <- mc_simulation(
    fits_std_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "std"
  )
}

mc_output_norm_l <- list()
for(i in seq_along(fits_norm_l)) {
  message(paste0("\nnormal: ", i, " of, ", num_fits, "\n"))
  mc_output_norm_l[[i]] <- mc_simulation(
    fits_norm_l[fit_id_l[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold,
    distribution = "normal"
  )
}
```


```{r chunk17b, eval=(params$run_sim && run_individual)}
saveRDS(mc_output, file = here("data", data_dir, "mc_output.RData"))
saveRDS(mc_output_std, file = here("data", data_dir, "mc_output_std.RData"))
saveRDS(mc_output_norm, file = here("data", data_dir, "mc_output_norm.RData"))
```

```{r chunk17c, eval=(params$run_sim && run_individual && params$include_long)}
saveRDS(mc_output_l, file = here("data", data_dir, "mc_output_l.RData"))
saveRDS(mc_output_std_l, file = here("data", data_dir, "mc_output_std_l.RData"))
saveRDS(mc_output_norm_l, file = here("data", data_dir, "mc_output_norm_l.RData"))
```

```{r chunk18, eval=(run_individual && !params$run_sim)}
mc_output <- readRDS(file = here("data", data_dir, "mc_output.RData"))
mc_output_std <- readRDS(file = here("data", data_dir, "mc_output_std.RData"))
mc_output_norm <- readRDS(file = here("data", data_dir, "mc_output_norm.RData"))
```

```{r chunk18b, eval=(run_individual && !params$run_sim && params$include_long)}
mc_output_l <- readRDS(file = here("data", data_dir, "mc_output_l.RData"))
mc_output_std_l <- readRDS(file = here("data", data_dir, "mc_output_std_l.RData"))
mc_output_norm_l<- readRDS(file = here("data", data_dir, "mc_output_norm_l.RData"))
```


```{r chunk19}
##                                     MAX SUM PLOTS
```

```{r chunk20, eval=(params$run_sim && run_individual)}
max_sum_plots <- list()
for(i in seq_along(fits)) {
  max_sum_plots[[i]] <- plot_max_sum(
    fits[fit_id[[i]]], 
    mc_num_paths,
    distribution = "sstd"
  )
}

max_sum_plots_std <- list()
for(i in seq_along(fits_std)) {
  max_sum_plots_std[[i]] <- plot_max_sum(
    fits_std[fit_id[[i]]], 
    mc_num_paths, 
    distribution = "std"
  )
}

max_sum_plots_norm <- list()
for(i in seq_along(fits_norm)) {
  max_sum_plots_norm[[i]] <- plot_max_sum(
    fits_norm[fit_id[[i]]], 
    mc_num_paths, 
    distribution = "norm"
  )
}
```

```{r chunk20a, eval=(params$run_sim && run_individual && params$include_long)}
max_sum_plots_l <- list()
for(i in seq_along(fits_l)) {
  max_sum_plots_l[[i]] <- plot_max_sum(
    fits_l[fit_id_l[[i]]], 
    mc_num_paths,
    distribution = "sstd"
  )
}

max_sum_plots_std_l <- list()
for(i in seq_along(fits_std_l)) {
  max_sum_plots_std_l[[i]] <- plot_max_sum(
    fits_std_l[fit_id_l[[i]]], 
    mc_num_paths, 
    distribution = "std"
  )
}

max_sum_plots_norm_l <- list()
for(i in seq_along(fits_norm_l)) {
  max_sum_plots_norm_l[[i]] <- plot_max_sum(
    fits_norm_l[fit_id_l[[i]]], 
    mc_num_paths, 
    distribution = "norm"
  )
}
```

```{r chunk20b, eval=(params$run_sim && run_individual)}
saveRDS(max_sum_plots, here("data", data_dir, "max_sum_plots.RData"))
saveRDS(max_sum_plots_std, here("data", data_dir, "max_sum_plots_std.RData"))
saveRDS(max_sum_plots_norm, here("data", data_dir, "max_sum_plots_norm.RData"))
```

```{r chunk20c, eval=(params$run_sim && run_individual && params$include_long)}
saveRDS(max_sum_plots_l, here("data", data_dir, "max_sum_plots_l.RData"))
saveRDS(max_sum_plots_std_l, here("data", data_dir, "max_sum_plots_std_l.RData"))
saveRDS(max_sum_plots_norm_l, here("data", data_dir, "max_sum_plots_norm_l.RData"))
```



```{r chunk21, eval=(run_individual && !params$run_sim)}
max_sum_plots <- readRDS(file = here("data", data_dir, "max_sum_plots.RData"))
max_sum_plots_std <- readRDS(file = here("data", data_dir, "max_sum_plots_std.RData"))
max_sum_plots_norm <- readRDS(file = here("data", data_dir, "max_sum_plots_norm.RData"))
```

```{r chunk21b, eval=(run_individual && !params$run_sim)}
max_sum_plots_l <- readRDS(file = here("data", data_dir, "max_sum_plots_l.RData"))
max_sum_plots_std_l <- readRDS(file = here("data", data_dir, "max_sum_plots_std_l.RData"))
max_sum_plots_norm_l <- readRDS(file = here("data", data_dir, "max_sum_plots_norm_l.RData"))
```


```{r chunk22}
##                                  IMPORTANCE SAMPLING
```

```{r chunk23, eval=(params$run_sim && run_individual && params$run_is_sim)}
is_g_fits <-  list()
is_output <- list()
for(i in seq_along(fits)) {
  is_g_fits[[i]] <- is_proposal(
    x_i_fit = fits[fit_id[[i]]],
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    obj_func_plot = TRUE,
    init_par = c(2, 0.5),
    method = "L-BFGS-B",
    lower = c(1, 0.1),
    upper = c(3, 0.9)
  )
  
  is_output[[i]] <- importance_sampling(
    x_i_fit = fits[fit_id[[i]]], 
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    g_n_params = is_g_fits[[i]]$par, 
    mode = 1
  )
}
```

```{r chunk23b, eval=(params$run_sim && run_individual && params$run_is_sim)}
saveRDS(is_g_fits, here("data", data_dir, "is_g_fits.RData"))
saveRDS(is_output, here("data", data_dir, "is_output.RData"))
```

```{r chunk24, eval=(run_individual && !params$run_sim && !params$run_is_sim && params$run_is_plot)}
is_g_fits <- readRDS(file = here("data", data_dir, "is_g_fits.RData"))
is_output <- readRDS(file = here("data", data_dir, "is_output.RData"))
```


```{r chunk25}
###########################################################################
##
##                            GENERATE REPORTS
##
###########################################################################
```

```{r chunk26, eval=params$run_exploratory, results='asis'}
exploratory_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-exploratory.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(exploratory_report), sep = '\n')
```

```{r chunk27, eval=params$run_comparison, results='asis'}
comparison_report <- knitr::knit_child(
    here("child_docs", "pension-returns_template-comparison.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(comparison_report), sep = '\n')
```

```{r chunk28, eval=params$run_individual, results='asis'}
long <- FALSE
individual_reports <- lapply(
  seq_along(data_df),
  function(i) {
    knitr::knit_child(
      here("child_docs", "pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```

```{r chunk28b, eval=params$run_individual, results='asis', eval=params$include_long}
long <- TRUE
individual_reports <- lapply(
  seq_along(data_df_l),
  function(i) {
    knitr::knit_child(
      here("child_docs", "pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```


# Findings

With the data deliberately frozen, we can interpret a fixed set of estimates rather
than chase a moving target. We present the salient results, analyse them against the
theory that motivates the study, and draw out the implications.

## 1. The plans are one portfolio held at different leverage

*Result.* Within each provider the high-risk plan is, to a close approximation, a scalar
multiple of the medium-risk plan: Regressing one cumulative log-path on the other gives a
slope of about 1.2 (Velliv) and 1.5 (PFA) with intercept indistinguishable from zero and
R^2 between 0.91 and 0.98. Across all four plans the first principal component accounts for
roughly 96% of the variation. The implied low-risk building block (2*B - D for PFA)
reconstructs a coherent, lower-volatility series, confirming a two-fund structure.

*Analysis.* This is not a statistical accident but the product design: PFA states that
profile B is approximately a 50/50 blend of its high- and low-risk funds and profile D is
100% the high-risk fund, so the profiles are fixed affine combinations of the same two
underlying funds. The choice of "risk profile" within a provider is therefore a choice of
exposure (leverage) to one portfolio, not a choice between different portfolios.

*Implication.* The investor's first decision reduces to *how
much* of one factor to hold.

## 2. "Risk" is amplified drawdown, and apparent path-dominance is a starting-point artefact

*Result.* Because the slope exceeds 1 with zero intercept, the high-risk plan amplifies the
common factor symmetrically: higher average growth, but deeper drawdowns (worst monthly
peak-to-trough -17.3% vs -15.2% for Velliv, -15.9% vs -11.9% for PFA) and underperformance
in flat or falling markets. The much-noted fact that the cumulative high-risk path never
dips below the medium-risk path holds only for an index started in 2012; rebased to just
before the COVID drawdown it falls below for about nine months (see "Path crossing").

*Analysis.* With W_high ≈ W_med^β and β>1, the high path lies below the medium path exactly
when the cumulative return since the start is negative. Non-crossing is thus a property of
the start date relative to the cycle, not of the plans.

## 3. How much can a fixed sample tell us? Parameter precision

This is the methodological core. We fit a four-parameter Fernández–Steel skewed Student-t
(location m, scale s, tail index nu, skew xi). The parameters split sharply by how much
data they need.

*The location and scale are estimable.* Monthly, m̂(vmr) = 0.0054 with standard error
0.0021 (t ≈ 2.6); the MLE is about as efficient as the Gaussian sample mean (naive σ/√n =
0.00225). Under a null of Gaussian(0, 0.02) the chance of observing a mean this large is
~0.02%, so the positive drift is real. The scale is similarly well determined.

*The tail and skew are the hard parameters — and they carry the risk.*
- Annual (n = 13): the observed Fisher information is singular for vmr, vhr and phr — the
  optimiser drives xi to ≈ 0 (vmr xî = 0.004), a degenerate extreme-left-skew corner, and
  standard errors are undefined. For the one non-degenerate fit, pmr, nu^ = 2.27 with
  SE = 2.68 (118% relative): nu is not identified.
- Monthly (n = 142): the fits are non-degenerate. nu^ = 3.4–4.3 with SE = 1.2–1.8 (35–43%
  relative); the 95% CI on nu for vmr is ≈ [1.1, 5.8], spanning infinite variance (nu < 2),
  infinite kurtosis (nu < 4) and near-Gaussian tails. xî = 0.70–0.77 with SE ≈ 0.085
  (10–12%); the CI [0.53, 0.87] lies wholly below 1 (≈ 3.5 SE from symmetry). The left skew
  is statistically established; the tail index is not pinned down.

In short: the *direction* (heavy left skew, fat tails) is robust where we have power; the
*magnitude* of the tail is not estimable annually and only loosely estimable monthly. (A
Monte-Carlo replication of this, simulated at pmr's parameters, is in `misc/sgt.Rmd`, where
the nu and xi estimators show very wide dispersion at small n.)

*Analysis — why nu and xi are hard while m is easy.*
- The Gaussian has only m and s; the t adds nu and xi, which govern the tail and have no
  Gaussian counterpart. The data burden lives entirely in these extra parameters.
- Taleb's recommended workflow (SCoFT, "Consequence 7", and the plug-in discussion) is to
  estimate the tail exponent by MLE — which "can work well… inverse-gamma distributed, low
  variance" — and then *derive* the mean analytically ("plug-in"), because the tail
  exponent "captures by extrapolation the low-probability deviations" better than the
  sample mean does. The analytical moment maps (Li & Nadarajah; Fernández–Steel,
  implemented in `misc/sgt.Rmd`) are precisely this plug-in. The catch is that plug-in
  presupposes an estimable nu; at n = 13 we have shown it is not, so the route Taleb
  recommends is unavailable annually and only partly available monthly.
- The κ metric (Taleb, Def. 8.1; Table 8.1) measures the preasymptotic deficit in the
  convergence of the *average*: κ = 0 for the Gaussian, and for Student-T(3) the table
  gives κ₁ = 0.29, κ₁,₃₀ = 0.19, κ₁,₁₀₀ = 0.16 (reproduced by our `f_kappa`). With
  n_min = n_g^(1/(1−κ)), matching 30 Gaussian observations takes about 120 Student-T(3)
  observations using κ₁ (Taleb), or ~67 using κ₁,₃₀; the figure is range-dependent because
  κ itself drifts with n — a symptom of the preasymptotics, not an inconsistency. Crucially,
  κ concerns the *mean*; the *tail* parameters are harder still, so n_min for the mean is a
  floor on what the tail would need.
- A caution on goodness-of-fit: Taleb notes that under thick tails the in-sample R² is
  *inflated* and, when variance is infinite, meaningless. We therefore report the
  probability-plot correlation as **PPCC** (the correlation itself, not its square) and read
  it, like AIC/BIC, as a statement about the *body* of the distribution -- a high PPCC does
  **not** validate the tail. For the tail we rely on the tail-weighted Anderson-Darling test
  (with a parametric-bootstrap p-value, since the parameters are estimated) and, more
  fundamentally, the max-sum plots, which ask whether the relevant moments exist at all.

## 4. Finite vs. infinite variance

*Result.* Some fits return nu ≈ 2.0–2.3 (pmr, phr), at or below the threshold where the
variance ceases to exist; the symmetric-t Monte Carlo for phr diverges to absurd values
(mean ~10^7). Monthly, the nu confidence interval still includes nu < 2.

*Analysis.* This is not a numerical error: for nu ≤ 2 the sample standard deviation does
not converge and E[100·e^X] is undefined ("the variance of a finite-variance variable with
tail exponent < 4 will be infinite"). We therefore cannot reject infinite variance even
with 142 monthly observations.

## 5. A precautionary reading

*The asymmetry of the inference.* The fits are heavy-left-skewed with a small positive
mean. Such a distribution predicts losses worse than any yet observed; the benign,
low-interest, bull-market sample has simply not delivered them. One could object that more
data might instead reveal large upside, symmetrising the fit and thinning the tail. Two
considerations weigh against relying on that hope:

- Taleb's escape — "better stay home" when no reliable estimator exists — is unavailable
  here: participation in a mandatory pension is compulsory. His own qualifier is that one
  "can still take risky decisions if bounded for maximum losses"; a pension is *not* bounded
  that way, which strengthens, rather than weakens, the case for caution.
- The data were collected in a single low-rate, bull-market regime, so the sample is more
  likely to understate than overstate tail losses.

*A speculative but coherent mechanism.* The observed shape — left skew with a positive mean
— is what a return-*capping* design would produce. Capping the upside is self-evidently
unappealing, yet it is a coherent way to engineer a positive *expected* return: trimming
large gains while leaving the occasional large loss to be booked as "chance" pushes the mean
up. We cannot establish design from the data, but the shape is consistent with it, and
prudence argues for taking the heavy left skew at face value until evidence says otherwise.

*The cost of this precaution.* Assuming a heavier left tail also *raises* the apparent value
of diversifying across the two providers (the tail-softening from holding both is larger the
fatter the assumed idiosyncratic tail). So precaution here is not free: it may lead us to
over-value diversification. That trade-off is properly weighed in the investor-facing
analysis, not in this report.

## 6. Bottom line

Qualitatively the conclusions are firm: the plans are leverage on one portfolio, the tails
are fat, drawdowns are asymmetric, and path-dominance is a starting-point artefact.
Quantitatively, the mean and scale are estimable (firmly monthly, loosely annually) and the
left-skew direction is established monthly, but the tail index nu is not pinned down even at
n = 142 and is unidentified at n = 13. Under a precautionary principle — and because
"staying home" is not an option — we treat the distribution as heavily left-skewed, while
flagging that this stance may over-state the benefit of diversifying across providers, a
question we take up separately.

# Appendix

## Infinite variance

Taleb, Statistical Consequences Of Fat Tails, p. 97:  
"the variance of a finite variance random variable with tail exponent $< 4$ will be infinite".

And p. 363:  
"The hedging errors for an option portfolio (under a daily revision regime) over 3000 days, under a constant volatility Student T with tail exponent $\alpha = 3$. Technically the errors should not converge in finite time as their distribution has infinite variance."


## QQ lines  

Note: QQ lines by design pass through 1st and 3rd quantiles. They are not trendlines in the sense of linear regression.  

## Arithmetic vs geometric mean
Let $m$ be the number of steps in each path and $n$ be the number of paths.
$a$ is the initial capital.
Use arithmetic mean for mean of all paths at time $t$:
$$\dfrac{a (e^{z_1} + e^{z_2} + \dots + e^{z_n})}{n}$$
where
$$z_i := x_{i, 1} + x_{i, 2} + \dots + x_{i, m}$$
Use geometric mean for mean of all steps in a single path $i$:
$$a e^{\frac{x_{i, 1} + x_{i, 2} + \dots + x_{i, m}}{m}} = a \sqrt[m]{e^{x_{i, 1} + x_{i, 2} + \dots + x_{i, m}}}$$

So for **Monte Carlo** of returns after $m$ periods, we 

+ fit a skewed t-distribution to log-returns and use that distribution to simulate $\{x_{i, j}\}_j^m$,
+ for each path $i$, calculate $100\cdot e^{z_i}$,
+ calculate the mean of $\{z_i\}_i^n$:
    + $$\bar{z} = 100\dfrac{e^{z_1} + e^{z_2} + \dots + e^{z_n}}{n}$$

For **Importance Sampling**, we

+ model log-returns on a skewed t-distribution,
+ for each path $i$, calculate $100\cdot e^{z_i}$,
+ fit a skewed t-distribution to $\{z_i\}_i^n$ and use it as our $f$ density function from which we simulate $\{h_i\}_i^n$,
    + In our case $h$ and $z$ are identical, because we have an idea for a distribution to simulate $z$, but in general for IS $h$ could be a function of $z$.
+ calculate $w* = \frac{f}{g^*}$, where $g*$ is our proposal distribution, which minimizes the variance of $h\cdot w$.
+ calculate the arithmetic mean of $\{h_i w_i^{*}\}_i^n$: 
    + $$100 \dfrac{e^{h_1 w_1^{*}} +  e^{h_2 w_2^{*}} + \dots +  e^{h_n w_n^{*}}}{n}$$


## Average of returns vs returns of average

### Math

$$\text{Avg. of returns} := \dfrac{ \left(\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} \right) }{2}$$
$$\text{Returns of avg.} := \left(\dfrac{ x_t + y_t }{2}\right) \Big/ \left(\dfrac{ x_{t-1} + y_{t-1} }{2}\right) \equiv \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

For which $x_1$ and $y_1$ are
$\text{Avg. of returns} = \text{Returns of avg.}$?

$$\dfrac{ \left(\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} \right) }{2} = \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

$$\dfrac{x_t}{x_{t-1}} + \dfrac{y_t}{y_{t-1}} = 2 \dfrac{ x_t + y_t }{ x_{t-1} + y_{t-1}}$$

$$(x_{t-1} + y_{t-1}) x_t y_{t-1} + (x_{t-1} + y_{t-1}) x_{t-1} y_t = 2 (x_{t-1}y_{t-1}x_t + x_{t-1}y_{t-1}y_t)$$

$$(x_{t-1}x_ty_{t-1} + y_{t-1}x_ty_{t-1}) + (x_{t-1}x_{t-1}y_t + x_{t-1}y_{t-1}y_t) = 2(x_{t-1}y_{t-1}x_t + x_{t-1}y_{t-1}y_t)$$
This is not generally true, but true if for instance
$x_{t-1} = y_{t-1}$.

### Example

```{r}
x0 <- 100
y0 <- 200
Rx <- 0.5
Ry <- 1.5
```

Definition: `R = 1+r`

```{r}
cat(paste0("Let x_0 be ", x0, ".\n"))
cat(paste0("Let y_0 be ", y0, ".\n"))
cat("So the initial value of the pf is", x0 + y0, ".\n")
cat("\n")
cat(paste0("Let R_x be ", Rx, ".\n"))
cat(paste0("Let R_y be ", Ry, ".\n"))
```

Then,

```{r}
cat(paste0("x_1 is R_x * x_0 = ", Rx * x0, ".\n"))
cat(paste0("y_1 is R_y * y_0 = ", Ry * y0, ".\n"))
```

Average of returns:

```{r}
cat("0.5 * (R_x + R_y) =", 0.5 * (Rx + Ry), "\n")
```

So here the value of the pf at t=1 should be unchanged from t=0:

```{r}
cat("(x_0 + y_0) * 0.5 * (R_x + R_y) =", (x0 + y0) * 0.5 * (Rx + Ry), "\n")
```

But this is clearly not the case:

```{r}
cat("0.5 * (x_1 + y_1) = 0.5 * (R_x * x_0 + R_y * y_0) =", 0.5 * (Rx * x0 + Ry * y0), "\n")
```

Therefore we should take returns of average, not average of returns!

Let's take the average of log returns instead:

```{r}
cat("0.5 * (log(R_x) + log(R_y)) =", 0.5 * (log(Rx) + log(Ry)), "\n")
```

We now get:

```{r}
cat("(x_0 + y_0) * exp(0.5 * (log(Rx) + log(Ry))) =", (x0 + y0) * exp(0.5 * (log(Rx) + log(Ry))), "\n")
```

So taking the average of log returns doesn't work either.

## Simulation of mix vs mix of simulations

Test if a simulation of a mix (average) of two returns series has the
same distribution as a mix of two simulated returns series.

```{r}
mc_sim <- function(
    num_runs = 1, 
    num_paths = 1000, 
    num_periods = 20,
    m_a = 0,
    s_a = 0.4,
    m_b = 10,
    s_b = 3
    ) {
  
  data_x <- rnorm(num_periods, m_a, s_a)
  data_y <- rnorm(num_periods, m_b, s_b)
  
  m_data_x <- mean(data_x)
  s_data_x <- sd(data_x)
  m_data_y <- mean(data_y)
  s_data_y <- sd(data_y)
  
  cat("m(data_x):", m_data_x, "\n")
  cat("s(data_x):", s_data_x, "\n")
  cat("m(data_y):", m_data_y, "\n")
  cat("s(data_y):", s_data_y, "\n")
  cat("\n")
  
  m_data_xy <- mean(0.5 * data_x + 0.5 * data_y)
  s_data_xy <- sd(0.5 * data_x + 0.5 * data_y)
  
  cat("m(data_x + data_y):", m_data_xy, "\n")
  cat("s(data_x + data_y):", s_data_xy, "\n")
  cat("\n")
  
  run_sim <- function(num_runs) {
    df <- data.frame(
      m_a = rep(0, num_runs), m_b = rep(0, num_runs), 
      s_a = rep(0, num_runs), s_b = rep(0, num_runs)
    )
    for(j in 1:num_runs) {
      sim_x <- rep(0, num_paths)
      sim_y <- rep(0, num_paths)
      sim_xy <- rep(0, num_paths)
      for(i in 1:num_paths) {
        sim_x[i] <- sum(rnorm(num_periods, m_data_x, s_data_x))
        sim_y[i] <- sum(rnorm(num_periods, m_data_y, s_data_y))
        sim_xy[i] <- sum(rnorm(num_periods, m_data_xy, s_data_xy))
      }

      df$m_a[j] <-  mean(0.5 * sim_x + 0.5 * sim_y)
      df$m_b[j] <-  mean(sim_xy)
      df$s_a[j] <-  sd(0.5 * sim_x + 0.5 * sim_y)
      df$s_b[j] <-  sd(sim_xy)
    }
    df
  }
  
  run_sim(num_runs)
}
```

```{r}
mc_sim_df <- mc_sim(
  num_runs = 10, 
  num_paths = 1000, 
  num_periods = 20,
  m_a = 0,
  s_a = 0.4,
  m_b = 10,
  s_b = 3
)
```

m and s of final state of all paths.\
`_a` is mix of simulated returns.\
`_b` is simulated mixed returns.

```{r}
knitr::kable(mc_sim_df, digits = 3)
```

```{r}
summary(mc_sim_df)
```

`_a` and `_b` are very close to equal.\
We attribute the differences to differences in estimating the
distributions in version a and b.

The final state is independent of the order of the preceding steps:

```{r}
vect1 <- c(rnorm(100))
vect2 <- c(sample(vect1, 100))
vect3 <- c(sample(vect1, 100))
path1 <- c(0, cumsum(vect1))
path2 <- c(0, cumsum(vect2))
path3 <- c(0, cumsum(vect3))
plot(path1, type = "l", col = "blue", 
     ylim = c(
       min(c(path1, path2, path3)), 
       max(c(path1, path2, path3))
     )
)
lines(path2, col = "red")
lines(path3, col = "green")
```

So does the order of the steps in the two processes matter, when mixing
simulated returns?

```{r}
vect1a <- c(rnorm(100, 0.05, 0.06))
vect1b <- c(sample(vect1a, 100))
vect2a <- c(rnorm(100, 0.05, 0.06))
vect2b <- c(sample(vect2a, 100))

path1a <- 100 * c(1, cumprod(1 + vect1a))
path1b <- 100 * c(1, cumprod(1 + vect1b))
path2a <- 100 * c(1, cumprod(1 + vect2a))
path2b <- 100 * c(1, cumprod(1 + vect2b))

mix_path_a <- 0.5 * path1a + 0.5 * path2a
mix_path_b <- 0.5 * path1b + 0.5 * path2b

plot(path1a, type = "l", lty = 1, col = "blue", 
     ylim = c(
       min(c(path1a, path1b, path2a, path2b)), 
       max(c(path1a, path1b, path2a, path2b))
     )
)
lines(path1b, lty = 2, col = "blue")
lines(path2a, lty = 1, col = "red")
lines(path2b, lty = 2, col = "red")
```

```{r}
plot(mix_path_a, type = "l", lty = 1, col = "blue", 
     ylim = c(
       min(c(mix_path_a, mix_path_b)), 
       max(c(mix_path_a, mix_path_b))
     )
)
lines(mix_path_b, lty = 2, col = "blue")
```

The order of steps in the individual paths do not matter, because the
mix of simulated paths is a sum of a sum, so the order of terms doesn't
affect the sum. If there is variation it is because the sets preceding
steps are not the same. For instance, the steps between step 1 and 60 in
the plot above are not the same for the two lines.

Recall,
$$\text{Var}(aX+bY) = a^2 \text{Var}(X) + b^2 \text{Var}(Y) + 2ab \text{Cov}(a, b)$$

```{r include=TRUE, echo=TRUE}
var(0.5 * data_df$vhr + 0.5 * data_df$phr)
0.5^2 * var(data_df$vhr) + 0.5^2 * var(data_df$phr) + 2 * 0.5 * 0.5 * cov(data_df$vhr, data_df$phr)
```

Our distribution estimate is based on 142 monthly observations. Is that
enough for a robust estimate? What if we suddenly hit a month like the
2008 crash? How would that affect our estimate?\
Let's try to include the Velliv data going back to 2007, which spans the GFC.\
We do this by sampling 142 observations from the long series `vmrl`
(`data_df_l$vmrl`, 207 monthly observations incl. 2007-2012).

```{r}
n <- 50
test_df <- data.frame(m = rep(0, n), s = rep(0, n))
for(i in 1:n) {
  vmrl_smp <- sample(data_df_l$vmrl, 142)
  test_df$m[i] <- mean(0.5 * vmrl_smp + 0.5 * data_df$phr)
  test_df$s[i] <- sd(0.5 * vmrl_smp + 0.5 * data_df$phr)
}
summary(test_df)
```

## The meaning of `xi`

The fit for `mhr` has the highest `xi` value of all. This suggests
right-skew:

```{r}
plot((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 1), xlim = c(-20, 20), ylim = c(0, 0.1), type = "l", xlab = "value", ylab = "likelihood", main = "Skew t-distribution density")
lines((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 0.5), col="red")
lines((-20000:20000)/1000, dsstd((-20000:20000)/1000, 0, 7, 5, 2), col="blue")
legend("topright", c("xi=1", "xi=0.5", "xi=2"), col = c("black", "red", "blue"), lty = 1)
```

## Max vs sum plot

If the Law Of Large Numbers holds true,
$$\dfrac{\max (X_1^p, ..., X^p)}{\sum_{i=1}^n X_i^p} \rightarrow 0$$ for
$n \rightarrow \infty$.

If not, $X$ doesn't have a $p$'th moment.

See Taleb: The Statistical Consequences Of Fat Tails, p. 192


