---
title: "Pension returns analysis"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
   - \usepackage[fontsize=8pt]{scrextend}
mainfont: SourceSansPro
output: 
  html_document:
    toc: true
    toc_depth: 3
    keep_md: yes
  pdf_document:
    toc: true
    toc_depth: 3
#fontsize: 10pt # for pdf. Limited to 10pt, 11pt and 12pt. Else use scrextend.
params:
  run_sim: TRUE ## TRUE: Run simulations and write output. FALSE: Read saved
                ## simulations output from disk instead of running the simulations.
  run_exploratory: TRUE ## Include exploratory report?
  run_individual: TRUE ## Include individual reports? Depends on run_individual.
  run_comparison: TRUE ## Include comparison report? Depends on run_individual.
  run_comments: TRUE
  run_appendix: TRUE
date: "`r format(Sys.time(), '%H:%M %d %B %Y')`"
---

```{r include=FALSE}
## ****************************** WARNING ******************************
## 
##             This document takes a long time to render!
##
## *********************************************************************
##
## To read previous simulation outputs from disk rather than run simulations,
## set YAML parameter run_sim: TRUE
##
## NOTE:
## When using the "Run All" command, YAML parameters will be ignored. An error
## message "Error in eval(ele) : object 'params' not found" will be printed,
## but the chunks will be evaluated.
```


```{r include=FALSE, eval=FALSE}
################################################################################
##
##                                      SETUP
##
################################################################################
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

run_individual <- (params$run_individual + params$run_comparison > 0)
run_exploratory <- (params$run_exploratory + run_individual > 0)

## Note: In the second line (run_exploratory <- ...), run_individual is correct that - not params$run_individual, because if we choose params$run_exploratory=TRUE and params$run_comparison=TRUE, but params$run_individual=FALSE, we still have to do the individual calculations.
```

```{r}
library(here)
```

```{r}
source(here("pension-returns_functions.R"))
```

```{r eval=FALSE, include=FALSE}
## Render this document to html and pdf
rmarkdown::render(here("pension-returns.Rmd"), output_format ="all")
```

```{bash eval=FALSE, include=FALSE}
## Convert md to pdf
## Run manually:
pandoc pension-returns.md -o pension-returns.pdf
```



```{r}
################################################################################
##
##                                      INPUTS
##
################################################################################
```

```{r}
mc_num_paths <- 1000
mc_num_periods <- 20
mc_dao <- TRUE
mc_threshold <- 0.01
```

Fit log returns to F-S skew standardized Student-t distribution.  
`m`  is the location parameter.  
`s` is the scale parameter.  
`nu` is the estimated shape parameter (degrees of freedom).  
`xi` is the estimated skewness parameter.  

# Returns data 2011-2023.  
For 2011, medium risk data is used in the high risk data set, as no high risk fund data is available prior to 2012.  
`vmrl` is a long version of Velliv medium risk data, from 2007 to 2023. For 2007 to 2011 (both included) no high risk data is available.
```{r}
## Headlines for individual sections
headlines <- c(
  "Velliv medium risk (vmr), 2011 - 2023",
  "Velliv medium risk (vmrl), 2007 - 2023",
  "Velliv high risk (vhr), 2011 - 2023",
  "PFA medium risk (pmr), 2011 - 2023",
  "PFA high risk (phr), 2011 - 2023",
  "Mix medium risk (mmr), 2011 - 2023",
  "Mix high risk (mhr), 2011 - 2023",
  "Mix vmr+phr (vm_ph), 2011 - 2023",
  "Mix vhr+pmr (mh_pm), 2011 - 2023"
)

## Log returns data 2011-2023.
## Velliv medium risk returns
vmr <- log(c(0.986, 1.098, 1.157, 1.136, 1.058, 1.044, 1.100, 0.963, 1.168, 1.097, 1.141, 0.868, 1.094))

## Velliv medium risk return, long period: 2007-2023
vmrl <- log(c(1.058, 0.801, 1.193, 1.129, 1.035, 0.996, 1.133, 1.092, 1.085, 1.013, 1.095, 1.011, 1.129, 1.101, 1.128, 0.988, 1.048))

## PFA medium risk returns
pmr <- log(c(1.084, 0.904, 1.107, 1.042, 1.119, 0.976, 1.089, 1.073, 1.084, 1.107, 1.111, 1.141, 1.004))

## Mix medium risk return
mmr <- log(c(1.035, 0.996, 1.133, 1.092, 1.085, 1.013, 1.095, 1.011, 1.129, 1.101, 1.128, 0.988, 1.048))

## Velliv high risk returns
vhr <- log(c(0.986, 1.093, 1.214, 1.160, 1.075, 1.039, 1.118, 0.952, 1.205, 1.099, 1.194, 0.849, 1.122))

## PFA high risk return
phr <- log(c(1.159, 0.878, 1.200, 1.068, 1.190, 0.943, 1.135, 1.082, 1.123, 1.128, 1.208, 1.182, 0.934))
  
## Mix high risk return
mhr <- log(c(1.073, 0.977, 1.207, 1.116, 1.128, 0.992, 1.126, 1.013, 1.164, 1.113, 1.201, 1.012, 1.014))

vmr_phr <- (c(100, 100 * cumprod(exp(vmr))) + c(100, 100 * cumprod(exp(phr))))/2
vmr_phr <- log(tail(vmr_phr, -1)/head(vmr_phr, -1))

vhr_pmr <- (c(100, 100 * cumprod(exp(vhr))) + c(100, 100 * cumprod(exp(pmr))))/2
vhr_pmr <- log(tail(vhr_pmr, -1)/head(vhr_pmr, -1))

data_df <- data.frame(
  vmr = vmr,
  vhr = vhr,
  pmr = pmr,
  phr = phr,
  mmr = mmr,
  mhr = mhr,
  vmr_phr = vmr_phr,
  vhr_pmr = vhr_pmr
)

data_df_l <- data.frame(
  vmrl = vmrl
)

fund_names <- colnames(data_df)
```

```{r}
all_data <- exp(c(vmr, vhr, pmr, phr, mmr, mhr))
plot(x = c(2011:2023), y = exp(vmr), type = "l",  lty = 1, lwd = 2, xlab = "year", ylab = "Gross returns", col = "#76EEC6", ylim = c(min(all_data), max(all_data)), main = "Gross returns 2011-2023")
lines(x = c(2011:2023), y = exp(vhr), col = "#76EEC6",  lty = 2, lwd = 1.5)
lines(x = c(2011:2023), y = exp(pmr), col = "#CD6600",  lty = 1, lwd = 2)
lines(x = c(2011:2023), y = exp(phr), col = "#CD6600",  lty = 2, lwd = 1.5)
lines(x = c(2011:2023), y = exp(mmr), col = "#FF00FF",  lty = 1, lwd = 2)
lines(x = c(2011:2023), y = exp(mhr), col = "#FF00FF",  lty = 2, lwd = 1.5)
legend("bottom", legend = c("vmr", "vhr", "pmr", "phr", "mmr", "mhr"), col = c("#7AC5CD", "#76EEC6", "#CD6600", "#CD6600", "#FF00FF", "#FF00FF"), lty = c(1, 2, 1, 2, 1, 2), lwd = 2, ncol = 3 )
```


```{r}
################################################################################
##
##                                   CALCULATIONS
##
################################################################################
```

```{r}
##                                       FITS
```

```{r eval=run_individual}
fits <- lapply(
  seq_along(data_df),
  function(i) {
    fit_skewed_t(unlist(data_df[i]))
  }
)
```

```{r eval=run_individual}
saveRDS(fits, file = here("data", "fits.RData"))
```

```{r}
##                                     MONTE CARLO
```

```{r eval=(params$run_sim && run_individual)}
fit_id <- list(
  1,
  2,
  3,
  4,
  c(1, 3),
  c(2, 4),
  c(1, 4),
  c(2, 3)
)

mc_output <- list()
for(i in seq_along(fits)) {
  mc_output[[i]] <- mc_simulation(
    fits[fit_id[[i]]], 
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    dao = mc_dao, 
    threshold = mc_threshold
  )
}

saveRDS(mc_output, file = here("data", "mc_output.RData"))
```

```{r eval=(run_individual && !params$run_sim)}
mc_output <- readRDS(file = here("data", paste0("mc_output.RData")))
```

```{r}
##                                     MAX SUM PLOTS
```

```{r eval=(params$run_sim && run_individual)}
max_sum_plots <- list()
for(i in seq_along(fits)) {
  max_sum_plots[[i]] <- plot_max_sum(fits[fit_id[[i]]], mc_num_paths)
}

saveRDS(max_sum_plots, here("data", "max_sum_plots.RData"))
```

```{r eval=(run_individual && !params$run_sim)}
max_sum_plots <- readRDS(file = here("data", "max_sum_plots.RData"))
```

```{r}
##                                  IMPORTANCE SAMPLING
```

```{r eval=(params$run_sim && run_individual)}
is_g_fits <-  list()
is_output <- list()
for(i in seq_along(fits)) {
  is_g_fits[[i]] <- is_proposal(
    x_i_fit = fits[fit_id[[i]]],
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    obj_func_plot = TRUE,
    init_par = c(2, 0.5),
    method = "L-BFGS-B",
    lower = c(1, 0.1),
    upper = c(3, 0.9)
  )
  
  is_output[[i]] <- importance_sampling(
    x_i_fit = fits[fit_id[[i]]], 
    x_n_vect = x_n_vect_from_mc_df(mc_output[[i]]$mc_df),
    num_paths = mc_num_paths, 
    num_periods = mc_num_periods, 
    g_n_params = is_g_fits[[i]]$par, 
    mode = 1
  )
}

saveRDS(is_g_fits, here("data", "is_g_fits.RData"))
saveRDS(is_output, here("data", "is_output.RData"))
```

```{r eval=(run_individual && !params$run_sim)}
is_g_fits <- readRDS(file = here("data", "is_g_fits.RData"))
is_output <- readRDS(file = here("data", "is_output.RData"))
```


```{r}
###########################################################################
##
##                            GENERATE REPORTS
##
###########################################################################
```

```{r eval=params$run_exploratory, results='asis'}
exploratory_report <- knitr::knit_child(
    here("pension-returns_template-exploratory.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(exploratory_report), sep = '\n')
```

```{r eval=params$run_comparison, results='asis'}
comparison_report <- knitr::knit_child(
    here("pension-returns_template-comparison.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(comparison_report), sep = '\n')
```

```{r eval=params$run_individual, results='asis'}
individual_reports <- lapply(
  seq_along(data_df),
  function(i) {
    knitr::knit_child(
      here("pension-returns_template-individual.Rmd"),
      envir = environment(), 
      quiet = TRUE
    )
  }
)

cat(unlist(individual_reports), sep = '\n')
```

```{r eval=params$run_comments, results='asis'}
comments_report <- knitr::knit_child(
    here("pension-returns_template-comments.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(comments_report), sep = '\n')
```

```{r eval=params$run_appendix, results='asis'}
appendix_report <- knitr::knit_child(
    here("pension-returns_template-appendix.Rmd"),
    envir = environment(), 
    quiet = TRUE
  )

cat(unlist(appendix_report), sep = '\n')
```



