## Path crossing and the starting point

In the "Velliv and PFA monthly prices" plot above every plan starts at index 100 in June 2012, and the riskier plans' paths never fall below the less-risky ones. It is tempting to read this as *more risk always pays*. It could look like the different risk plans for each provider are simply different weighted shares of returns from a single portfolio. What is actually going on here?

The observation does not generalise: It is an artefact of **where the index starts**. Each high-risk plan is, in effect, a leveraged exposure to one common return factor. The high-risk profile is, to a good approximation, a *scaled* version of the medium-risk one -- a slope above 1, essentially zero intercept, and a high $R^2$:

```{r}
.cor_v <- cor(data_df$vmr, data_df$vhr); .cor_p <- cor(data_df$pmr, data_df$phr)
.fv <- lm(vhr ~ vmr, data = data_df);    .fp <- lm(phr ~ pmr, data = data_df)
cat(sprintf("Velliv  corr(medium, high) = %.3f ;  high = %.4f + %.2f * medium ,  R^2 = %.3f\n",
            .cor_v, coef(.fv)[2], coef(.fv)[1], summary(.fv)$r.squared))
cat(sprintf("PFA     corr(medium, high) = %.3f ;  high = %.4f + %.2f * medium ,  R^2 = %.3f\n",
            .cor_p, coef(.fp)[1], coef(.fp)[2], summary(.fp)$r.squared))
```

"Approximately" matters: The $R^2$ is below 1, and that residual is where the paths separate. The profiles are not one stream scaled, but blends of two genuinely different funds -- a low-risk and a high-risk building block. Backing PFA's two funds out of the medium and high profiles (high fund $H = D$, low fund $L = 2B - D$, exact for simple returns):

Model each PFA profile as a long-only blend of a low-risk fund $L$ and a high-risk fund $H$, in simple returns: $\text{profile}_X = w_X H + (1-w_X) L$, with $w_X$ the high-risk weight. PFA's product description gives two weights: profile D is 100% the high-risk fund ($w_D=1$) and profile B is a 50/50 blend ($w_B=0.5$). Then  
$$w_D=1 \Rightarrow H=D, \qquad w_B=\tfrac12 \Rightarrow B=\tfrac12H+\tfrac12L \Rightarrow L=2B-D.$$  
These are exact identities at the value level, given the two stated weights. They are not estimated. The data only checks that the reconstructed $L$ is coherent, which it is: low volatility, correlation 0.59 with $H$, volatility ratio 3.1.

```{r}
.simp <- function(lr) exp(lr) - 1
.H <- .simp(data_df$phr); .L <- 2 * .simp(data_df$pmr) - .simp(data_df$phr)
cat(sprintf("PFA building blocks (simple returns): corr(low fund, high fund) = %.2f ; volatility ratio high/low = %.1f\n",
            cor(.L, .H), sd(.H) / sd(.L)))
```

The two underlying funds are only moderately correlated and differ several-fold in volatility -- far from interchangeable -- even though the *profiles* built from them are nearly collinear. That collinearity is not surprising: It is what two equity-dominated portfolios must produce. A "low-risk *plan*" is not a low-risk *fund*. Taking PFA's published 2026 strategic allocation (this is after a structural change in 2024) as an example, even the low-risk plan is roughly 76% equity-like (global, emerging, private equity, real estate) and only ~24% bonds. To see how little bond content it takes, treat a plan as a weight $w$ in equities and $(1-w)$ in bonds and ask which $w$ reproduces a given correlation with a 100%-equity plan (illustrative capital-market assumptions, not estimated from the data):

```{r}
.corr_mix <- function(w, sE, sB, rho)
  (w*sE^2 + (1-w)*rho*sE*sB) /
  (sE * sqrt(w^2*sE^2 + (1-w)^2*sB^2 + 2*w*(1-w)*rho*sE*sB))
.solve_w  <- function(target, sE, sB, rho)
  uniroot(function(w) .corr_mix(w, sE, sB, rho) - target, c(1e-3, 1))$root
.scen <- data.frame(
  `equity vol` = c("15%", "12%", "18%"),
  `bond vol`   = c("4%",  "5%",  "4%"),
  `equity weight for corr 0.986` =
    sprintf("%.0f%%", 100*c(.solve_w(.986,.15,.04,0), .solve_w(.986,.12,.05,0), .solve_w(.986,.18,.04,0))),
  `equity weight for corr 0.84`  =
    sprintf("%.0f%%", 100*c(.solve_w(.84,.15,.04,0),  .solve_w(.84,.12,.05,0),  .solve_w(.84,.18,.04,0))),
  check.names = FALSE)
knitr::kable(.scen)
```

An equity weight of only about two-thirds already yields a ~0.986 correlation with a pure-equity plan. The observed correlations are therefore unremarkable -- exactly what equity-heavy "low-risk" plans produce -- and not evidence of anything hidden.  
  
The implication for the choice is the point of this section. High correlation means the plans rise and fall together, so over a benign, rising sample the higher-risk plan simply ends up higher and the choice looks free. But correlation describes co-movement, not magnitude and not the tails. The lower-risk plan holds a real bond sleeve and a distinct, lower-volatility fund, and -- as the crossings above show -- that cushion is exactly what bites in a drawdown. The plans look alike because both are mostly equities; they are not alike when equities fall.  
  
### One common factor

The correlations point to a single common factor. A principal component analysis of the four
plans confirms it.

```{r pca-plans}
.pca <- prcomp(data_df[, c("vmr", "vhr", "pmr", "phr")], center = TRUE, scale. = TRUE)
.ve  <- .pca$sdev^2 / sum(.pca$sdev^2)
.pca_tab <- data.frame(
  "variance share" = sprintf("%.1f%%", 100 * .ve),
  "cumulative"     = sprintf("%.1f%%", 100 * cumsum(.ve)),
  check.names = FALSE,
  row.names = paste0("PC", seq_along(.ve)))
knitr::kable(.pca_tab, align = "r",
  caption = "Principal components of the four plans' monthly log returns (correlation scaling).")
```

The first component explains 97.7% of the variance and the first two explain 99.4%. The first
component loads almost equally on all four plans, near 0.50 each, so it is a common market
factor. The second component, under two percent, is a Velliv-versus-PFA contrast. To a close
approximation the four plans are therefore one factor seen at four exposures, which is the
structure the leverage reading below assumes. The within-provider relation is the special
case of this: The high plan is a scaled version of the medium plan, as the regression above
shows. PFA's two profiles are in addition fixed blends of the same two building blocks, with
the high-risk fund equal to profile D and the low-risk fund recovered as $2B-D$. That last
decomposition rests on PFA's stated profile weights, not on the data, which only confirm that
the recovered low-risk fund is coherent.

### The crossing condition

For each plan $X\in\{m,h\}$ (medium, high) and period $u=1,2,\dots$ from the chosen start
date, let $s_{X,u}$ be the simple return (gross return $1+s_{X,u}$), $r_{X,u}=\log(1+s_{X,u})$
the log return, $R_{X,t}=\sum_{u=1}^{t}r_{X,u}$ the cumulative log return through period $t$,
and $W_{X,t}=100\,e^{R_{X,t}}$ the wealth index rebased to $100$ at the start. Since
$x\mapsto 100\,e^{x}$ is strictly increasing,
$$
W_{h,t}<W_{m,t}\iff R_{h,t}<R_{m,t},
$$
exactly. What makes this usable is a model for how $R_{h,t}$ relates to $R_{m,t}$.

**Model A (proportionality in log returns).** The regression above fits the affine relation in
log returns, $r_{h,u}=\alpha+\beta\,r_{m,u}+\varepsilon_u$ with $\beta>1$. Summing over
$u=1,\dots,t$ and writing $E_t=\sum_{u\le t}\varepsilon_u$ for the cumulative residual gives
$R_{h,t}=\alpha t+\beta R_{m,t}+E_t$, so
$$
\begin{aligned}
W_{h,t}<W_{m,t}
&\iff \alpha t+\beta R_{m,t}+E_t<R_{m,t}\\
&\iff \alpha t+E_t<(1-\beta)\,R_{m,t}\\
&\iff R_{m,t}<-\frac{\alpha t+E_t}{\beta-1},
\end{aligned}
$$
the last step dividing by $1-\beta<0$, which reverses the inequality. Writing
$\theta_t=-(\alpha t+E_t)/(\beta-1)$ for the threshold, the crossing is $R_{m,t}<\theta_t$. In
the idealisation $\alpha=0$ and $\varepsilon_u\equiv0$ (hence $E_t\equiv0$) the threshold is
$\theta_t=0$, giving the clean rule
$$
W_{h,t}<W_{m,t}\iff R_{m,t}<0\iff W_{m,t}<100:
$$
the high plan sits below the medium plan exactly over those windows in which the medium
(safer) plan is itself below its starting value.

The clean rule replaces $\theta_t$ by $0$; its error $\theta_t$ has an intercept part
$-\alpha t/(\beta-1)$ and a residual part $-E_t/(\beta-1)$. With an intercept in the
regression, least squares forces $\sum_{u=1}^{T}\hat\varepsilon_u=0$ over the full sample of
length $T$, so $E_T=0$ and the residual part vanishes over the whole window; over a strict
sub-window (the COVID rebasing below is one), the partial sum
$E_t=\sum_{u\le t}\hat\varepsilon_u$ is generally nonzero and is the slack between the clean
rule and the truth there.

More fundamentally, $\beta$ is the population least-squares slope: Minimising
$E\!\left[(r_h-a-b\,r_m)^2\right]$ over $(a,b)$ gives the first-order conditions
$a=E[r_h]-b\,E[r_m]$ and $E[r_m r_h]=a\,E[r_m]+b\,E[r_m^2]$, which eliminate $a$ to leave
$\operatorname{Cov}(r_h,r_m)=b\,\operatorname{Var}(r_m)$, hence
$$
\beta=\frac{\operatorname{Cov}(r_h,r_m)}{\operatorname{Var}(r_m)}.
$$
This is defined only when the second moments are finite, and stably estimable only under
stronger conditions. The fitted $\nu\approx3$ to $4$ with a confidence interval reaching
$\nu\le2$, together with the max-sum plots, is precisely the evidence that finite variance
cannot be assumed: The crossing rule is therefore not a geometric identity but a statement
conditional on the existence of the second moments, the same caveat that governs every tail
estimate in this study.

**Model B (constant leverage).** Read "the high plan is a $\beta$-leveraged version of the
medium plan" literally, as a position rebalanced each period to hold constant exposure $\beta$.
Ignoring financing this scales the simple returns, $s_{h,u}=\beta\,s_{m,u}$ (a per-period
borrowing rate $c$ gives $s_{h,u}=\beta s_{m,u}-(\beta-1)c$; take $c=0$). Log returns are now
not proportional, and the crossing condition is exact,
$$
W_{h,t}<W_{m,t}\iff\sum_{u=1}^{t}\bigl[\log(1+\beta s_{m,u})-\log(1+s_{m,u})\bigr]<0
\iff\prod_{u=1}^{t}\frac{1+\beta s_{m,u}}{1+s_{m,u}}<1,
$$
valid while $s_{m,u}>-1/\beta$ (the leveraged plan is not wiped out). Expanding with
$\log(1+x)=x-\tfrac12 x^2+O(x^3)$ gives
$\log(1+\beta s)-\log(1+s)=(\beta-1)s-\tfrac12(\beta^2-1)s^2+O(s^3)$; summing, setting the
total $<0$, dividing by $\beta-1>0$ and using $\beta^2-1=(\beta-1)(\beta+1)$,
$$
\bar s_m<\tfrac12(\beta+1)\,\overline{s_m^{2}},\qquad
\overline{s_m^{2}}=\tfrac1n\sum_{u}s_{m,u}^2=\widehat{\operatorname{Var}}(s_m)+\bar s_m^2
\approx\hat\sigma_m^2 .
$$
For i.i.d. returns with $\mu=E[s_m]$ and $\sigma^2=\operatorname{Var}(s_m)$, the long-run
growth rate is $g=E[\log(1+s)]\approx\mu-\tfrac12\sigma^2$, so $g_m\approx\mu-\tfrac12\sigma^2$
and $g_h\approx\beta\mu-\tfrac12\beta^2\sigma^2$. The leveraged path stays below iff
$g_h<g_m$:
$$
(\beta-1)\mu<\tfrac12(\beta^2-1)\sigma^2\iff\mu<\tfrac12(\beta+1)\sigma^2\iff\beta>2\beta^\star-1,
$$
with the growth-optimal (Kelly) leverage $\beta^\star=\mu/\sigma^2$.

**The two models contrasted.** Both conditions read $\bar s_m<k\,\overline{s_m^{2}}$:
$k=\tfrac12$ for Model A (since
$R_m<0\iff\sum_u\log(1+s_{m,u})<0\iff\bar s_m<\tfrac12\overline{s_m^2}$ to second order) and
$k=\tfrac{\beta+1}{2}$ for Model B. With $\beta>1$ Model B crosses below over a wider set of
paths; in drift terms the thresholds on the medium plan's growth are $g_m<0$ (A) versus
$g_m<\tfrac12\beta\sigma^2$ (B). Model B's is strictly positive (a constant-leverage high
plan can sit below while the medium plan still grows), because leverage multiplies the
variance by $\beta^2$ but the mean only by $\beta$. The two constructions differ in which
variable carries the proportionality:
$$
g_h^{A}=E[\beta r_m]=\beta g_m,\qquad
g_h^{B}=\beta\mu-\tfrac12\beta^2\sigma^2=\beta g_m-\underbrace{\tfrac12\beta(\beta-1)\sigma^2}_{\text{leverage drag}} .
$$
Model A imposes $r_h=\beta r_m$, i.e. $1+s_h=(1+s_m)^{\beta}$, a convex power rule
($s_h=\beta s_m+\tfrac12\beta(\beta-1)s_m^2+O(s^3)$, holding more than $\beta s_m$ in large
moves), and that extra convexity has expectation $\tfrac12\beta(\beta-1)\sigma^2$, exactly
cancelling the drag, so $g_h^{A}=\beta g_m$ with no separate penalty.

For the four plans, the high plan regressed on the medium plan (monthly log returns):

```{r crossing-leverage-numbers}
.lev <- function(lo, hi) {
  rl <- data_df[[lo]]; rh <- data_df[[hi]]               # monthly log returns
  f  <- lm(rh ~ rl)
  sl <- exp(rl) - 1; sh <- exp(rh) - 1                   # simple returns
  b  <- unname(coef(f)[2]); a <- unname(coef(f)[1])
  mu <- mean(sl); v <- var(sl)                           # medium plan, monthly, simple
  list(b = b, a = a,
       r2 = summary(f)$r.squared, r2s = summary(lm(sh ~ sl))$r.squared,
       mu_a = 12 * mu, sd_a = sqrt(12 * v),
       shift = 12 * 0.5 * b * v, drag = 12 * 0.5 * b * (b - 1) * v,
       kelly = mu / v)
}
.v <- .lev("vmr", "vhr"); .p <- .lev("pmr", "phr")
.pc <- function(x) sprintf("%.2f%%", 100 * x)

.fit_tab <- data.frame(
  "$\\beta$"              = sprintf("%.3f",  c(.v$b,   .p$b)),
  "$\\alpha$ (per month)" = sprintf("%+.4f", c(.v$a,   .p$a)),
  "$R^2$ (log)"           = sprintf("%.3f",  c(.v$r2,  .p$r2)),
  "$R^2$ (simple)"        = sprintf("%.3f",  c(.v$r2s, .p$r2s)),
  check.names = FALSE,
  row.names = c("Velliv (medium to high)", "PFA (medium to high)"))
knitr::kable(.fit_tab, align = "r",
  caption = "High plan regressed on medium plan, monthly log returns (n = 142).")

.diag_tab <- data.frame(
  "$\\mu$ (ann.)"    = c(.pc(.v$mu_a),  .pc(.p$mu_a)),
  "$\\sigma$ (ann.)" = c(.pc(.v$sd_a),  .pc(.p$sd_a)),
  "$\\tfrac12\\beta\\sigma^2$"         = c(.pc(.v$shift), .pc(.p$shift)),
  "$\\tfrac12\\beta(\\beta-1)\\sigma^2$" = c(.pc(.v$drag),  .pc(.p$drag)),
  "Kelly $\\beta^\\star$" = sprintf("%.1f",  c(.v$kelly, .p$kelly)),
  check.names = FALSE,
  row.names = c("Velliv", "PFA"))
knitr::kable(.diag_tab, align = "r",
  caption = "Medium-plan moments and leverage diagnostics (annualised).")
```

$\hat\beta>1$ with $\hat\alpha\approx0$ for both, so Model A's clean rule is the operative one.
The log- and simple-return $R^2$ differ by only about $0.001$ because returns are small, so
Models A and B are observationally almost identical here; the distinction is conceptual. The
constant-leverage threshold sits only $\tfrac12\beta\hat\sigma^2\approx0.3\%$ to $0.4\%$ per
year above zero, because $\beta$ is near $1$; for a genuine 2x or 3x product it would
dominate. The in-sample Kelly $\hat\beta^\star\approx11$ to $18$ is not a tangency to chase:
it is $\hat\mu/\hat\sigma^2$ measured on a crash-free window, the ratio of the two
least-robust quantities: The moment-existence caveat once more.
  
June 2012 was a low point just before a long bull market, so the safer plans never
revisited 100 and the paths never crossed. Start the index instead just before a drawdown and the high-risk path falls below. The COVID crash is a clean realised example: Rebasing every plan to 100 at the end of February 2020, the high-risk plans drop **below** the medium-risk plans through the March-2020 trough -- and stay there for about nine months -- before recovering.  

```{r fig.width=10}
## Rebase the cumulative index to 100 at a chosen start month, then plot medium vs high.
crossing_start <- as.Date("2020-02-01")
.cs <- monthly_logreturns_df %>% filter(date >= crossing_start)
.idx <- function(lr) 100 * exp(cumsum(c(0, lr)))     # 100 at the start month
crossing_df <- data.frame(
  date = c(crossing_start, .cs$date),
  `Velliv medium` = .idx(.cs$vmr),
  `Velliv high`   = .idx(.cs$vhr),
  `PFA medium`    = .idx(.cs$pmr),
  `PFA high`      = .idx(.cs$phr),
  check.names = FALSE
) %>% gather(key = "plan", value = "price", -date)

ggplot(crossing_df, aes(x = date, y = price, color = plan)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 100, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("6 months"), labels = date_format("%b %Y")) +
  labs(
    title = "Same plans, index started just before the COVID crash (end-Feb 2020 = 100)",
    subtitle = "High-risk paths fall below medium-risk through the drawdown, then recover",
    x = "date", y = "price index"
  )
```

The high-risk plans regain and overtake the medium-risk plans only once the cumulative return since the start date turns positive again. Whether the realised paths cross is therefore not a property of the plans but of the chosen starting point relative to the market cycle. (A deeper or more prolonged regime -- e.g. one in which the index spends years below its start, as in 2008-2009 -- would keep the high-risk plan below for correspondingly longer.)

```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv and PFA monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```
```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    group_by(date = lubridate::year(date)) %>%
    summarize(
      vmr = sum(vmr),
      vhr = sum(vhr),
      pmr = sum(pmr),
      phr = sum(phr),
      mmr = sum(mmr),
      mhr = sum(mhr),
      vmr_phr = sum(vmr_phr),
      vhr_pmr = sum(vhr_pmr)
    ) %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)),
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  #scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  scale_x_continuous(breaks = lubridate::year(monthly_logreturns_df$date)) +
  labs(
    title = paste("Velliv+PFA annual gross returns"), 
    subtitle = "Aggregated from monthly gross returns", 
    x = "date", 
    y = "gross return"
  )
```


## When the high plan falls behind

Absolute wealth paths are hard to compare near the terminal date. There they converge. The
lead of the high plan over the medium plan is easier to read. It is the high index minus the
medium index, with both rebased to 100 at the entry. It crosses zero exactly when the high
plan falls behind.

Take first a fixed terminal and a varying entry. Invest 100 at every month in the sample. Hold
each entry to the last month in the data. The terminal wealth of an entry at month $i$ is
$W_i=100\,e^{R_i}$ with $R_i=\sum_{u=i}^{t} r_{X_u}$ and $t$ the last month. The lead is the
high terminal wealth minus the medium terminal wealth, read across entry months.

```{r entry-month-lead, fig.width=10}
## Lead of the high plan over the medium plan by entry month, terminal fixed at the last month.
## Invest 100 at each month and hold to the terminal month t. R_i = sum_{u=i}^t r_u, so the
## terminal wealth of 100 entered at month i is W_i = 100 e^{R_i}. The lead is the high
## terminal wealth minus the medium terminal wealth, read across entry months.
.suffix <- function(lr) rev(cumsum(rev(lr)))       # R_i = sum_{u=i}^t lr_u
.tw     <- function(lr) 100 * exp(.suffix(lr))     # terminal wealth of 100 entered at month i
.entry_lead <- data.frame(
  date   = monthly_logreturns_df$date,
  Velliv = .tw(monthly_logreturns_df$vhr) - .tw(monthly_logreturns_df$vmr),
  PFA    = .tw(monthly_logreturns_df$phr) - .tw(monthly_logreturns_df$pmr),
  check.names = FALSE
) %>% gather(key = "provider", value = "lead", -date)

ggplot(.entry_lead, aes(x = date, y = lead, color = provider)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  labs(
    title = "Lead of the high plan over the medium plan by entry month, held to the terminal month",
    subtitle = "High terminal wealth minus medium terminal wealth, 100 invested at each entry; above zero means the high plan is ahead",
    x = "entry month", y = "high minus medium (index points)"
  )
```

The lead never crosses zero. It is largest for the earliest entries. It shrinks toward zero as
the entry nears the terminal month. The reason is the terminal date. Our sample ends near its
high. Every entry has therefore gained by the end. So the high plan leads at every entry.
Running through all entry months shows no crossing.

Now reverse the exercise. Fix the entry at the first month and vary the terminal. Invest 100 at
the first month. Read the lead at each terminal month over the entire period.

```{r high-plan-lead, fig.width=10}
## Lead of the high plan over the medium plan by terminal month, entry fixed at the first month.
## Mirror of the entry-month plot above: Invest 100 at the first month and read the value at
## each terminal month. S_tau = sum_{u=1}^tau r_u, so the value at terminal tau is 100 e^{S_tau}.
## The lead is the high value minus the medium value, read across terminal months.
.prefix <- function(lr) cumsum(lr)                  # S_tau = sum_{u=1}^tau lr_u
.idx    <- function(lr) 100 * exp(.prefix(lr))      # value of 100 invested at the first month
.lead <- data.frame(
  date   = monthly_logreturns_df$date,
  Velliv = .idx(monthly_logreturns_df$vhr) - .idx(monthly_logreturns_df$vmr),
  PFA    = .idx(monthly_logreturns_df$phr) - .idx(monthly_logreturns_df$pmr),
  check.names = FALSE
) %>% gather(key = "provider", value = "lead", -date)

ggplot(.lead, aes(x = date, y = lead, color = provider)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  labs(
    title = "Lead of the high plan over the medium plan by terminal month, entered at the first month",
    subtitle = "High value minus medium value, 100 invested at the first month; above zero means the high plan is ahead",
    x = "terminal month", y = "high minus medium (index points)"
  )
```

The lead never turns negative. It builds over the bull market. It dips through the March-2020
trough but stays above zero. The two plots are mirror images. The full-sample lead sits at the
left edge of the first plot and at the right edge of this one. From the 2012 low looking
forward, and from the 2024 high looking back, the high plan stays ahead.

Neither full-sample view shows a below-zero lead. That does not rule one out. Each view fixes
one endpoint and moves the other. To be certain the lead never turns negative we would have to
run every entry against every terminal, every sub-window of the sample. We do not do that. A
below-zero lead appears whenever the entry sits at an interior peak. The plot below is one
example. The entry is the eve of the COVID crash, end-February 2020, a market peak.

```{r covid-entry-lead, fig.width=10}
## Example of a below-zero lead: A fixed entry at an interior peak, the eve of the COVID crash.
## Invest 100 at the entry and follow the lead forward over the holding period. Both indices
## rebased to 100 at the entry; the lead dips below zero while the medium plan is below its entry.
.entry <- as.Date("2020-02-01")
.cs <- monthly_logreturns_df %>% filter(date >= .entry)
.idx <- function(lr) 100 * exp(cumsum(c(0, lr)))   # rebased to 100 at the entry month
.lead <- data.frame(
  date   = c(.entry, .cs$date),
  Velliv = .idx(.cs$vhr) - .idx(.cs$vmr),
  PFA    = .idx(.cs$phr) - .idx(.cs$pmr),
  check.names = FALSE
) %>% gather(key = "provider", value = "lead", -date)

ggplot(.lead, aes(x = date, y = lead, color = provider)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("6 months"), labels = date_format("%b %Y")) +
  labs(
    title = "Lead of the high plan over the medium plan, entry end-Feb 2020",
    subtitle = "High index minus medium index, both rebased to 100 at entry; below zero means the high plan is behind",
    x = "date", y = "high minus medium (index points)"
  )
```

The lead turns negative through the March-2020 trough. It stays negative for about nine months
before recovering. This is the realised crossing, now visible as the lead dipping below zero.
The condition is the one derived above. The high plan trails the medium plan exactly while the
medium plan is below its entry value.

**A counterfactual.** The non-crossing in the two full-sample views is a feature of this
sample. The sample is a bull market that ends near its high. To see how much rests on that,
reverse every monthly return. Replace each $r_{X_u}$ with $-r_{X_u}$. The counterfactual
terminal wealth is $W'_i=100\,e^{R'_i}$ with $R'_i=\sum_{u=i}^{t} -r_{X_u}$. This is the same
world with the opposite luck. A rising market becomes a falling one. The leverage relation
survives the reversal. The reversed high plan is still a $\beta$-scaled version of the reversed
medium plan, because negating both sides of $h_u=\alpha+\beta m_u$ leaves the slope. Only the
sign of the cumulative return flips.

```{r cf-entry-month-lead, fig.width=10}
## Counterfactual plot A: Every monthly return reversed. R'_i = sum_{u=i}^t (-r_u), so the
## terminal wealth of 100 entered at month i is W'_i = 100 e^{R'_i}. Lead is high minus medium.
.suffix <- function(lr) rev(cumsum(rev(lr)))       # sum_{u=i}^t lr_u
.tw     <- function(lr) 100 * exp(.suffix(lr))     # terminal wealth of 100 entered at month i
.cf_entry_lead <- data.frame(
  date   = monthly_logreturns_df$date,
  Velliv = .tw(-monthly_logreturns_df$vhr) - .tw(-monthly_logreturns_df$vmr),
  PFA    = .tw(-monthly_logreturns_df$phr) - .tw(-monthly_logreturns_df$pmr),
  check.names = FALSE
) %>% gather(key = "provider", value = "lead", -date)

ggplot(.cf_entry_lead, aes(x = date, y = lead, color = provider)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  labs(
    title = "Counterfactual: Lead by entry month with every monthly return reversed",
    subtitle = "Reversed-return terminal wealth, high minus medium; below zero means the high plan is behind",
    x = "entry month", y = "high minus medium (index points)"
  )
```

Now the lead is below zero everywhere. The high plan falls behind at every entry. The mirror
view by terminal month agrees.

```{r cf-terminal-month-lead, fig.width=10}
## Counterfactual plot B: Every monthly return reversed. S'_tau = sum_{u=1}^tau (-r_u), so the
## value of 100 invested at the first month is 100 e^{S'_tau}. Lead is high minus medium.
.prefix <- function(lr) cumsum(lr)                 # sum_{u=1}^tau lr_u
.idx    <- function(lr) 100 * exp(.prefix(lr))     # value of 100 invested at the first month
.cf_term_lead <- data.frame(
  date   = monthly_logreturns_df$date,
  Velliv = .idx(-monthly_logreturns_df$vhr) - .idx(-monthly_logreturns_df$vmr),
  PFA    = .idx(-monthly_logreturns_df$phr) - .idx(-monthly_logreturns_df$pmr),
  check.names = FALSE
) %>% gather(key = "provider", value = "lead", -date)

ggplot(.cf_term_lead, aes(x = date, y = lead, color = provider)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  labs(
    title = "Counterfactual: Lead by terminal month with every monthly return reversed",
    subtitle = "Reversed-return value, high minus medium, 100 invested at the first month; below zero means the high plan is behind",
    x = "terminal month", y = "high minus medium (index points)"
  )
```

The reversed market ends near a low. Every cumulative return is negative. The high plan
amplifies losses exactly as it amplifies gains. So it sits below the medium plan throughout.
The realised non-crossing was the luck of a rising sample. It was never a property of the
plans.

## The downward bound

A pension account cannot fall below zero. In return terms the simple return satisfies
$s\ge-1$: At worst a period wipes out the capital, but it cannot take more. This bound sits
oddly beside a left-skewed and unbounded distribution of log returns. The two are reconciled
by the way wealth compounds. Wealth is $W=100\,e^{R}=100\prod_u(1+s_u)$, and $e^{R}>0$ for
every real $R$, so exponentiating a log-return model can never produce a negative value. The
simulated paths approach zero but never breach it. The wealth distribution is therefore
right-skewed and bounded below at zero, even though the log-return distribution that generates
it is left-skewed and unbounded. The heavy left tail in log space becomes a cluster of paths
near zero, not a set of negative balances.

This also settles which leverage reading applies. Under log-return scaling,
$1+s_h=(1+s_m)^{\beta}$, so as the medium plan approaches a total loss the high plan does too,
and neither breaches the bound. Under constant financial leverage, $s_h=\beta s_m$, a loss in
the medium plan beyond $1/\beta$ would drive the high plan below $-1$ and wipe it out. The
plans never do this, because they are long-only blends: The weights sum to one and no money is
borrowed, so the value is a weighted sum of non-negative fund values and stays non-negative.
The leverage here is higher exposure to the volatile fund, not borrowed money, which is why
the data follow the log-scaling relation rather than constant financial leverage.

## The Sharpe ratio: Leverage ray or curve?

The crossing model $r_{h,u}=\alpha+\beta\,r_{m,u}+\varepsilon_u$ has a second consequence --
about reward per unit of risk rather than path order.

**Definition.** The annualised Sharpe ratio of plan $X$ is $S_X=(\mu_X-r_f)/\sigma_X$, with
$\mu_X$ and $\sigma_X$ the mean and standard deviation of the monthly log returns and $r_f$ the
risk-free rate. Danish short rates are near zero over the sample, so we set $r_f=0$ and
$S_X=\mu_X/\sigma_X$, estimated by $\hat S_X=(\bar x_X/\hat\sigma_X)\sqrt{12}$; the $\sqrt{12}$
annualisation assumes serially independent returns. Geometrically, plotting each plan as a
point $(\sigma_X,\mu_X)$, the Sharpe is the slope of the line from the origin to that point,
so a set of plans shares one Sharpe if and only if the points lie on a single ray through the
origin.

**A ray is the signature of leverage.** Under Model A with a clean fit ($\alpha=0$,
$\varepsilon_u\equiv0$), scaling the log returns by $\beta$ scales both moments by $\beta$ (so
$\mu_h=\beta\mu_m$ and $\sigma_h=\beta\sigma_m$), giving $S_h=\beta\mu_m/(\beta\sigma_m)=S_m$: The
two plans share a Sharpe and lie exactly on a ray. This is the same log-scaling that produced
the clean crossing rule. With a real fit ($\alpha\approx0$, $R^2<1$), let
$\rho=\operatorname{Cov}(r_h,r_m)/(\sigma_h\sigma_m)$ be the correlation; since
$\operatorname{Cov}(r_h,r_m)=\beta\sigma_m^2$ we have $\rho=\beta\sigma_m/\sigma_h$, hence
$$
S_h=\frac{\alpha+\beta\mu_m}{\sigma_h}=\rho\left(S_m+\frac{\alpha}{\beta\sigma_m}\right)\approx\rho\,S_m\quad(\alpha\approx0).
$$
The high plan's Sharpe is the medium plan's Sharpe scaled by the correlation $\rho=\sqrt{R^2}$.
A perfect leverage ray is $\rho=1$ ($S_h=S_m$); the menu bends by exactly the log-collinearity
shortfall $1-\rho$, the same residual the crossing section flagged. (Equivalently, in
simple-return space, constant leverage financed at $r_f$, i.e. Model B, traces the
constant-excess-Sharpe Capital Allocation Line; a ray is the signature of leverage in either
reading.)

**Why the curvature cannot be read off.** Two obstructions, both the study's theme. First,
$S=\mu/\sigma$ requires a finite variance ($\nu>2$); the fitted $\nu$ has a confidence
interval reaching $\nu\le2$, so the denominator may not exist, and where it does not the
Sharpe is not a population quantity. Second, even granting a finite variance, the sampling
variance of the estimator is (Lo 2002; Mertens 2002, under serial independence)
$$
\operatorname{Var}(\hat S)\approx\frac1T\left(1+\tfrac12 S^2-\gamma_3 S+\tfrac14(\gamma_4-3)S^2\right),
$$
with $\gamma_3$ the skewness and $\gamma_4$ the kurtosis, which is finite only for
$\nu>4$. The fitted $\nu\approx3.4$ to $4.3$ straddles $4$: The Sharpe is computable, but its own
standard error need not be finite. The figures below use the i.i.d.-normal standard error
$\operatorname{SE}(\hat S)\approx\sqrt{12\,(1+S^2/24)/n}$, which is a lower bound: The sample
skewness ($\approx-1$) and kurtosis ($\approx6$) inflate it, and the population kurtosis may
be infinite.

```{r sharpe-ray-curve}
.sharpe <- function(x) mean(x) / sd(x) * sqrt(12)   # annualised; rates ~ 0 over the sample
.wins   <- c("2012-06-01", "2016-01-01", "2019-01-01")
.plans  <- c(vmr = "Velliv medium", vhr = "Velliv high",
             pmr = "PFA medium",    phr = "PFA high")
.tab <- t(sapply(.wins, function(w) {
  .s <- monthly_logreturns_df %>% filter(date >= as.Date(w))
  round(sapply(names(.plans), function(p) .sharpe(.s[[p]])), 2)
}))
rownames(.tab) <- paste0("from ", substr(.wins, 1, 7), "  (n=",
  sapply(.wins, function(w) sum(monthly_logreturns_df$date >= as.Date(w))), ")")
colnames(.tab) <- .plans
knitr::kable(.tab, caption = "Annualised Sharpe ratio by plan and start date.")
```

The full-sample decomposition $S_h=\rho(S_m+\alpha/(\beta\sigma_m))$ and the sampling error of
each estimate:

```{r sharpe-leverage-numbers}
.sh <- function(lo, hi) {
  rm <- data_df[[lo]]; rh <- data_df[[hi]]; n <- length(rm)
  Sm  <- mean(rm) / sd(rm) * sqrt(12)                    # annualised Sharpe, medium
  Sh  <- mean(rh) / sd(rh) * sqrt(12)                    # annualised Sharpe, high
  rho <- cor(rm, rh)
  se  <- sqrt(12 * (1 + Sm^2 / 24) / n)                  # i.i.d.-normal SE of annualised Sharpe
  data.frame(
    "$\\rho$"           = sprintf("%.3f", rho),
    "$S_m$"             = sprintf("%.2f", Sm),
    "$S_h$"             = sprintf("%.2f", Sh),
    "$\\rho S_m$"       = sprintf("%.2f", rho * Sm),
    "$\\mathrm{SE}(S)$" = sprintf("%.2f", se),
    "gap / SE"          = sprintf("%.2f", (Sm - Sh) / se),
    check.names = FALSE)
}
.stab <- rbind(.sh("vmr", "vhr"), .sh("pmr", "phr"))
rownames(.stab) <- c("Velliv", "PFA")
knitr::kable(.stab, align = "r",
  caption = "Within-provider Sharpe decomposition and sampling error (full sample, n = 142; annualised; SE i.i.d.-normal, a lower bound).")
```

**Reading.** For Velliv $\hat\rho=0.997$: The medium and high plans are a near-perfect leverage
ray, $\hat S_h$ is reproduced by $\hat\rho(\hat S_m+\hat\alpha/(\hat\beta\hat\sigma_m))$ to the
displayed precision, and the gap $\hat S_m-\hat S_h\approx0.03$ is about $0.1$ of one standard
error. Choosing the Velliv risk level is pure risk appetite, with nothing to optimise: The
same $\beta$-scaling that gives Velliv's clean crossing rule. For PFA $\hat\rho=0.977$: The
menu is measurably more bent (by about $2\%$), but the Sharpe gap is still only about $0.2$ of
a standard error, and which profile leads flips with the start date (the by-window table
above). So although PFA's menu does curve, no best-Sharpe blend is identifiable at this sample
size, and the standard error itself is only a lower bound (the kurtosis it omits may
be infinite). It is not worth paying, in added risk or in fees, to chase a tangency that the
data cannot locate.

The risk-return map shows the same result. Each plan sits at its annualised volatility and mean
return. Plans on one ray through the origin share a Sharpe ratio; a bend off the ray is a Sharpe
difference.

```{r mean-var-plot, fig.width = 8, fig.height = 5}
.mvd <- cbind(velliv_monthly_logreturns_df[, c("low_risk", "medium_risk", "high_risk")],
              pfa_monthly_logreturns_df[, c("pfa_a", "pfa_b", "pfa_c", "pfa_d")])
.mv <- data.frame(
  plan     = c("Velliv low", "Velliv medium", "Velliv high", "PFA A (low)", "PFA B", "PFA C", "PFA D (high)"),
  provider = c(rep("Velliv", 3), rep("PFA", 4)),
  vol = sapply(.mvd, function(x) sd(x) * sqrt(12)),
  ret = sapply(.mvd, function(x) mean(x) * 12))
.vray <- with(.mv[.mv$provider == "Velliv", ], mean(ret / vol))   # mean Velliv Sharpe (rf = 0)
ggplot(.mv, aes(vol, ret, colour = provider)) +
  geom_abline(slope = .vray, intercept = 0, linetype = 2, colour = "grey60") +
  geom_point(size = 3) +
  ggrepel::geom_text_repel(aes(label = plan), size = 3, show.legend = FALSE) +
  labs(title = "Risk vs return by plan (annualised, full sample)",
       subtitle = "Dashed line: The Velliv leverage ray through the origin",
       x = "volatility", y = "mean return") +
  expand_limits(x = 0, y = 0)
```

The three Velliv plans lie close to one ray, the signature of leverage. PFA's profiles sit a
little off it, the small bend the decomposition above quantified.

## Diversifying across providers: Variance drain and certainty equivalent

The crossing and Sharpe results concern one provider's menu. A separate question is what the
return moments say about holding *both* providers at one risk level, for example Velliv high
together with PFA high. This is the diversification the investor tool weighs against fees. We
derive its value here from the moments alone, and the tool applies it to costs.

**Variance drain.** Compound growth sits below the arithmetic mean by half the variance. If a
plan's simple returns have arithmetic mean $a$ and variance $\sigma^2$, wealth compounds at
about $g\approx a-\tfrac12\sigma^2$ per period. Cutting variance therefore lifts the compound
rate one part in two: A reduction $\Delta\sigma^2$ raises $g$ by $\Delta\sigma^2/2$.

**A 50/50 blend cuts variance.** Take two plans of equal variance $\sigma^2$ and correlation
$\rho$. The equal-weight blend has variance $\tfrac12\sigma^2(1+\rho)$, a reduction of
$\tfrac12\sigma^2(1-\rho)$. Write $\sigma^2_{\mathrm{gap}}=\operatorname{Var}(r_A-r_B)=2\sigma^2(1-\rho)$
for the variance of the gap between the two plans. The reduction is then $\sigma^2_{\mathrm{gap}}/4$,
and the compound-growth lift is
$$\Delta g=\tfrac12\cdot\frac{\sigma^2_{\mathrm{gap}}}{4}=\frac{\sigma^2_{\mathrm{gap}}}{8}.$$

**Certainty equivalent.** A saver with constant relative risk aversion $\gamma$ values a variance
cut at $\tfrac12\gamma$ times its size, so the certainty-equivalent value of holding both plans
rather than one is
$$\mathrm{CE}\approx\tfrac12\gamma\cdot\frac{\sigma^2_{\mathrm{gap}}}{4}=\frac{\gamma\,\sigma^2_{\mathrm{gap}}}{8}.$$
The variance drain is the case $\gamma=1$, log utility, where the risk-adjusted value equals the
raw growth lift. The value scales linearly in risk aversion above that.

**Two sources over a horizon.** Over $T$ periods the gap between the two plans' cumulative log
returns has two parts. One is path noise, the accumulation of the per-period gap, which grows
linearly in $T$. The other is uncertainty about the *relative* drift, since which provider
compounds faster is not known in advance, and this grows as $T^2$, so it dominates at long
horizons. Both enter $\sigma^2_{\mathrm{gap}}$ and so the certainty equivalent. The estimation
part is the larger stake for a long-horizon saver, and it is exactly what the data cannot pin
down.

```{r diversification-value}
.divval <- function(a, b, label) {
  gapvar <- var(a - b)                               # per-period (monthly) gap variance
  ce     <- function(g) 1e4 * 12 * g * gapvar / 8    # certainty equivalent, basis points per year
  data.frame(
    blend = label,
    "$\\sigma_{\\mathrm{gap}}$ (ann.)" = sprintf("%.1f%%", 100 * sqrt(12 * gapvar)),
    "$\\Delta g$ (bp/yr)"              = sprintf("%.1f", ce(1)),
    "CE, $\\gamma=2$ (bp/yr)"          = sprintf("%.1f", ce(2)),
    "CE, $\\gamma=4$ (bp/yr)"          = sprintf("%.1f", ce(4)),
    check.names = FALSE)
}
knitr::kable(
  rbind(.divval(monthly_logreturns_df$vhr, monthly_logreturns_df$phr, "Velliv high + PFA high"),
        .divval(monthly_logreturns_df$vmr, monthly_logreturns_df$pmr, "Velliv medium + PFA medium")),
  row.names = FALSE, align = "r",
  caption = "Diversification value of a 50/50 cross-provider blend at one risk level, from the return moments (full sample, n = 142). $\\Delta g$ is the variance drain, equal to the certainty equivalent at $\\gamma=1$; $\\Delta g$ and CE are in basis points per year. A mean-variance, body-of-the-distribution figure.")
```

**A body result, not a tail result.** This is a mean-variance argument, valid for the centre of
the distribution where a variance is a meaningful summary. The report's central finding is that
the tail is heavy and the variance may not be finite, in which case $\sigma^2$ and everything
built on it lose their meaning. The value above is therefore the diversification benefit in the
body. Whether holding both providers helps in a crash is a tail question, and the investor tool
takes it up with a joint simulation.

### The provider gap is one realised path

The certainty equivalent above prices the uncertainty about *which* provider compounds faster.
That uncertainty is real: Over the sample the two high-risk plans drift apart, but by a margin the
data cannot distinguish from zero.

```{r provider-drift}
.dh  <- monthly_logreturns_df$phr - monthly_logreturns_df$vhr   # PFA high minus Velliv high, monthly
.seh <- sd(.dh) / sqrt(length(.dh))
knitr::kable(data.frame(
  "annualised gap" = sprintf("%+.1f pp/yr", 1200 * mean(.dh)),
  "standard error" = sprintf("%.1f pp/yr", 1200 * .seh),
  "t"              = sprintf("%.1f", mean(.dh) / .seh),
  check.names = FALSE),
  caption = "Realised return gap, PFA high minus Velliv high (full sample, n = 142).")
```

```{r provider-drift-plot, fig.width = 10}
.idx <- function(lr) 100 * exp(cumsum(lr))
.pp  <- data.frame(
  date = monthly_logreturns_df$date,
  `Velliv high` = .idx(monthly_logreturns_df$vhr),
  `PFA high`    = .idx(monthly_logreturns_df$phr),
  check.names = FALSE) %>% gather(key = "plan", value = "index", -date)
ggplot(.pp, aes(date, index, colour = plan)) +
  geom_line(linewidth = 0.6) +
  scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  labs(title = "Cumulative index of the two high-risk plans (100 at the start)",
       subtitle = "One realised path over the sample",
       x = "date", y = "price index")
```

The paths separate, yet with $t\approx`r round(mean(.dh) / .seh, 1)`$ the gap is not
distinguishable from zero. Which provider leads is not something the data establish in advance,
which is exactly the uncertainty the certainty equivalent prices.





