## Path crossing and the starting point

In the "Velliv and PFA monthly prices" plot above every plan starts at index 100 in June 2012, and the riskier plans' paths never fall below the less-risky ones. It is tempting to read this as *more risk always pays*. It could look like the different risk plans for each provider are simply different weighted shares of returns from a single portfolio. What is actually going on here?

The observation does not generalise: It is an artefact of **where the index starts**. Each high-risk plan is, in effect, a leveraged exposure to one common return factor. The high-risk profile is, to a good approximation, a *scaled* version of the medium-risk one -- a slope above 1, essentially zero intercept, and a high $R^2$:

```{r}
.cor_v <- cor(data_df$vmr, data_df$vhr); .cor_p <- cor(data_df$pmr, data_df$phr)
.fv <- lm(vhr ~ vmr, data = data_df);    .fp <- lm(phr ~ pmr, data = data_df)
cat(sprintf("Velliv  corr(medium, high) = %.3f ;  high = %.4f + %.2f * medium ,  R^2 = %.3f\n",
            .cor_v, coef(.fv)[2], coef(.fv)[1], summary(.fv)$r.squared))
cat(sprintf("PFA     corr(medium, high) = %.3f ;  high = %.4f + %.2f * medium ,  R^2 = %.3f\n",
            .cor_p, coef(.fp)[1], coef(.fp)[2], summary(.fp)$r.squared))
```

"Approximately" matters: The $R^2$ is below 1, and that residual is where the paths separate. The profiles are not one stream scaled, but blends of two genuinely different funds -- a low-risk and a high-risk building block. Backing PFA's two funds out of the medium and high profiles (high fund $H = D$, low fund $L = 2B - D$, exact for simple returns):

```{r}
.simp <- function(lr) exp(lr) - 1
.H <- .simp(data_df$phr); .L <- 2 * .simp(data_df$pmr) - .simp(data_df$phr)
cat(sprintf("PFA building blocks (simple returns): corr(low fund, high fund) = %.2f ; volatility ratio high/low = %.1f\n",
            cor(.L, .H), sd(.H) / sd(.L)))
```

The two underlying funds are only moderately correlated and differ several-fold in volatility -- far from interchangeable -- even though the *profiles* built from them are nearly collinear. That collinearity is not surprising: It is what two equity-dominated portfolios must produce. A "low-risk *plan*" is not a low-risk *fund*. Taking PFA's published 2026 strategic allocation (this is after a structural change in 2024) as an example, even the low-risk plan is roughly 76% equity-like (global, emerging, private equity, real estate) and only ~24% bonds. To see how little bond content it takes, treat a plan as a weight $w$ in equities and $(1-w)$ in bonds and ask which $w$ reproduces a given correlation with a 100%-equity plan (illustrative capital-market assumptions, not estimated from the data):

```{r}
.corr_mix <- function(w, sE, sB, rho)
  (w*sE^2 + (1-w)*rho*sE*sB) /
  (sE * sqrt(w^2*sE^2 + (1-w)^2*sB^2 + 2*w*(1-w)*rho*sE*sB))
.solve_w  <- function(target, sE, sB, rho)
  uniroot(function(w) .corr_mix(w, sE, sB, rho) - target, c(1e-3, 1))$root
.scen <- data.frame(
  `equity vol` = c("15%", "12%", "18%"),
  `bond vol`   = c("4%",  "5%",  "4%"),
  `equity weight for corr 0.986` =
    sprintf("%.0f%%", 100*c(.solve_w(.986,.15,.04,0), .solve_w(.986,.12,.05,0), .solve_w(.986,.18,.04,0))),
  `equity weight for corr 0.84`  =
    sprintf("%.0f%%", 100*c(.solve_w(.84,.15,.04,0),  .solve_w(.84,.12,.05,0),  .solve_w(.84,.18,.04,0))),
  check.names = FALSE)
knitr::kable(.scen)
```

An equity weight of only about two-thirds already yields a ~0.986 correlation with a pure-equity plan. The observed correlations are therefore unremarkable -- exactly what equity-heavy "low-risk" plans produce -- and not evidence of anything hidden.  
  
The implication for the choice is the point of this section. High correlation means the plans rise and fall together, so over a benign, rising sample the higher-risk plan simply ends up higher and the choice looks free. But correlation describes co-movement, not magnitude and not the tails. The lower-risk plan holds a real bond sleeve and a distinct, lower-volatility fund, and -- as the crossings above show -- that cushion is exactly what bites in a drawdown. The plans look alike because both are mostly equities; they are not alike when equities fall.  
  
### The crossing condition

For each plan $X\in\{m,h\}$ (medium, high) and period $u=1,2,\dots$ from the chosen start
date, let $s_{X,u}$ be the simple return (gross return $1+s_{X,u}$), $r_{X,u}=\log(1+s_{X,u})$
the log return, $R_{X,t}=\sum_{u=1}^{t}r_{X,u}$ the cumulative log return through period $t$,
and $W_{X,t}=100\,e^{R_{X,t}}$ the wealth index rebased to $100$ at the start. Since
$x\mapsto 100\,e^{x}$ is strictly increasing,
$$
W_{h,t}<W_{m,t}\iff R_{h,t}<R_{m,t},
$$
exactly. What makes this usable is a model for how $R_{h,t}$ relates to $R_{m,t}$.

**Model A (proportionality in log returns).** The regression above fits the affine relation in
log returns, $r_{h,u}=\alpha+\beta\,r_{m,u}+\varepsilon_u$ with $\beta>1$. Summing over
$u=1,\dots,t$ and writing $E_t=\sum_{u\le t}\varepsilon_u$ for the cumulative residual gives
$R_{h,t}=\alpha t+\beta R_{m,t}+E_t$, so
$$
\begin{aligned}
W_{h,t}<W_{m,t}
&\iff \alpha t+\beta R_{m,t}+E_t<R_{m,t}\\
&\iff \alpha t+E_t<(1-\beta)\,R_{m,t}\\
&\iff R_{m,t}<-\frac{\alpha t+E_t}{\beta-1},
\end{aligned}
$$
the last step dividing by $1-\beta<0$, which reverses the inequality. Writing
$\theta_t=-(\alpha t+E_t)/(\beta-1)$ for the threshold, the crossing is $R_{m,t}<\theta_t$. In
the idealisation $\alpha=0$ and $\varepsilon_u\equiv0$ (hence $E_t\equiv0$) the threshold is
$\theta_t=0$, giving the clean rule
$$
W_{h,t}<W_{m,t}\iff R_{m,t}<0\iff W_{m,t}<100:
$$
the high plan sits below the medium plan exactly over those windows in which the medium
(safer) plan is itself below its starting value.

The clean rule replaces $\theta_t$ by $0$; its error $\theta_t$ has an intercept part
$-\alpha t/(\beta-1)$ and a residual part $-E_t/(\beta-1)$. With an intercept in the
regression, least squares forces $\sum_{u=1}^{T}\hat\varepsilon_u=0$ over the full sample of
length $T$, so $E_T=0$ and the residual part vanishes over the whole window; over a strict
sub-window -- the COVID rebasing below is one -- the partial sum
$E_t=\sum_{u\le t}\hat\varepsilon_u$ is generally nonzero and is the slack between the clean
rule and the truth there.

More fundamentally, $\beta$ is the population least-squares slope: minimising
$E\!\left[(r_h-a-b\,r_m)^2\right]$ over $(a,b)$ gives the first-order conditions
$a=E[r_h]-b\,E[r_m]$ and $E[r_m r_h]=a\,E[r_m]+b\,E[r_m^2]$, which eliminate $a$ to leave
$\operatorname{Cov}(r_h,r_m)=b\,\operatorname{Var}(r_m)$, hence
$$
\beta=\frac{\operatorname{Cov}(r_h,r_m)}{\operatorname{Var}(r_m)}.
$$
This is defined only when the second moments are finite, and stably estimable only under
stronger conditions. The fitted $\nu\approx3$–$4$ with a confidence interval reaching
$\nu\le2$, together with the max-sum plots, is precisely the evidence that finite variance
cannot be assumed: the crossing rule is therefore not a geometric identity but a statement
conditional on the existence of the second moments -- the same caveat that governs every tail
estimate in this study.

**Model B (constant leverage).** Read "the high plan is a $\beta$-leveraged version of the
medium plan" literally -- a position rebalanced each period to hold constant exposure $\beta$.
Ignoring financing this scales the simple returns, $s_{h,u}=\beta\,s_{m,u}$ (a per-period
borrowing rate $c$ gives $s_{h,u}=\beta s_{m,u}-(\beta-1)c$; take $c=0$). Log returns are now
not proportional, and the crossing condition is exact,
$$
W_{h,t}<W_{m,t}\iff\sum_{u=1}^{t}\bigl[\log(1+\beta s_{m,u})-\log(1+s_{m,u})\bigr]<0
\iff\prod_{u=1}^{t}\frac{1+\beta s_{m,u}}{1+s_{m,u}}<1,
$$
valid while $s_{m,u}>-1/\beta$ (the leveraged plan is not wiped out). Expanding with
$\log(1+x)=x-\tfrac12 x^2+O(x^3)$ gives
$\log(1+\beta s)-\log(1+s)=(\beta-1)s-\tfrac12(\beta^2-1)s^2+O(s^3)$; summing, setting the
total $<0$, dividing by $\beta-1>0$ and using $\beta^2-1=(\beta-1)(\beta+1)$,
$$
\bar s_m<\tfrac12(\beta+1)\,\overline{s_m^{2}},\qquad
\overline{s_m^{2}}=\tfrac1n\sum_{u}s_{m,u}^2=\widehat{\operatorname{Var}}(s_m)+\bar s_m^2
\approx\hat\sigma_m^2 .
$$
For i.i.d. returns with $\mu=E[s_m]$ and $\sigma^2=\operatorname{Var}(s_m)$, the long-run
growth rate is $g=E[\log(1+s)]\approx\mu-\tfrac12\sigma^2$, so $g_m\approx\mu-\tfrac12\sigma^2$
and $g_h\approx\beta\mu-\tfrac12\beta^2\sigma^2$. The leveraged path stays below iff
$g_h<g_m$:
$$
(\beta-1)\mu<\tfrac12(\beta^2-1)\sigma^2\iff\mu<\tfrac12(\beta+1)\sigma^2\iff\beta>2\beta^\star-1,
$$
with the growth-optimal (Kelly) leverage $\beta^\star=\mu/\sigma^2$.

**The two models contrasted.** Both conditions read $\bar s_m<k\,\overline{s_m^{2}}$:
$k=\tfrac12$ for Model A (since
$R_m<0\iff\sum_u\log(1+s_{m,u})<0\iff\bar s_m<\tfrac12\overline{s_m^2}$ to second order) and
$k=\tfrac{\beta+1}{2}$ for Model B. With $\beta>1$ Model B crosses below over a wider set of
paths; in drift terms the thresholds on the medium plan's growth are $g_m<0$ (A) versus
$g_m<\tfrac12\beta\sigma^2$ (B). Model B's is strictly positive -- a constant-leverage high
plan can sit below while the medium plan still grows -- because leverage multiplies the
variance by $\beta^2$ but the mean only by $\beta$. The two constructions differ in which
variable carries the proportionality:
$$
g_h^{A}=E[\beta r_m]=\beta g_m,\qquad
g_h^{B}=\beta\mu-\tfrac12\beta^2\sigma^2=\beta g_m-\underbrace{\tfrac12\beta(\beta-1)\sigma^2}_{\text{leverage drag}} .
$$
Model A imposes $r_h=\beta r_m$, i.e. $1+s_h=(1+s_m)^{\beta}$ -- a convex power rule,
$s_h=\beta s_m+\tfrac12\beta(\beta-1)s_m^2+O(s^3)$, holding more than $\beta s_m$ in large
moves -- and that extra convexity has expectation $\tfrac12\beta(\beta-1)\sigma^2$, exactly
cancelling the drag, so $g_h^{A}=\beta g_m$ with no separate penalty.

For the four plans, the high plan regressed on the medium plan (monthly log returns):

```{r crossing-leverage-numbers}
.lev <- function(lo, hi) {
  rl <- data_df[[lo]]; rh <- data_df[[hi]]               # monthly log returns
  f  <- lm(rh ~ rl)
  sl <- exp(rl) - 1; sh <- exp(rh) - 1                   # simple returns
  b  <- unname(coef(f)[2]); a <- unname(coef(f)[1])
  mu <- mean(sl); v <- var(sl)                           # medium plan, monthly, simple
  list(b = b, a = a,
       r2 = summary(f)$r.squared, r2s = summary(lm(sh ~ sl))$r.squared,
       mu_a = 12 * mu, sd_a = sqrt(12 * v),
       shift = 12 * 0.5 * b * v, drag = 12 * 0.5 * b * (b - 1) * v,
       kelly = mu / v)
}
.v <- .lev("vmr", "vhr"); .p <- .lev("pmr", "phr")
.pc <- function(x) sprintf("%.2f%%", 100 * x)

.fit_tab <- data.frame(
  "β"              = sprintf("%.3f",  c(.v$b,   .p$b)),
  "α (per month)"  = sprintf("%+.4f", c(.v$a,   .p$a)),
  "R² (log)"       = sprintf("%.3f",  c(.v$r2,  .p$r2)),
  "R² (simple)"    = sprintf("%.3f",  c(.v$r2s, .p$r2s)),
  check.names = FALSE,
  row.names = c("Velliv (medium→high)", "PFA (medium→high)"))
knitr::kable(.fit_tab, align = "r",
  caption = "High plan regressed on medium plan, monthly log returns (n = 142).")

.diag_tab <- data.frame(
  "μ (ann.)"        = c(.pc(.v$mu_a),  .pc(.p$mu_a)),
  "σ (ann.)"        = c(.pc(.v$sd_a),  .pc(.p$sd_a)),
  "½βσ²"            = c(.pc(.v$shift), .pc(.p$shift)),
  "½β(β-1)σ²"       = c(.pc(.v$drag),  .pc(.p$drag)),
  "Kelly β*"        = sprintf("%.1f",  c(.v$kelly, .p$kelly)),
  check.names = FALSE,
  row.names = c("Velliv", "PFA"))
knitr::kable(.diag_tab, align = "r",
  caption = "Medium-plan moments and leverage diagnostics (annualised).")
```

$\hat\beta>1$ with $\hat\alpha\approx0$ for both, so Model A's clean rule is the operative one.
The log- and simple-return $R^2$ differ by only about $0.001$ because returns are small, so
Models A and B are observationally almost identical here; the distinction is conceptual. The
constant-leverage threshold sits only $\tfrac12\beta\hat\sigma^2\approx0.3$–$0.4\%$ per year
above zero, because $\beta$ is near $1$; for a genuine 2x–3x product it would dominate. The
in-sample Kelly $\hat\beta^\star\approx11$–$18$ is not a tangency to chase: it is
$\hat\mu/\hat\sigma^2$ measured on a crash-free window, the ratio of the two least-robust
quantities -- the moment-existence caveat once more.
  
June 2012 was a low point just before a long bull market, so the safer plans never
revisited 100 and the paths never crossed. Start the index instead just before a drawdown and the high-risk path falls below. The COVID crash is a clean realised example: rebasing every plan to 100 at the end of February 2020, the high-risk plans drop **below** the medium-risk plans through the March-2020 trough -- and stay there for about nine months -- before recovering.  

```{r fig.width=10}
## Rebase the cumulative index to 100 at a chosen start month, then plot medium vs high.
crossing_start <- as.Date("2020-02-01")
.cs <- monthly_logreturns_df %>% filter(date >= crossing_start)
.idx <- function(lr) 100 * exp(cumsum(c(0, lr)))     # 100 at the start month
crossing_df <- data.frame(
  date = c(crossing_start, .cs$date),
  `Velliv medium` = .idx(.cs$vmr),
  `Velliv high`   = .idx(.cs$vhr),
  `PFA medium`    = .idx(.cs$pmr),
  `PFA high`      = .idx(.cs$phr),
  check.names = FALSE
) %>% gather(key = "plan", value = "price", -date)

ggplot(crossing_df, aes(x = date, y = price, color = plan)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 100, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("6 months"), labels = date_format("%b %Y")) +
  labs(
    title = "Same plans, index started just before the COVID crash (end-Feb 2020 = 100)",
    subtitle = "High-risk paths fall below medium-risk through the drawdown, then recover",
    x = "date", y = "price index"
  )
```

The high-risk plans regain and overtake the medium-risk plans only once the cumulative return since the start date turns positive again. Whether the realised paths cross is therefore not a property of the plans but of the chosen starting point relative to the market cycle. (A deeper or more prolonged regime -- e.g. one in which the index spends years below its start, as in 2008-2009 -- would keep the high-risk plan below for correspondingly longer.)

```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv and PFA monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```
```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    group_by(date = lubridate::year(date)) %>%
    summarize(
      vmr = sum(vmr),
      vhr = sum(vhr),
      pmr = sum(pmr),
      phr = sum(phr),
      mmr = sum(mmr),
      mhr = sum(mhr),
      vmr_phr = sum(vmr_phr),
      vhr_pmr = sum(vhr_pmr)
    ) %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)),
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  #scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  scale_x_continuous(breaks = lubridate::year(monthly_logreturns_df$date)) +
  labs(
    title = paste("Velliv+PFA annual gross returns"), 
    subtitle = "Aggregated from monthly gross returns", 
    x = "date", 
    y = "gross return"
  )
```


## The Sharpe ratio: leverage ray or curve?

The crossing model $r_{h,u}=\alpha+\beta\,r_{m,u}+\varepsilon_u$ has a second consequence --
about reward per unit of risk rather than path order.

**Definition.** The annualised Sharpe ratio of plan $X$ is $S_X=(\mu_X-r_f)/\sigma_X$, with
$\mu_X$ and $\sigma_X$ the mean and standard deviation of the monthly log returns and $r_f$ the
risk-free rate. Danish short rates are near zero over the sample, so we set $r_f=0$ and
$S_X=\mu_X/\sigma_X$, estimated by $\hat S_X=(\bar x_X/\hat\sigma_X)\sqrt{12}$; the $\sqrt{12}$
annualisation assumes serially independent returns. Geometrically, plotting each plan as a
point $(\sigma_X,\mu_X)$, the Sharpe is the slope of the line from the origin to that point,
so a set of plans shares one Sharpe if and only if the points lie on a single ray through the
origin.

**A ray is the signature of leverage.** Under Model A with a clean fit ($\alpha=0$,
$\varepsilon_u\equiv0$), scaling the log returns by $\beta$ scales both moments by $\beta$ --
$\mu_h=\beta\mu_m$, $\sigma_h=\beta\sigma_m$ -- so $S_h=\beta\mu_m/(\beta\sigma_m)=S_m$: the
two plans share a Sharpe and lie exactly on a ray. This is the same log-scaling that produced
the clean crossing rule. With a real fit ($\alpha\approx0$, $R^2<1$), let
$\rho=\operatorname{Cov}(r_h,r_m)/(\sigma_h\sigma_m)$ be the correlation; since
$\operatorname{Cov}(r_h,r_m)=\beta\sigma_m^2$ we have $\rho=\beta\sigma_m/\sigma_h$, hence
$$
S_h=\frac{\alpha+\beta\mu_m}{\sigma_h}=\rho\left(S_m+\frac{\alpha}{\beta\sigma_m}\right)\approx\rho\,S_m\quad(\alpha\approx0).
$$
The high plan's Sharpe is the medium plan's Sharpe scaled by the correlation $\rho=\sqrt{R^2}$.
A perfect leverage ray is $\rho=1$ ($S_h=S_m$); the menu bends by exactly the log-collinearity
shortfall $1-\rho$ -- the same residual the crossing section flagged. (Equivalently, in
simple-return space, constant leverage financed at $r_f$ -- Model B -- traces the
constant-excess-Sharpe Capital Allocation Line; a ray is the signature of leverage in either
reading.)

**Why the curvature cannot be read off.** Two obstructions, both the study's theme. First,
$S=\mu/\sigma$ requires a finite variance ($\nu>2$); the fitted $\nu$ has a confidence
interval reaching $\nu\le2$, so the denominator may not exist, and where it does not the
Sharpe is not a population quantity. Second, even granting a finite variance, the sampling
variance of the estimator is (Lo 2002; Mertens 2002, under serial independence)
$$
\operatorname{Var}(\hat S)\approx\frac1T\left(1+\tfrac12 S^2-\gamma_3 S+\tfrac14(\gamma_4-3)S^2\right),
$$
with $\gamma_3$ the skewness and $\gamma_4$ the kurtosis -- and $\gamma_4$ is finite only for
$\nu>4$. The fitted $\nu\approx3.4$–$4.3$ straddles $4$: the Sharpe is computable, but its own
standard error need not be finite. The figures below use the i.i.d.-normal standard error
$\operatorname{SE}(\hat S)\approx\sqrt{12\,(1+S^2/24)/n}$, which is a lower bound -- the sample
skewness ($\approx-1$) and kurtosis ($\approx6$) inflate it, and the population kurtosis may
be infinite.

```{r sharpe-ray-curve}
.sharpe <- function(x) mean(x) / sd(x) * sqrt(12)   # annualised; rates ~ 0 over the sample
.wins   <- c("2012-06-01", "2016-01-01", "2019-01-01")
.plans  <- c(vmr = "Velliv medium", vhr = "Velliv high",
             pmr = "PFA medium",    phr = "PFA high")
.tab <- t(sapply(.wins, function(w) {
  .s <- monthly_logreturns_df %>% filter(date >= as.Date(w))
  round(sapply(names(.plans), function(p) .sharpe(.s[[p]])), 2)
}))
rownames(.tab) <- paste0("from ", substr(.wins, 1, 7), "  (n=",
  sapply(.wins, function(w) sum(monthly_logreturns_df$date >= as.Date(w))), ")")
colnames(.tab) <- .plans
knitr::kable(.tab, caption = "Annualised Sharpe ratio by plan and start date.")
```

The full-sample decomposition $S_h=\rho(S_m+\alpha/(\beta\sigma_m))$ and the sampling error of
each estimate:

```{r sharpe-leverage-numbers}
.sh <- function(lo, hi) {
  rm <- data_df[[lo]]; rh <- data_df[[hi]]; n <- length(rm)
  Sm  <- mean(rm) / sd(rm) * sqrt(12)                    # annualised Sharpe, medium
  Sh  <- mean(rh) / sd(rh) * sqrt(12)                    # annualised Sharpe, high
  rho <- cor(rm, rh)
  se  <- sqrt(12 * (1 + Sm^2 / 24) / n)                  # i.i.d.-normal SE of annualised Sharpe
  data.frame(
    "ρ (corr)"  = sprintf("%.3f", rho),
    "S med"     = sprintf("%.2f", Sm),
    "S high"    = sprintf("%.2f", Sh),
    "ρ × S med" = sprintf("%.2f", rho * Sm),
    "SE(S)"     = sprintf("%.2f", se),
    "gap / SE"  = sprintf("%.2f", (Sm - Sh) / se),
    check.names = FALSE)
}
.stab <- rbind(.sh("vmr", "vhr"), .sh("pmr", "phr"))
rownames(.stab) <- c("Velliv", "PFA")
knitr::kable(.stab, align = "r",
  caption = "Within-provider Sharpe decomposition and sampling error (full sample, n = 142; annualised; SE i.i.d.-normal, a lower bound).")
```

**Reading.** For Velliv $\hat\rho=0.997$: the medium and high plans are a near-perfect leverage
ray, $\hat S_h$ is reproduced by $\hat\rho(\hat S_m+\hat\alpha/(\hat\beta\hat\sigma_m))$ to the
displayed precision, and the gap $\hat S_m-\hat S_h\approx0.03$ is about $0.1$ of one standard
error. Choosing the Velliv risk level is pure risk appetite, with nothing to optimise -- the
same $\beta$-scaling that gives Velliv's clean crossing rule. For PFA $\hat\rho=0.977$: the
menu is measurably more bent (by about $2\%$), but the Sharpe gap is still only about $0.2$ of
a standard error, and which profile leads flips with the start date (the by-window table
above). So although PFA's menu does curve, no best-Sharpe blend is identifiable at this sample
size -- and the standard error itself is only a lower bound, since the kurtosis it omits may
be infinite. It is not worth paying, in added risk or in fees, to chase a tangency that the
data cannot locate.





