## Path crossing and the starting point

In the "Velliv and PFA monthly prices" plot above every plan starts at index 100 in June 2012, and the riskier plans' paths never fall below the less-risky ones. It is tempting to read this as *more risk always pays*. It could look like the different risk plans for each provider are simply different weighted shares of returns from a single portfolio. What is actually going on here?

The observation does not generalise: It is an artefact of **where the index starts**. Each high-risk plan is, in effect, a leveraged exposure to one common return factor. The high-risk profile is, to a good approximation, a *scaled* version of the medium-risk one -- a slope above 1, essentially zero intercept, and a high $R^2$:

```{r}
.cor_v <- cor(data_df$vmr, data_df$vhr); .cor_p <- cor(data_df$pmr, data_df$phr)
.fv <- lm(vhr ~ vmr, data = data_df);    .fp <- lm(phr ~ pmr, data = data_df)
cat(sprintf("Velliv  corr(medium, high) = %.3f ;  high = %.4f + %.2f * medium ,  R^2 = %.3f\n",
            .cor_v, coef(.fv)[2], coef(.fv)[1], summary(.fv)$r.squared))
cat(sprintf("PFA     corr(medium, high) = %.3f ;  high = %.4f + %.2f * medium ,  R^2 = %.3f\n",
            .cor_p, coef(.fp)[1], coef(.fp)[2], summary(.fp)$r.squared))
```

"Approximately" matters: The $R^2$ is below 1, and that residual is where the paths separate. The profiles are not one stream scaled, but blends of two genuinely different funds -- a low-risk and a high-risk building block. Backing PFA's two funds out of the medium and high profiles (high fund $H = D$, low fund $L = 2B - D$, exact for simple returns):

```{r}
.simp <- function(lr) exp(lr) - 1
.H <- .simp(data_df$phr); .L <- 2 * .simp(data_df$pmr) - .simp(data_df$phr)
cat(sprintf("PFA building blocks (simple returns): corr(low fund, high fund) = %.2f ; volatility ratio high/low = %.1f\n",
            cor(.L, .H), sd(.H) / sd(.L)))
```

The two underlying funds are only moderately correlated and differ several-fold in volatility -- far from interchangeable -- even though the *profiles* built from them are nearly collinear. That collinearity is not surprising: It is what two equity-dominated portfolios must produce. A "low-risk *plan*" is not a low-risk *fund*. Taking PFA's published 2026 strategic allocation (this is after a structural change in 2024) as an example, even the low-risk plan is roughly 76% equity-like (global, emerging, private equity, real estate) and only ~24% bonds. To see how little bond content it takes, treat a plan as a weight $w$ in equities and $(1-w)$ in bonds and ask which $w$ reproduces a given correlation with a 100%-equity plan (illustrative capital-market assumptions, not estimated from the data):

```{r}
.corr_mix <- function(w, sE, sB, rho)
  (w*sE^2 + (1-w)*rho*sE*sB) /
  (sE * sqrt(w^2*sE^2 + (1-w)^2*sB^2 + 2*w*(1-w)*rho*sE*sB))
.solve_w  <- function(target, sE, sB, rho)
  uniroot(function(w) .corr_mix(w, sE, sB, rho) - target, c(1e-3, 1))$root
.scen <- data.frame(
  `equity vol` = c("15%", "12%", "18%"),
  `bond vol`   = c("4%",  "5%",  "4%"),
  `equity weight for corr 0.986` =
    sprintf("%.0f%%", 100*c(.solve_w(.986,.15,.04,0), .solve_w(.986,.12,.05,0), .solve_w(.986,.18,.04,0))),
  `equity weight for corr 0.84`  =
    sprintf("%.0f%%", 100*c(.solve_w(.84,.15,.04,0),  .solve_w(.84,.12,.05,0),  .solve_w(.84,.18,.04,0))),
  check.names = FALSE)
knitr::kable(.scen)
```

An equity weight of only about two-thirds already yields a ~0.986 correlation with a pure-equity plan. The observed correlations are therefore unremarkable -- exactly what equity-heavy "low-risk" plans produce -- and not evidence of anything hidden.  
  
The implication for the choice is the point of this section. High correlation means the plans rise and fall together, so over a benign, rising sample the higher-risk plan simply ends up higher and the choice looks free. But correlation describes co-movement, not magnitude and not the tails. The lower-risk plan holds a real bond sleeve and a distinct, lower-volatility fund, and -- as the crossings above show -- that cushion is exactly what bites in a drawdown. The plans look alike because both are mostly equities; they are not alike when equities fall.  
  
In cumulative log terms $\log(W_\text{high}/100) \approx \beta\,\log(W_\text{med}/100)$ with slope $\beta > 1$ and intercept $\approx 0$. With $\beta > 1$ this gives $W_\text{high} > W_\text{med}$ **exactly when** $W_\text{med} > 100$. So the two paths can only cross while the *safer* plan is itself below its starting value -- equivalently, the high-risk plan sits below the medium-risk plan precisely over any window in which the cumulative return *since the start date* is negative.  
  
June 2012 was a low point just before a long bull market, so the safer plans never
revisited 100 and the paths never crossed. Start the index instead just before a drawdown and the high-risk path falls below. The COVID crash is a clean realised example: rebasing every plan to 100 at the end of February 2020, the high-risk plans drop **below** the medium-risk plans through the March-2020 trough -- and stay there for about nine months -- before recovering.  

```{r fig.width=10}
## Rebase the cumulative index to 100 at a chosen start month, then plot medium vs high.
crossing_start <- as.Date("2020-02-01")
.cs <- monthly_logreturns_df %>% filter(date >= crossing_start)
.idx <- function(lr) 100 * exp(cumsum(c(0, lr)))     # 100 at the start month
crossing_df <- data.frame(
  date = c(crossing_start, .cs$date),
  `Velliv medium` = .idx(.cs$vmr),
  `Velliv high`   = .idx(.cs$vhr),
  `PFA medium`    = .idx(.cs$pmr),
  `PFA high`      = .idx(.cs$phr),
  check.names = FALSE
) %>% gather(key = "plan", value = "price", -date)

ggplot(crossing_df, aes(x = date, y = price, color = plan)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 100, linetype = 2, colour = "grey50") +
  scale_x_date(breaks = date_breaks("6 months"), labels = date_format("%b %Y")) +
  labs(
    title = "Same plans, index started just before the COVID crash (end-Feb 2020 = 100)",
    subtitle = "High-risk paths fall below medium-risk through the drawdown, then recover",
    x = "date", y = "price index"
  )
```

The high-risk plans regain and overtake the medium-risk plans only once the cumulative return since the start date turns positive again. Whether the realised paths cross is therefore not a property of the plans but of the chosen starting point relative to the market cycle. (A deeper or more prolonged regime -- e.g. one in which the index spends years below its start, as in 2008-2009 -- would keep the high-risk plan below for correspondingly longer.)

```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)), 
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  scale_x_date(breaks=date_breaks("2 years"), labels=date_format("%Y")) +
  labs(
    title = paste("Velliv and PFA monthly gross returns"), 
    subtitle = "", 
    x = "date", 
    y = "gross return"
  )
```
```{r fig.width=10}
ggplot(
  monthly_logreturns_df %>% 
    group_by(date = lubridate::year(date)) %>%
    summarize(
      vmr = sum(vmr),
      vhr = sum(vhr),
      pmr = sum(pmr),
      phr = sum(phr),
      mmr = sum(mmr),
      mhr = sum(mhr),
      vmr_phr = sum(vmr_phr),
      vhr_pmr = sum(vhr_pmr)
    ) %>% 
    gather(key = "fund", value = "gross_return", -date) %>% 
    mutate(gross_return = exp(gross_return)),
  aes(x = date, y = gross_return)
) +
  geom_line(aes(color = fund), linewidth = 0.5) +
  #scale_x_date(breaks = date_breaks("2 years"), labels = date_format("%Y")) +
  scale_x_continuous(breaks = lubridate::year(monthly_logreturns_df$date)) +
  labs(
    title = paste("Velliv+PFA annual gross returns"), 
    subtitle = "Aggregated from monthly gross returns", 
    x = "date", 
    y = "gross return"
  )
```


## The Sharpe ratio: Leverage ray or curve?

The crossing argument established that, within a provider, the high-risk plan is to a good approximation a *scaled* version of the medium-risk one. If that scaling were **pure leverage**, the two plans would share the same reward-to-risk ratio -- the same annualised Sharpe -- and the risk level would carry no consequence for return *per unit of risk*. The test is whether the Sharpe ratios coincide, and whether that holds across sub-periods rather than being an artefact of one window. (Danish short rates over the sample are near zero, so we treat the return as the excess return.)

```{r sharpe-ray-curve}
.sharpe <- function(x) mean(x) / sd(x) * sqrt(12)   # annualised; rates ~ 0 over the sample
.wins   <- c("2012-06-01", "2016-01-01", "2019-01-01")
.plans  <- c(vmr = "Velliv medium", vhr = "Velliv high",
             pmr = "PFA medium",    phr = "PFA high")
.tab <- t(sapply(.wins, function(w) {
  .s <- monthly_logreturns_df %>% filter(date >= as.Date(w))
  round(sapply(names(.plans), function(p) .sharpe(.s[[p]])), 2)
}))
rownames(.tab) <- paste0("from ", substr(.wins, 1, 7), "  (n=",
  sapply(.wins, function(w) sum(monthly_logreturns_df$date >= as.Date(w))), ")")
colnames(.tab) <- .plans
knitr::kable(.tab, caption = "Annualised Sharpe ratio by plan and start date.")
```

**Velliv behaves like a leverage ray.** Its medium and high plans have nearly the same Sharpe in every sub-period -- the gap stays within a few hundredths and all but vanishes in the most recent window -- the signature of one portfolio held at two levels of exposure. Choosing the Velliv risk level is therefore a pure risk-appetite decision: More risk buys proportionally more expected return, with no change in reward per unit of risk and nothing to "optimise".

**PFA's two profiles curve -- but which one is better is not identifiable.** PFA's medium and high profiles do *not* share a Sharpe, so the menu bends rather than running straight; yet *which* end looks better flips with the start date -- the medium profile has the higher Sharpe measured from 2012, the high profile from 2019, with the two essentially tied in between. The differences sit well inside the sampling noise documented elsewhere in this report. The textbook response to a curved menu is to locate the higher-Sharpe blend; here you cannot reliably tell which blend that is, so it is not worth paying -- in added risk or in fees -- to chase one.





